### 2022-08-05 appendix/appendix_a.Rmd
Add files via upload

# Participant background information

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
h1 = c("No", "Status", "Sex", "AOA", "Age", "LOE", "Deaf schools", "Home language", "Parents' hearing")

df_table_a <- read.csv("appendix_a_table.csv", sep = ";")
names(df_table_a) <- h1
apa_table(df_table_a,
          caption ="Participant demographic information based on self-report",
          note = "AOA = age of acquisition, LOE = level of education, HS = high school, AD = associate degree, MS = middle school, BA/BS = Bachelor of Arts/Science, TR = (Spoken) Turkish.",
          escape = FALSE,
          placement = "h",
          font_size = "footnotesize",
          landscape = FALSE
)
```