### 2022-08-05 appendix/appendix_a.tex

\clearpage

\begin{appendix}
\hypertarget{participant-background-information}{%
\section{Participant background
information}\label{participant-background-information}}

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-12}Participant demographic information based on self-report}

\footnotesize{

\begin{tabular}{lllllllll}
\toprule
No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{Deaf schools} & \multicolumn{1}{c}{Home language} & \multicolumn{1}{c}{Parents' hearing}\\
\midrule
1 & Native & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign & Both deaf\\
2 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign & Both deaf\\
3 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign & Both deaf\\
4 & Native & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign & Both deaf\\
5 & Native & M & 0-3 & 28 & 12 & PS, MS & TİD & Both deaf\\
6 & Native & M & 0-3 & 32 & 12 & PS, MS & TİD & F deaf, M hearing\\
7 & Native & M & 0-3 & 35 & 12 & PS, MS, HS & TİD & Both deaf\\
8 & Native & F & 0-3 & 23 & 12 & PS, MS, HS & TİD & Both deaf\\
9 & Native & F & 0-3 & 30 & 12 & PS, MS, HS & TİD & Both deaf\\
10 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD & Both deaf\\
11 & Native & F & 0-3 & 24 & 14 & PS, MS, HS & TİD & Both deaf\\
12 & Native & M & 0-3 & 18 & 12 & PS, MS & TİD & Both deaf\\
13 & Native & F & 0-3 & 21 & 16 & PS, MS, HS & TİD & Both deaf\\
14 & Native & M & 0-3 & 19 & 16 & PS, MS, HS & TİD & Both deaf\\
15 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD & Both deaf\\
16 & Late & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR & Both hearing\\
17 & Late & M & 4-7 & 36 & 12 & MS, HS & TR, home sign & Both hearing\\
18 & Late & M & 4-7 & 33 & 12 & PS, HS & TR & Both hearing\\
19 & Late & F & 4-7 & 33 & 12 & PS & TR, home sign & Both hearing\\
20 & Late & F & 4-7 & 31 & 14 & PS, MS & Home sign & Both hearing\\
21 & Late & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign & Both hearing\\
22 & Late & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR & Both hearing\\
23 & Late & M & 8-12 & 43 & 16 & PS, MS & TR & Both hearing\\
24 & Late & M & 8-12 & 28 & 16 & MS & TR & Both hearing\\
25 & Late & F & 8-12 & 24 & 12 & PS, MS, HS & TR, TİD & Both hearing\\
26 & Late & F & 8-12 & 24 & 8 & PS, MS & Home sign & Both hearing\\
27 & Late & M & 8-12 & 33 & 12 & PS, MS & Home sign & Both hearing\\
28 & Late & M & 13-17 & 50 & 8 & MS & TR, home sign & Both hearing\\
29 & Late & M & 13-17 & 31 & 16 & PS, MS, HS & TR & Both hearing\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, HS = high school, AD = associate degree, MS = middle school, BA/BS = Bachelor of Arts/Science, TR = (Spoken) Turkish.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}
\end{appendix}