### 2022-08-05 appendix/tables/edited_table.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  doc,mask]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Age of Acquisition Effects on Reference Tracking in Turkish Sign Language (TİD)},
  pdflang={en-EN},
  pdfkeywords={Reference tracking, Turkish Sign Language (TİD), Age of acqusition, Signed narrative, Pragmatic competence},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{AOA Effects on Reference Tracking in TİD}
\keywords{Reference tracking, Turkish Sign Language (TİD), Age of acqusition, Signed narrative, Pragmatic competence\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{floatrow}
\floatsetup[table]{capposition=top}
\floatsetup[figure]{capposition=top}
\usepackage{times}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{langsci-gb4e}
\usepackage{multicol}
\setlength{\columnsep}{-3cm}
\usepackage{setspace}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\AtBeginEnvironment{tablenotes}{\doublespacing}
\captionsetup[table]{font={stretch=1.5}}
\captionsetup[figure]{font={stretch=1.5}}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Age of Acquisition Effects on Reference Tracking in Turkish Sign Language (TİD)}
\author{Onur Keleş\textsuperscript{1}, Furkan Atmaca\textsuperscript{2}, \& Kadir Gökgöz\textsuperscript{1}}
\date{}

\authornote{

This work was supported by Boğaziçi Üniversitesi Bilimsel Araştırma Projeleri {[}Boğaziçi University Research Fund{]}, {[}14458{]}. We thank all of our 31 participants.

The authors made the following contributions. Onur Keleş: Conceptualization, Data Curation, Investigation, Methodology, Writing - Original Draft Preparation, Writing - Review \& Editing; Furkan Atmaca: Formal Analysis, Methodology, Software, Validation, Visualization, Writing - Review \& Editing; Kadir Gökgöz: Funding Acquisition, Project Administration, Supervision, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Onur Keleş, Bebek, 34342 Beşiktaş/İstanbul. E-mail: \href{mailto:keles.onur@hotmail.com}{\nolinkurl{keles.onur@hotmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Boğaziçi University\\\textsuperscript{2} Leipzig University}

\abstract{
We investigated the reference tracking strategies among native and late adult deaf signers of Turkish Sign Language (TİD). Given the previous research that suggests over-redundancy in reference tracking and poorer morphosyntactic competence for late signers, we expected that late signers use fewer instances of null marking and mixing of perspectives in their tracking than the native group. Using a story-telling paradigm, we tested 15 native and 16 late signers. Participants were shown 10 short wordless video clips from a cartoon, and asked to retell them. The overall results indicated that the observed distribution of referring expressions was overall consistent with the suggested models of reference control within the topicality and saliency paradigm for both spoken and signed languages. However, native acquisition increased the use of null marking, with no significant effect in nominal production. We also found that native signers used mixed perspectives in narrative discourse more than late signers. Given these findings, we suggest that language deprivation might negatively affect late signers' pragmatic competence (e.g., sensitivity to economy of form) and their use of complex morphosyntactic structures in reference tracking, which are known to be age-sensitive.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Narratives are extensive discourses that construct an important part of human communication. In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities. Certain forms such as nominal expressions are commonly used for the first mentions of a referent in a discourse, and other forms such as pronominals and null marking may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse. The selection of these referring expressions and anaphoric tools (henceforth REATs) is contingent on cognitive abilities, such that the encoder must rely on the activation of the working memory to keep in mind previously referred expressions and long strings of information (Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play a pivotal role in the tracking of referents: The form with the least amount of redundant information must be selected based on the accessibility levels of the referent (Ahn, 2019; Ariel, 1990). This denotes that acquiring adult-like narrative competence is not a trivial job for children. Speakers and signers of a native language start to refer to entities and use complex anaphoric or cataphoric expressions appropriately following age 10 (Aksu-Koç \& Nicolopoulou, 2015; Hickmann, Hendriks, Roland, \& Liang, 1996; Rathmann, Mann, \& Morgan, 2007). Similarly, second language learners find it difficult to produce a coherent and structured narrative where it is important to use REATs appropriately in line with the cognitive statuses of the referents. This finding comes both from research examining L2 acquisition of spoken narratives (Gundel, Hedberg, \& Zacharski, 1993; Hickmann, Hendriks, Roland, \& Liang, 1996; Williams, 1988) and signed narratives (Bel, Ortells, \& Morgan, 2015; Frederiksen \& Mayberry, 2019) which are conveyed in different communication modalities. Yet, sign languages and deaf populations have remained mostly understudied with respect to the acquisition of narrative and pragmatic competence.

Sign languages are the natural languages of the visual-spatial modality prominently used among deaf populations. The cognitive and linguistic development of native signers (i.e., those who start acquiring a signed language from one or more deaf caregivers following birth) parallels that of native speakers (Chamberlain, Morford, \& Mayberry, 1999; Petitto, 1997). However, reports indicate that at least over 90\% of all deaf individuals are born into hearing families who do not know how to sign (Weaver \& Starner, 2011; Woll, 2013). This means that the auditory modality used in the household is not accessible to many deaf children. That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school. Although the lack of a systematic and comprehensible form of communication in the early years of life is compensated with an internally consistent primitive gesture-based system called home sign (Goldin-Meadow, 2005), this neither fully resembles natural languages nor reinstates first language acquisition. Consequently, language deprivation has been found to have negative effects on certain cognitive and linguistic measures among late signing children. To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic abilities (Boudreault \& Mayberry, 2006; Karadöller, Sümer, \& Özyürek, 2017; Newport, 1990), phonological processing and phonemic awareness (Corina, Hafer, \& Welch, 2014; MacSweeney, Waters, Brammer, Woll, \& Goswami, 2008), vocabulary knowledge (Anderson, 2005; Lu, Jones, \& Morgan, 2016), and narrative abilities (Becker, 2009).

Given that there is a paucity of studies that examine deaf adult signers with delayed exposure to a sign language, it is not very clear whether the narrative abilities of late signers differ from those of native signers. To expand on the previous literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in Turkish Sign Language (TİD) narratives through a story-telling paradigm. We explore the nature of REATs selection and perspective taking as a possible indicator of pragmatic competence between our acquisition groups. The paper is organized as follows: Section 1 (i) gives an overview of REATs selection and perspective taking in signed languages, (ii) describes hitherto attested REATs in TİD, and (iii) reviews previous findings on the effect of delayed first language acquisition on signers' narrative competence. Section 2 and 3 present the methodology and results of this paper, respectively. Finally, Section 4 provides a general discussion of the reference tracking strategies among native and late signers in relation to the effects of delayed first language acquisition on pragmatic competence.

\hypertarget{reference-tracking-strategies-in-natural-languages}{%
\subsection{Reference Tracking Strategies in Natural Languages}\label{reference-tracking-strategies-in-natural-languages}}

\hypertarget{reats-selection}{%
\subsubsection{REATs Selection}\label{reats-selection}}

Natural languages can be systematically used to keep track of previous referents in the discourse and mark the changing cognitive statuses of the REATs to create a more coherent narrative. Among the theoretical frameworks that aim to explain that different linguistic forms are selected in accordance with the addressee's mental representations of the referents in the discourse is the Givenness Hierarchy (Gundel, Hedberg, \& Zacharski, 1993). According to this model, given or familiar information is expressed in more attenuated forms such as pronominal use and null marking since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information (Chafe, 1976). The Givenness Hierarchy is akin to other previously proposed models of reference control within the topicality and saliency paradigm such as the Accessibility Hierarchy (Ariel, 1990), Topic Continuity (Givon, 1983), and Familiarity Scale (Prince, 1981). Gundel, Hedberg, and Zacharski (1993) argued that the distribution of forms is not random and abides by the Gricean Maxim of Quantity (Grice, 1975) in that the encoder must make the most informative and succinct choice possible. On this ground, for highly accessible referents, linguistic forms with the least phonetic content are selected for the least amount of processing effort based on the economy principle.

It is also possible to categorize these referring expressions into three according to their discourse statuses (Gullberg, 2006). A referent can be introduced, maintained, that is, continued across two or more clauses, or reintroduced into the discourse following a topic shift. Based on the principles of reference accessibility and relevance, the introduction and reintroduction of a referent are often accomplished by full noun phrases which can be commonly modified by an adjective or demonstrative in spoken languages and also by optional markers such as constructed action, classifier constructions, or index points in sign languages. Maintenance or topic continuity, on the other hand, is typically marked with null marking, pronominal forms, and classifiers across signed languages since the continued referent is still highly accessible by the addressee.

\hypertarget{perspective-taking-in-signed-narratives}{%
\subsubsection{Perspective taking in signed narratives}\label{perspective-taking-in-signed-narratives}}

Sign language research has identified a split, albeit using different terminology, between what is commonly referred to as the character perspective and the observer perspective to explain some of the attested structures in signed narratives. The character perspective (also referred to as the large-scale use of the space) depends on the signer's mental conceptualization of the narrated event. In other words, the signers engage with the space in front of themselves as if they are interacting with the objects and characters in the story. This is often achieved with the signer taking on the role of a referent in the narrative and performing the actions by employing a linguistic device called constructed action (Metzger, 1995) or referential shift (Emmorey, Corina, \& Bellugi, 1995) that are marked by certain non-manual cues such as the facial expression and the breaking of the eye contact. This is to say, the narrator's head, face, torso, arms, and as an optional marker, hands become one with those of the associated referent(s) (Smith \& Cormier, 2014). Constructed action does not necessarily entail a body shift to a locus in the signing space. Instead, the signer mentally rotates the space and faces the addressee in most cases.In the observer perspective or the small-scale use of the signing space, referents are fully or partially represented not by the signer's body but by the manually articulated classifier, verbal or nominal constructions signed in front of the body.

Signers typically switch between these perspectives multiple times while narrating. While the observer perspective is commonly used to describe an intransitive event and to show change of location (e.g., path movement), the character perspective is used to inform the addressee about transitive events and to depict what the constructed character is doing. Switching between perspectives can be done sequentially: Signers can talk about different characters in a narrative by providing locative information and describe intransitive events from the observer perspective through different verbal and classifier constructions. The character perspective can immediately follow an observer perspective representation and describe the transitive or manual activities carried out by the character(s).

Capitalizing on the visual-spatial modality, signers can also engage in simultaneous mixing of perspectives. Morgan (2005) noted that these two perspectives, shifted referential space versus fixed referential space in his terminology, can be mixed for the simultaneous articulation of multiple referents. In this representation, while the signer constructs the actions of a referent via constructed action, the hands or part of the signer's body can be partitioned (see Dudis, 2004) and leave the mentally rotated space. The partitioned limbs (e.g., selected fingers) can then function as a whole entity classifier for another referent. The signer's construction or enactment of the animate referent can interact with the objects situated in the observer perspective through body part classifiers or verbal constructions. This way, the signer provides spatial information from both the observer and character perspective. This is to ensure the achievement of spatial coherence and efficiency of information. Simultaneous encoding of spatial information is then motivated by the economy principle of language.

\hypertarget{tracking-referents-in-turkish-sign-language-tid}{%
\subsection{Tracking Referents in Turkish Sign Language (TİD)}\label{tracking-referents-in-turkish-sign-language-tid}}

The linguistic inventory for discourse referents in TİD includes nominal constructions (bare nouns, fingerspelled nouns, and modified nouns) for low accessibility referents. Null marking (e.g., plain verb, agreement verb, and constructed action), classifier predicates (e.g., whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX) are selected for high accessibility referents. This interim section describes the REAT types attested in TİD under four main headings.

\begin{multicols}{2}
\begin{exe}
\ex
\begin{xlist}
\ex Full noun \\\glll
\includegraphics[width=2.5cm]{pictures/1a_1.png} \includegraphics[width=2.5cm]{pictures/1a_2.png} \\ \textsc{mouse} \textsc{cat} \\ \\ 
\ex Noun modified by adjective \\\glll
\includegraphics[width=2.5cm]{pictures/1c_1.png} \includegraphics[width=2.5cm]{pictures/1c_2.png} \\ \textsc{mouse} \textsc{sick} \\ \\ 
\ex Fingerspelled noun\\\glll 
\includegraphics[width=2.5cm]{pictures/1b_1.png} \includegraphics[width=2.5cm]{pictures/1b_2.png} \includegraphics[width=2.5cm]{pictures/1b_3.png}\\ \textsc{t-} \textsc{-o-} \textsc{-m} \\ \\
\ex Noun modified by index\\\glll 
\includegraphics[width=2.5cm]{pictures/1d_1.png} \includegraphics[width=2.5cm]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
\end{xlist}
\end{exe}
\end{multicols}
\vspace{-1.25cm}
\newpage
\ea 
\ea Pronominal \\\glll
\includegraphics[width=2.5cm]{pictures/2_1.png} \includegraphics[width=2.5cm]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
\z \z
\vspace{-0.75cm}
\begin{multicols}{4}
\begin{exe}
\ex
\begin{xlist}
\ex Whole entity \\\glll 
\includegraphics[width=2.25cm]{pictures/3a.png} \\ {\textsc{dh: mouse}} \\ {\textsc{ndh: tunnel}}\\ \\
\ex Body part \\\glll
\includegraphics[width=2.25cm]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
\ex Handling \\\glll 
\includegraphics[width=2.25cm]{pictures/3c.png} \\ \textsc{hold} \\ \\
\ex Extension \\\glll 
\includegraphics[width=2.25cm]{pictures/3d_1.png} \includegraphics[width=2.25cm]{pictures/3d_2.png}\\ \textsc{wood} \textsc{plank}\\ \\
\end{xlist}
\end{exe}
\end{multicols}

\begin{multicols}{3}
\begin{exe}
\ex
\begin{xlist}
\ex Constructed action \\\glll 
\includegraphics[width=2.5cm]{pictures/4c.png} \\ \textsc{struggle} \\ \\

\ex Agreement verb  \\\glll 
\includegraphics[width=2.5cm]{pictures/4b.png} \\ {\textsc{$_{3a}$kiss$_{3b}$}} \\ \\

\ex Plain verb  \\\glll
\includegraphics[width=2.5cm]{pictures/4a_1.png} \includegraphics[width=2.5cm]{pictures/4a_2.png}  \\ \textsc{restroom} \textsc{enter}\\ \\
\end{xlist}
\end{exe}
\end{multicols}

\hypertarget{nominal}{%
\subsubsection{\texorpdfstring{Nominal (\textsc{nom})}{Nominal ()}}\label{nominal}}

In TİD, it is possible to use bare noun phrases as in (1a) and fingerspelled nouns as in (1b) to introduce or reintroduce referents into the discourse. The first mention of a referent can also be tracked with noun phrases modified by adjectives as in (1c) which either follow or precede a nominal (Özsoy \& Nuhbalaoğlu, 2014). Alternatively, the noun phrase can retain its basic meaning and the modifier can help determine its position on the accessibility hierarchy by expressing the definiteness of the nominal. Definiteness in TİD can be expressed with a prenominal or postnominal \textsc{ix} functioning as a determiner or a demonstrative as in (1d).

\hypertarget{pronominal}{%
\subsubsection{\texorpdfstring{Pronominal \textsc{ix} (\textsc{pro})}{Pronominal  ()}}\label{pronominal}}

As discussed previously, the \textsc{ix} sign (extended index finger) has various functions in signed languages, one of which is the \textsc{pro} use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when reintroducing a topic in discourse (Emmorey, 1996). That is to say, pronominal \textsc{ix} is licensed only if the referent is physically present in the space or the non-present antecedent is already associated with a locus. Once such a spatial relation is set up by the signer, it is possible to bring those referents back into the discourse by pointing to their location. The same \textsc{ix} function has also been identified for TİD (Sevinç, 2006). The example in (2) illustrates a spatially-anchored pronominal reference for topic reintroduction. These signs are thus rarely used for the first mentions of an entity (Perniss \& Özyürek, 2015). For the maintenance condition where the referent is highly accessible, some studies reported a relatively few instances of pronominal \textsc{ix} as signers favored other grammatical tools for reference such as null marking or classifier constructions (Cormier, Smith, \& Zwets, 2013; Frederiksen \& Mayberry, 2016, 2019). However, recent work pointed out what is referred to as the referentially underspecified pronominal \textsc{ix} which does not require the non-present antecedent to be spatially anchored to a previously established locus. Although other interpretations consider pointing signs without an noun phrases as examples of demonstratives (Koulidobrova \& Lillo-Martin, 2016), for the purposes of the present study, we will refer to those constructions as pronominal \textsc{ix} and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.

\hypertarget{classifier}{%
\subsubsection{\texorpdfstring{Classifier (\textsc{cl})}{Classifier ()}}\label{classifier}}

\textsc{cl} predicates are highly complex morphological constructions which are very common in sign languages (Supalla, 1986). These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape (Talmy, 2003). As in most sign languages hitherto studied, \textsc{cl} constructions are also common and found in TİD (see Dikyuva, Makaroğlu, \& Arık, 2015 for a review). Following the classification of Engberg-Pedersen (1993), we distinguish between four types of \textsc{ix} constructions. Whole entity classifiers (\textsc{wecl}) represent the entirety of any referent whereas body part classifiers (\textsc{bpcl}) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth). Handling classifiers (\textsc{hcl}) occur with transitive verbs and simply depict how an object is handled or manipulated by the agent. On this ground, \textsc{hcl} depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space. Extension classifiers (\textsc{extcl}), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape. We exemplify these constructions in (3). As discussed before, \textsc{hcl} is transitive in nature and depicts the handling action performed by an animate agent. This is why most research tend to consider \textsc{hcl} as an extension to constructed action and group them with the character perspective where the signer's body aligns with the referent (Cormier, Smith, \& Zwets, 2013; Dudis, 2004; Özyürek \& Perniss, 2010). However, the other \textsc{ix} constructions co-occur with the observer perspective. That is to say, the signer does not mentally rotate the space nor employ constructed action but instead narrates the event from the observer perspective by fully or partly stating the referent via different linguistic forms. As expected, these constructions are argued to be higher on the accessibility hierarchy than \textsc{nom} (\textsc{wecl} more so than \textsc{bpcl} due to its less transparent and iconic nature). This is because they take on certain semantic properties of the referent, but they are reported to rank lower than \textsc{pro} pointing signs which are less explicit in form (Morgan, 2005). This contrasts with the finding that \textsc{pro} may be dispreferred in favor of \textsc{cl} constructions (Cormier, Smith, \& Zwets, 2013; Frederiksen \& Mayberry, 2016) contra Bel, Ortells, and Morgan (2015); Morgan (1999) who observed an abundance of \textsc{pro} within their participants' signed discourse. Considering the Givenness Hierarchy (Givon, 1983) that suggests more attenuated content for highly accessible referents, we expect \textsc{pro} to be higher on the hierarchy than \textsc{cl} in the present article. Nevertheless, \textsc{cl} constructions are commonly used by signers to maintain previous referring expressions although some studies reported a sporadic use of \textsc{cl} for referent introduction and reintroduction (e.g., Morgan, 2005).

\hypertarget{null-marking}{%
\subsubsection{Null Marking}\label{null-marking}}

Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null marking (\textsc{NULL}) when referring to highly accessible referents which are still active in the mental representation of the addressee. In signed discourse, this is achieved by employing verbal constructions when the event is narrated via the observer perspective. To identify the predicate types in ASL, Padden (1986) suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs. According to this classification, plain verbs (\textsc{plain}) (4a) do not agree with the subject or the object. This means that they are not inflected for number or person. This contrasts with agreeing verbs (\textsc{agr}) (4b) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put3a} `I placed the book there') take locational affixes but like plain verbs, they are not inflected for number or person. TİD too licenses \textsc{null} arguments. Following this classification, Sevinç (2006) argued that although it is common in TİD to drop arguments, there were certain limitations on the \textsc{null} marking of the arguments for \textsc{agr} and \textsc{plain} verbs. It is also possible and common for signers to use constructed action (\textsc{ca}) as in (4c) and take on the role of a character in the narrative while omitting arguments. This means that they do not assign locations to previously established referents but instead designate their own body as the specified locus. This change in the spatial scale is often marked with non-manual markers, with a break in eye gaze and the facial expression being the most informative cues (Engberg-Pedersen, 1993).

\hypertarget{mixing-of-perspectives}{%
\subsubsection{\texorpdfstring{\textbf{Mixing of perspectives}}{Mixing of perspectives}}\label{mixing-of-perspectives}}

As described in previous sections, it is also possible to mix both perspectives to bring two or more referents simultaneously into the discourse (Dudis, 2004; Janzen, 2004; Liddell, 2003). The fusion of character perspective and observer perspective is attested in TİD (Perniss \& Özyürek, 2015). In this mix, the signer's face and torso are mapped onto the narrated character with \textsc{ca} and the signer's hands or arms are partitioned off from the character perspective to sign verbal or \textsc{cl} predicates. These manual constructions are viewed externally by the signer from the observer perspective. In the example (4a), the signer takes on the role of Tom, an animate character in the narrative, holding a gun. The signer's right-hand leaves the mentally rotated space and starts to depict the gun being bent over toward Tom (the signer's head) by constructing an agreeing whole entity classifier. Two referents with varying degrees of cognitive statuses are then tracked simultaneously via two different REATs. Likewise, the signer in (4b) first introduces the female cat in the narrative by lexically identifying it as the syntactic object. The agentive subject (Tom), however, is highly accessible to the addressee. For this reason, the signer omits the subject argument and only signs the predicate \textsc{see} which agrees with unspecified loci. The second predicate \textsc{KISS} is represented with \textsc{bpcl} the endpoint of which this time aligns with the location of the signer who takes on the role of the previously established referent by using the character perspective. In the following clause, the event shifts from the perspective of the female cat to Tom. Here, the signer depicts Tom's hiding behind a couch and his left hand is partitioned off to represent a third referent (the black cat) with a \textsc{WECL}. In the second mix, Tom secretly watches the black cat running around. By employing these morphologically complex constructions that result in the mixture or blend of the two perspectives, signers can make the most of the visual modality for the simultaneous tracking of multiple referents with varying accessibility levels. Prioritizing informativeness and efficiency of information, this observation is in line with the presuppositions of the Gricean Maxim of Quantity (Grice, 1975) and the Relevance Theory (Sperber \& Wilson, 1986).

\newpage
\begin{exe}
\ex 
\begin{xlist}
\ex Perspective mixing with CA and WECL \\\glll
\includegraphics[width=2.5cm]{pictures/5a_1.png} \includegraphics[width=2.5cm]{pictures/5a_2.png} \includegraphics[width=2.5cm]{pictures/5a_3.png}\\ {\textsc{cp: hold.gun}} \\ \textsc{op:} {\textsc{ }} {\textsc{$_{3a}$gun.bend$_1$}} \\
\glt `The gun is bending toward Tom.'
\ex Perspective mixing across two sentences \\\glll 
\includegraphics[width=2.5cm]{pictures/5b_1.png} \includegraphics[width=2.5cm]{pictures/5b_3.png} \includegraphics[width=2.5cm]{pictures/5b_2.png} \includegraphics[width=2.5cm]{pictures/5b_4.png}\\ \textsc{cp:} {\textsc{}} {\textsc{being.kissed}} {\textsc{hide}} \\ {\textsc{op: girl}} {\textsc{$_{3a}$see$_{3b}$}} {\textsc{$_{3b}$kiss$_1$}} {\textsc{$_{3a}$run$_{3b}$}} \\
\glt `Tom sees and kisses the female cat. Then the gray cat comes running while Tom hides behind the couch.'
\end{xlist}
\end{exe}

\hypertarget{learning-how-to-track-referents-in-signed-languages}{%
\subsection{Learning How to Track Referents in Signed Languages}\label{learning-how-to-track-referents-in-signed-languages}}

Considering the work on sign language discourse that indicates native sign language development of deaf children parallels that of hearing children, all else being equal, it is not surprising that the pragmatic competence is also achieved similarly, albeit with different milestones (Morgan, 2000). Deaf children of deaf parents have been observed to lag behind hearing age-matched peers at 4 years of age in terms of the organization of a narrative, but the general understanding is that they catch up within two years (Vercaingne-Ménard, Godard, \& Labelle, 2001). In both communication modalities, children start to produce more complex and coherent stories with appropriate and informative linguistic forms to indicate anaphoric or cataphoric references after age 10 (Aksu-Koç \& Nicolopoulou, 2015; Hickmann, Hendriks, Roland, \& Liang, 1996; Rathmann, Mann, \& Morgan, 2007). Prior to this age, both signing and speaking children fail to refer to entities consistently since their pragmatic abilities are yet to develop. The manipulation of linguistic forms for discourse functions to signal topic continuance or shift is difficult to master until the beginning of formal education. These findings are also comparable across different sign languages including British Sign Language (BSL) (Morgan, 2005), American Sign Language (ASL) (Loew, 1984), German Sign Language (Deutsche Gebärdensprache - DGS) (Becker, 2009), and Canadian Sign Language (LSQ) (Vercaingne-Ménard, Godard, \& Labelle, 2001), pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.

Given the preliminary evidence that topic continuity and salience marking is also difficult for second language learners regardless of modality Hickmann \& Hendriks (1999), it has been suggested under the Interface Hypothesis that the syntax-discourse interface is vulnerable and especially challenging to acquire (Sorace, 2011). For this reason, it is thus one of the latest components of language to be learned. Similarly, Sorace and Serratrice (2009) discuss that this interface entails much more than morphosyntactic knowledge but also includes cognitive and pragmatic resources that help provide discourse coherence. They propose that the second language discourse is over-redundant. This means that second language learners produce more overt pronouns or fuller forms of language for highly accessible referents possibly as a compensatory strategy in an attempt to reduce the increasing load on cognitive processing.

Previously cited studies indicated that acquisition of reference tracking and thereby the mastery of perspectives in discourse production is challenging for deaf children of deaf parents (i.e., native signers) and even more so for second language learners. Still, it is unclear whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking. In response to this question, Becker (2009) examined DGS signing deaf native and late children's narrative abilities. She reported that late signing children generated more explicit forms such as \textsc{nom} when referring to the characters in the story compared to native learners. Her findings also revealed that late learners failed to establish unambiguous references and hence were less successful in the use of the signing space. Becker (2009) concluded with the claim that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax. A similar line of research came from Smith and Cormier (2014). Likewise, they explored the use of the two perspectives employed with or without \textsc{ca} among BSL signing deaf native and late children. The results indicated that only native signers employed a non-prototypical alignment of perspectives to increase efficiency. In the study, the native signers used the character perspective without \textsc{ca}, with the opposite pattern observed for late learners. The authors reasoned that the former had been a more complex instance of reference tracking and the difference might stem from the fact that late signers have hearing parents and thus engage in less joint attention which translates into fewer opportunities to learn appropriate eye gaze patterns. Comparably, Cormier, Smith, and Zwets (2013) investigated the use of constructed action in BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity. In other words, they produced more explicit forms like \textsc{nom}, and fewer instances of \textsc{null} when maintaining previous referents while Ferrara and Johnston (2012) failed to find such a preference for less phonetic content for highly accessible referents by native signers. Most relevantly to the present study, Gür (2018) replicated some of these findings for TİD in her work on the age of sign language acquisition effects on narrative abilities. She found that signers with early acquisition engaged in more correct instances of role shift while narrating. Given the limited amount of research involving late adult signers, the effects of delayed exposure to language on later discourse competence have been understudied. As discussed in the first section, the morphosyntactic competence of late signers falls behind that of native learners (e.g., Boudreault \& Mayberry, 2006; Newport, 1990) and these differences persist into adulthood. As a result, this discrepancy may continue to have an inhibiting effect on the syntax-discourse interface late in adulthood.

\hypertarget{present-study}{%
\subsection{Present Study}\label{present-study}}

In an attempt at answering this question, the present study aims to examine the reference tracking strategies among native and late deaf adult signers of TİD in relation to REATs selection and perspective taking. Based on previous research, we hypothesize that late signers will become more over-redundant in their reference tracking compared to native signers. This would translate to a preference by late signers for \textsc{nom} over \textsc{null} especially in the maintenance and re-introduction settings. In contrast, we would expect that native signers use more \textsc{null} markers in their tracking based on the economy principle. As for the perspective taking, we hypothesize that late learners will employ fewer instances of perspective mixing to refer to multiple references because of less developed morphosyntactic and pragmatic abilities in comparison to native signers. In contrast, we predict that native signers will use the signing space more economically and use more zero anaphora along with better use of the spatial scale.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We tested 29 Deaf adult signers, 15 with exposure to TİD from birth (i.e., native signers), 14 with exposure between ages 3 and 17 (i.e., late signers) (MAOA = 7.7 years, SD = 3.3 years). 26 participants were pre-lingually deaf, that is they were either born deaf or they had acquired deafness before the age of 3. The remaining 3 participants became deaf between the ages of 4-7. Participants' level of education (LOE) (MLOE = 12.3 years, SD = 2.0 years) and the average year of language use following the start of exposure to TİD did not differ much between the acquisition groups (see Appendix for more detailed demographic information). All participating native signers were born into deaf families and started acquiring TİD from at least one of their caregivers after birth. The parents of 14 participating native signers were deaf. One native signer had a deaf father and a hearing mother. Comparably, late learners were all born into hearing families and started learning TİD later in life after age 3, usually with the start of formal education. The age of acquisition for late signers ranged between years 3-17 (MAOA = 8.70 years, SD = 3.52 years). The parents of all participating late signers were hearing, and the reported home languages spoken in childhood consisted mainly of Turkish and Home Sign. All participants indicated that their preferred language of communication was TİD which they used for most contexts. On a 1 (poor) to 5 (proficient) point scale, participants' self-reported proficiency in TİD was on average 4.97 points.

\hypertarget{task-procedure}{%
\subsection{\texorpdfstring{\textbf{Task Procedure}}{Task Procedure}}\label{task-procedure}}

We used a story-telling paradigm to examine how native and late signers of TİD track references in signed discourse. Participants were presented with 10 short extracts from a cartoon and asked to retell them. The stimulus stories were extracted from the cartoon Tom and Jerry. The duration of clips ranged from 17 seconds to 28 seconds. Each clip had six to ten events which consisted of one or multiple (up to three) recurring animate characters for which the present context required the introduction, maintenance, and re-introduction of referents.

Participants sat in front of a laptop computer which had the stimulus stories. The instructions were all given by a deaf research assistant who was present in all sessions and ensured that the task requirements were clear. The informed consent form was translated into TİD and explained to the participants by the same assistant before data recording. Participants were recorded by a digital camera, and they narrated the content of each clip to an additional experimenter who sat across the participants. The order of clips was randomized for each participant.

\hypertarget{coding}{%
\subsection{Coding}\label{coding}}

We coded the discourse status (Introduction, Maintenance, Re-introduction), REAT type (Nominal, Pronominal, Classifier, Null), and perspectives (Character, Observer, Mixed) (see Table 2 for the observed instances), identifying 4523 reference markings in total. We used ELAN (Sloetjes \& Wittenburg, 2008) for data annotation. First Author annotated all the narratives. We first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod). On one tier, we had the grammatical tools for a single reference with the subcategories Nominal, Pronominal, Zero Anaphora, and Classifier. Following the classification of Frederiksen and Mayberry (2019), we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence. Only subjects were coded for maintenance and re-introduction. We coded subject referents as maintained only if they had been signed in the previous clause. To this end, the referents which had been discontinued across one or more sentences were all regarded reintroduced into the discourse. We classified all handling classifiers which depicted the action of a referent under constructed action in line with previous interpretations (Cormier, Smith, \& Zwets, 2013).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{reats-selection-1}{%
\subsection{REATs Selection}\label{reats-selection-1}}

We collected 4168 data points as a result of the experiment. Each data point is a referent tracked by the participant. Each of the referents is classified in terms of its discourse function (Introduction, Maintenance, Re-introduction), its grammatical function (Classifier, Nominal, Pronominal, Null), and the group of the participant (Native, Late). Figure \ref{fig:count-plot} shows the numerical distribution of the responses according to the aforementioned variables. The use of nominal or null is more numerous than the classifier and pronominal reference tracking. There is a noticeable difference between participant groups in the use of null marking for maintaining a referent. Native signers of TİD have used it more compared to the late signers.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/count-plot-1.pdf}
\caption{\label{fig:count-plot}Referent counts by Discourse function (Introduction, Maintanence, Re-introduction), Grammatical function (Classifier, Nominal, Pronominal, Zero anaphora), and Participant group (Native, Late)}
\end{figure}

For more inference on the data and have pairwise comparisons, we fit a regression model to the grammatical function of the responses using Discourse Status and Group as predictors. We used the brms package (Bürkner, 2018) in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late). We did not define priors ourselves but used the preset uninformed priors. Figure \ref{fig:model-gram} shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50\% credible intervals, and the thin line is the 95\% credible intervals.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/model-gram-1.pdf}
\caption{\label{fig:model-gram}Regression model plot for the Grammatical Function (Nominal, Pronominal, Zero) with the predictors Discourse function (Introduction, Maintenance, Re-introduction) and Participant group (Native, Late)}
\end{figure}

In a categorical distribution, the model takes one of the responses as a reference and compares all other responses to it. For example the first coefficient, nominal intercept, in Figure \ref{fig:model-gram} is the comparison of Nominal referent use versus a classifier referent use. It means that participants produced more nominal referents than classifier referents in general regardless of participant group and discourse function. The second coefficient shows that zero anaphora referents are also used more than classifier referents by the participants in general, but not as much as nominal referents.

In the comparison of nominal versus classifier referents, being a native acquirer of TiD did not have an effect. Nominal referents were used more in the introduction setting of a discourse compared to a re-introduction setting, again the whole comparison is being made against a classifier referent. In contrast to the introduction setting, nominal referents were used less in the maintain discourse setting compared to the re-introduction setting, within the greater comparison to classifier referents. There does not seem to be an interaction between introduction setting of discourse and being a native acquirer of TiD, however there looks like an interaction between the Maintain setting of the discourse with being a native acquirer of TiD. In that, native signers produced less nominal referents in the maintain setting of the discourse compared to the other combinations.

In zero anaphora referents compared to classifier referents, being a native acquirer of TiD increased the odds of using zero anaphora in general. Zero anaphora were used less as introduction setting of discourse compared to the re-introduction setting. This contrasts with the nominal referent (4th coefficient) as it was used more in the introduction setting versus re-introduction. In contrast to the introduction setting of the discourse, zero anaphora referents were used more as maintain setting of discourse compared to the re-introduction setting, yet again within the greater comparison to the classifier referents. There seems to be an interaction between the introduction setting of discourse and being a native acquirer of TiD. In that, the odds of using zero anaphora referent further increases if the participant is a native acquirer of TiD and the referent is used in the introduction setting. There is another interaction with regards to the maintain discourse setting, contrasting the first interaction. In that, the odds of using zero anaphora decrease if the participant is a native acquirer of TiD and the referent is used in the maintain setting of discourse.

We infer the following from the results of our first model. Participants used nominal referents and zero anaphora most of the time compared to classifier referents. Being a native acquirer of TiD increases the overall chances of using zero anaphora referents. However, this effect interacts with the discourse setting. The chances of a native acquirer of TiD using zero anaphora referent increases if they are to use it as a maintain function of the discourse. This interaction is reversed in the case of nominal referents. The chances of a native acquirer of TiD using nominal referents decrease if they are to use it as a maintaion function of the discourse. Furthermore, the chances of using zero anaphora referents decreases if the native participants are to use it as the introduction function of the discourse.

In addition to the discourse and grammatical function of a referent, we also coded the perspective of the participant in their response. In the responses where the participants simulated the story in their summary, they could have: (i) acted as the character, (ii) mixed character and observer perspectives, or (iii) been a complete observer in their reference tracking. Figure \ref{fig:plot-perspective} shows the distribution of the responses by the Discourse function, Perspective (Character, Mixed, Observer), and the participant group.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/plot-perspective-1.pdf}
\caption{\label{fig:plot-perspective}Distribution of the simulated reference tracking by Discourse function (Introduction, Maintenance, Re-introduction), Perspective (Character, Mixed, Observer), and the Participant group (Native, Late).}
\end{figure}

For more inference on the data and have pairwise comparisons, we fit a regression model to the perspective of the responses using Discourse function and the participant group as predictors. We used the brms package in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse (Intro, Maintenance, Re-intro) and Participant group (Native, Late). We did not define priors and used the preset uninformed priors. Figure \ref{fig:model-gram} shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50\% credible intervals, and the thin line is the 95\% credible intervals.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/model-pers-1.pdf}
\caption{\label{fig:model-pers}Regression model plot for the Perspective of the reference with the predictors Participant group (Native, Late) and Discourse function (Introduction, Maintain, Re-introduction)}
\end{figure}

The results of this model too are interpreted with a reference level of the response in mind, here the Observer perspective is the reference level with regards to which we make the comparisons. Participants in general used less mixed perspective in their referents compared to the observer perspective (1st coefficient), they also used less character perspective compared to the observer perspective. The magnitude of the decrease for Mixed versus Observer is greater than that of Character versus Observer. This means that participants used more Observer and Character perspectives than Mixed perspectives. The posterior probability distributions for both intercepts are left skewed, this shows that there is higher variance towards the lower end of the estimate. This means that there are extreme cases where Mixed and Character perspectives were used even less than the Observer perspective.

Being a native acquirer of TiD increased the chances of using mixed perspective compared to using the observer perspective. Mixed perspective was used less in the introduction setting of the discourse compared to the re-introduction setting. In contrast, mixed perspective was used more in the maintain setting of the discourse compared to the re-introduction setting. This means that mixed perspective was used more as introduction and re-introduction settings than the maintain setting. There does not seem to be an interaction between being a native acquirer of TiD and the discourse function that effects the chances of using mixed perspective compared to the observer perspective.

Being a native acuqirer of TiD increased the chances of using character perspective compared to suing the observer perspective. In a similar spirit to the effect of introduction setting in mixed versus observer (4th coefficient), character observer was used less as the introduction setting of the discourse compared to the re-introduction setting. In contrast, character perspective was used more in the maintain setting of the discourse compared to the re-introduction setting. This means that character perspective was used more as introduction and re-introduction settings than the maintain setting. There seems to be interactions of being a native acquirer of TiD and discourse settings this time. The chances of character perspective further increase if the participant is a native acquirer of TiD and they are to use the referent in the introduction setting of the discourse. The chances of character perspective decrease if the participant is a native acquirer of TiD and they are to use the referent in the maintain setting of the discourse.

We infer the following from the results of our second model. Participants used Observer and Character perspective more than Mixed perspective in general. Being a native acquirer of TiD increases the chances of using Mixed or Character perspective compared to the observer perspective. The discourse setting introduction decreases the chances of using Character or Mixed perspective, but this is partly undone for the character perspective by this setting's interaction with being a native acquirer of TiD.

In general, the posterior probability distributions of our second model's coefficients are left or right skewed. This means that there was high variance towards the extreme in a way that our data or experiment setting was not enough to prune them out. We are noting this difference of the first and second models to show that, although we treat posterior probability distributions as categorical statements of an effect, not all interpretations of the estimates are the same.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-ahn2019}{}%
Ahn, D. (2019). \emph{The determinacy scale: A competition mechanism for anaphoric expressions} (PhD thesis).

\leavevmode\hypertarget{ref-aksu-kouxe72015}{}%
Aksu-Koç, A., \& Nicolopoulou, A. (2015). Character reference in young children's narratives: A crosslinguistic comparison of English, Greek, and Turkish. \emph{Lingua}, \emph{155}, 62--84. \url{https://doi.org/10.1016/j.lingua.2014.04.006}

\leavevmode\hypertarget{ref-anderson2005}{}%
Anderson, D. (2005). \emph{Lexical Development of Deaf Children Acquiring Signed Languages}. Oxford University Press. Retrieved from \url{https://www.oxfordscholarship.com/view/10.1093/acprof:oso/9780195180947.001.0001/acprof-9780195180947-chapter-6}

\leavevmode\hypertarget{ref-ariel1990}{}%
Ariel, M. (1990). \emph{Accessing Noun-Phrase Antecedents}. Routledge.

\leavevmode\hypertarget{ref-bamberg1997}{}%
Bamberg, M. G. (1997). Positioning between structure and performance. \emph{Journal of Narrative and Life History}, \emph{7}(1-4), 335342.

\leavevmode\hypertarget{ref-becker2009}{}%
Becker, C. (2009). Narrative competences of Deaf children in German Sign Language. \emph{Sign Language \& Linguistics}, \emph{12}(2), 113--160. \url{https://doi.org/10.1075/sll.12.2.02bec}

\leavevmode\hypertarget{ref-bel2015}{}%
Bel, A., Ortells, M., \& Morgan, G. (2015). Reference control in the narratives of adult sign language learners. \emph{International Journal of Bilingualism}, \emph{19}(5), 608--624. \url{https://doi.org/10.1177/1367006914527186}

\leavevmode\hypertarget{ref-boudreault2006}{}%
Boudreault, P., \& Mayberry, R. I. (2006). Grammatical processing in american sign language: Age of first-language acquisition effects in relation to syntactic structure. \emph{Language and Cognitive Processes}, \emph{21}(5), 608--635. \url{https://doi.org/10.1080/01690960500139363}

\leavevmode\hypertarget{ref-buxfcrkner2018}{}%
Bürkner, P.-C. (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. \emph{The R Journal}, \emph{10}(1), 395. \url{https://doi.org/10.32614/RJ-2018-017}

\leavevmode\hypertarget{ref-chafe1976}{}%
Chafe, W. (1976). \emph{Givenness, contrastiveness, definiteness, subjects, topics, and point of view}. Retrieved from \href{https:///paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}{/paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}

\leavevmode\hypertarget{ref-chamberlain_language_1999}{}%
Chamberlain, C., Morford, J. P., \& Mayberry, R. I. (Eds.). (1999). \emph{Language acquisition by eye}. Mahwah, N.J: Psychology Press.

\leavevmode\hypertarget{ref-corina2014}{}%
Corina, D. P., Hafer, S., \& Welch, K. (2014). Phonological Awareness for American Sign Language. \emph{The Journal of Deaf Studies and Deaf Education}, \emph{19}(4), 530--545. \url{https://doi.org/10.1093/deafed/enu023}

\leavevmode\hypertarget{ref-cormier2013}{}%
Cormier, K., Smith, S., \& Zwets, M. (2013). Framing constructed action in British Sign Language narratives. \emph{Journal of Pragmatics}, \emph{55}, 119--139. \url{https://doi.org/10.1016/j.pragma.2013.06.002}

\leavevmode\hypertarget{ref-dikyuva2015}{}%
Dikyuva, H., Makaroğlu, B., \& Arık, E. (2015). Türk i{ş}aret dili dilbilgisi kitab{ı}. \emph{Aile Ve Sosyal Politikalar Bakanl{ı}{ğ}{ı} Yay{ı}nlar{ı}: Ankara}.

\leavevmode\hypertarget{ref-dudis2004}{}%
Dudis, P. G. (2004). Body partitioning and real-space blends. \emph{Cognitive Linguistics}, \emph{15}(2). \url{https://doi.org/10.1515/cogl.2004.009}

\leavevmode\hypertarget{ref-emmorey1996}{}%
Emmorey, K. (1996). \emph{The confluence of space and language in signed languages}. In \emph{Language, Speech, and Communication} (pp. 171--209). Cambridge, MA, US: The MIT Press.

\leavevmode\hypertarget{ref-emmorey1995}{}%
Emmorey, K., Corina, D., \& Bellugi, U. (1995). \emph{Differential processing of topographic and referential functions of space.}

\leavevmode\hypertarget{ref-engberg-pedersen1993}{}%
Engberg-Pedersen, E. (1993). \emph{Space in danish sign language: The semantics and morphosyntax of the use of space in a visual language}.

\leavevmode\hypertarget{ref-ferrara2012}{}%
Ferrara, L., \& Johnston, T. (2012). Enactment in discourse: Conceptualization through language and gesture. \emph{Language, Culture and Mind V: Integrating Semiotic Resources in Communication and Creativity. Catholic University of Portugal, Lisbon}.

\leavevmode\hypertarget{ref-frederiksen2016}{}%
Frederiksen, A. T., \& Mayberry, R. I. (2016). Who's on First? Investigating the referential hierarchy in simple native ASL narratives. \emph{Lingua}, \emph{180}, 49--68. \url{https://doi.org/10.1016/j.lingua.2016.03.007}

\leavevmode\hypertarget{ref-frederiksen2019}{}%
Frederiksen, A. T., \& Mayberry, R. I. (2019). Reference tracking in early stages of different modality L2 acquisition: Limited over-explicitness in novice ASL signers{'} referring expressions. \emph{Second Language Research}, \emph{35}(2), 253--283. \url{https://doi.org/10.1177/0267658317750220}

\leavevmode\hypertarget{ref-givon1983}{}%
Givon, T. (1983). \emph{Topic continuity in discourse}. John Benjamins Publishing Company. Retrieved from \url{https://benjamins.com/catalog/tsl.3}

\leavevmode\hypertarget{ref-goldin-meadow2005}{}%
Goldin-Meadow, S. (2005). \emph{The Resilience of Language: What Gesture Creation in Deaf Children Can Tell Us about how All Children Learn Language}. Psychology Press.

\leavevmode\hypertarget{ref-grice1975}{}%
Grice, H. P. (1975). \emph{Logic and conversation}. Brill.

\leavevmode\hypertarget{ref-gullberg2006}{}%
Gullberg, M. (2006). Handling Discourse: Gestures, Reference Tracking, and Communication Strategies in Early L2: Language Learning Vol. 56, No. 1. \emph{Language Learning}, \emph{56}(1), 155--196. \url{https://doi.org/10.1111/j.0023-8333.2006.00344.x}

\leavevmode\hypertarget{ref-gundel1993}{}%
Gundel, J. K., Hedberg, N., \& Zacharski, R. (1993). Cognitive status and the form of referring expressions in discourse. \emph{Language}, \emph{69}(2), 274--307. \url{https://doi.org/10.2307/416535}

\leavevmode\hypertarget{ref-guxfcr2018}{}%
Gür, C. (2018). \emph{The Effects of Late Language Acquisition on the Development of Narrative Skills: Insights from Turkish Sign Language} (PhD thesis). Erzurum.

\leavevmode\hypertarget{ref-hickmann1999}{}%
Hickmann, M., \& Hendriks, H. (1999). Cohesion and anaphora in children's narratives: a comparison of English, French, German, and Mandarin Chinese. \emph{Journal of Child Language}, \emph{26}(2), 419--452. \url{https://doi.org/10.1017/S0305000999003785}

\leavevmode\hypertarget{ref-hickmann1996}{}%
Hickmann, M., Hendriks, H., Roland, F., \& Liang, J. (1996). The marking of new information in children's narratives: a comparison of English, French, German and Mandarin Chinese*. \emph{Journal of Child Language}, \emph{23}(3), 591--619. \url{https://doi.org/10.1017/S0305000900008965}

\leavevmode\hypertarget{ref-janzen2004}{}%
Janzen, T. (2004). Space rotation, perspective shift, and verb morphology in ASL. \emph{Cognitive Linguistics}, \emph{15}(2). \url{https://doi.org/10.1515/cogl.2004.006}

\leavevmode\hypertarget{ref-karaduxf6ller2017}{}%
Karadöller, D. Z., Sümer, B., \& Özyürek, A. (2017). \emph{Effects of Delayed Language Exposure on Spatial Language Acquisition by Signing Children and Adults}. 5.

\leavevmode\hypertarget{ref-koulidobrova2016}{}%
Koulidobrova, E., \& Lillo-Martin, D. (2016). A {`}point{'}of inquiry: The case of the (non-) pronominal IX in ASL. \emph{The Impact of Pronominal Form on Interpretation}, \emph{221}, 250.

\leavevmode\hypertarget{ref-liddell2003}{}%
Liddell, S. K. (2003). \emph{Grammar, Gesture, and Meaning in American Sign Language}. Cambridge University Press.

\leavevmode\hypertarget{ref-loew1984}{}%
Loew, R. C. (1984). \emph{Roles and reference in American sign language : a developmental perspective}. Retrieved from \href{https:///paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}{/paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}

\leavevmode\hypertarget{ref-lu2016}{}%
Lu, J., Jones, A., \& Morgan, G. (2016). The impact of input quality on early sign development in native and non-native language learners*. \emph{Journal of Child Language}, \emph{43}(3), 537--552. \url{https://doi.org/10.1017/S0305000915000835}

\leavevmode\hypertarget{ref-macsweeney2008}{}%
MacSweeney, M., Waters, D., Brammer, M. J., Woll, B., \& Goswami, U. (2008). Phonological processing in deaf signers and the impact of age of first language acquisition. \emph{NeuroImage}, \emph{40}(3), 1369--1379. \url{https://doi.org/10.1016/j.neuroimage.2007.12.047}

\leavevmode\hypertarget{ref-metzger1995}{}%
Metzger, M. (1995). Constructed action and constructed dialogue in american sign language. \emph{Sociolinguistics in Deaf Communities}, 255271.

\leavevmode\hypertarget{ref-morgan1999}{}%
Morgan, G. (1999). Event packaging in british. \emph{Storytelling and Conversation: Discourse in Deaf Communities}, \emph{5}, 27.

\leavevmode\hypertarget{ref-morgan2000}{}%
Morgan, G. (2000). \emph{Discourse cohesion in sign and speech}. \url{https://doi.org/10.1177/13670069000040030101}

\leavevmode\hypertarget{ref-morgan2005}{}%
Morgan, G. (2005). \emph{The Development of Narrative Skills in British Sign Language} (B. Schick, M. Marschark, \& P. E. Spencer, Eds.). Oxford University Press. \url{https://doi.org/10.1093/acprof:oso/9780195180947.001.0001}

\leavevmode\hypertarget{ref-newport1990}{}%
Newport, E. L. (1990). Maturational Constraints on Language Learning. \emph{Cognitive Science}, \emph{14}(1), 11--28. \url{https://doi.org/10.1207/s15516709cog1401_2}

\leavevmode\hypertarget{ref-uxf6zsoy2014}{}%
Özsoy, S., \& Nuhbalaoğlu, D. (2014). \emph{Formal and experimental approaches in sign languages}. 9--11. University of Venice.

\leavevmode\hypertarget{ref-uxf6zyuxfcrek2010}{}%
Özyürek, A., \& Perniss, P. (2010). \emph{Event representations in signed languages} (E. Pederson \& J. Bohnemeyer, Eds.). In (pp. 84--107). Cambridge: Cambridge University Press. \url{https://doi.org/10.1017/CBO9780511782039.005}

\leavevmode\hypertarget{ref-padden1986}{}%
Padden, C. (1986). \emph{Verbs and role-shifting in american sign language}. \emph{44}, 57. National Association of the Deaf Silver Spring, MD.

\leavevmode\hypertarget{ref-perniss2015}{}%
Perniss, P., \& Özyürek, A. (2015). Visible Cohesion: A Comparison of Reference Tracking in Sign, Speech, and Co-Speech Gesture. \emph{Topics in Cognitive Science}, \emph{7}(1), 36--60. \url{https://doi.org/10.1111/tops.12122}

\leavevmode\hypertarget{ref-petitto1997}{}%
Petitto, L. A. (1997). \emph{In the Beginning: On the Genetic and Environmental Factors that Make Early Language Acquisition Possible} (M. Gopnik, Ed.). England: Oxford University Press. Retrieved from \url{https://www.academia.edu/8722491/In_the_Beginning_On_the_Genetic_and_Environmental_Factors_that_Make_Early_Language_Acquisition_Possible}

\leavevmode\hypertarget{ref-prince1981}{}%
Prince, E. (1981). \emph{Toward a taxonomy of given - new information} (P. Cole, Ed.). New York: Academic Press.

\leavevmode\hypertarget{ref-rathmann2007}{}%
Rathmann, C., Mann, W., \& Morgan, G. (2007). Narrative Structure and Narrative Development in Deaf Children. \emph{Deafness \& Education International}, \emph{9}(4), 187--196. \url{https://doi.org/10.1179/146431507790559932}

\leavevmode\hypertarget{ref-sevinuxe72006}{}%
Sevinç, A. M. (2006). Grammatical relations and word order in turkish sign language (t{İ}d). \emph{Unpublished Master's Thesis, Middle East Technical University, Turkey}.

\leavevmode\hypertarget{ref-sloetjes2008}{}%
Sloetjes, H., \& Wittenburg, P. (2008, May). \emph{Annotation by category: ELAN and ISO DCR}. Marrakech, Morocco: European Language Resources Association (ELRA). Retrieved from \url{http://www.lrec-conf.org/proceedings/lrec2008/pdf/208_paper.pdf}

\leavevmode\hypertarget{ref-smith2014}{}%
Smith, S., \& Cormier, K. (2014). In or Out?: Spatial Scale and Enactment in Narratives of Native and Nonnative Signing Deaf Children Acquiring British Sign Language. \emph{Sign Language Studies}, \emph{14}(3), 275--301. \url{https://doi.org/10.1353/sls.2014.0008}

\leavevmode\hypertarget{ref-sorace2011}{}%
Sorace, A. (2011). Pinning down the concept of {``}interface{''} in bilingualism. \emph{Linguistic Approaches to Bilingualism}, \emph{1}(1), 133.

\leavevmode\hypertarget{ref-sorace2009}{}%
Sorace, A., \& Serratrice, L. (2009). Internal and external interfaces in bilingual language development: Beyond structural overlap. \emph{International Journal of Bilingualism}, \emph{13}(2), 195--210. \url{https://doi.org/10.1177/1367006909339810}

\leavevmode\hypertarget{ref-sperber1986}{}%
Sperber, D., \& Wilson, D. (1986). \emph{Relevance: Communication and cognition} (Vol. 142). Harvard University Press Cambridge, MA.

\leavevmode\hypertarget{ref-supalla1986}{}%
Supalla, T. (1986). The classifier system in american sign language. \emph{Noun Classes and Categorization}, \emph{7}, 181214.

\leavevmode\hypertarget{ref-talmy2003}{}%
Talmy, L. (2003). The representation of spatial structure in spoken and signed language. \emph{Perspectives on Classifier Constructions in Sign Language}, 169195.

\leavevmode\hypertarget{ref-vercaingne-muxe9nard2001}{}%
Vercaingne-Ménard, A., Godard, L., \& Labelle, M. (2001). The emergence of narrative discourse in two young deaf children. \emph{Signed Languages: Discoveries from International Research}, 120134.

\leavevmode\hypertarget{ref-weaver2011}{}%
Weaver, K. A., \& Starner, T. (2011). \emph{The proceedings of the 13th international ACM SIGACCESS conference}. 91. Dundee, Scotland, UK: ACM Press. \url{https://doi.org/10.1145/2049536.2049554}

\leavevmode\hypertarget{ref-williams1988}{}%
Williams, J. (1988). Zero Anaphora in Second Language Acquisition: A Comparison among Three Varieties of English. \emph{Studies in Second Language Acquisition}, \emph{10}(3), 339--370. \url{https://doi.org/10.1017/S0272263100007488}

\leavevmode\hypertarget{ref-woll2013}{}%
Woll, B. (2013). \emph{Sign language and spoken language development in young children: Measuring vocabulary by means of the CDI} (L. Meurant, A. Sinte, M. Van Herreweghe, \& M. Vermeerbergen, Eds.). Berlin, Boston: DE GRUYTER. \url{https://doi.org/10.1515/9781614511472.15}

\end{CSLReferences}

\newpage

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:table-a}Participant demographic information based on self-report}

\small{

\begin{tabular*}{\textwidth}{c @{\extracolsep{\fill}}lllllllll}
\toprule
No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{\shortstack{Hearing \\loss}} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{\shortstack{Home \\language}} & \multicolumn{1}{c}{\shortstack{Parents' \\hearing}}\\
\midrule
1 & Native & M & 0-3 & 33 & Prelingual & HS & TİD, HS & Both deaf\\
2 & Native & F & 0-3 & 27 & Prelingual & HS & TİD, HS & Both deaf\\
3 & Native & F & 0-3 & 27 & Prelingual & HS & HS & Both deaf\\
4 & Native & M & 0-3 & 29 & Prelingual & HS & TİD, HS & Both deaf\\
5 & Native & M & 0-3 & 28 & 4-7 years & HS & TİD & Both deaf\\
6 & Native & M & 0-3 & 32 & Prelingual & HS & TİD & F deaf, M hearing\\
7 & Native & M & 0-3 & 35 & Prelingual & HS & TİD & Both deaf\\
8 & Native & F & 0-3 & 23 & Prelingual & HS & TİD & Both deaf\\
9 & Native & F & 0-3 & 30 & Prelingual & HS & TİD & Both deaf\\
10 & Native & F & 0-3 & 25 & Prelingual & HS & TİD & Both deaf\\
11 & Native & F & 0-3 & 24 & Prelingual & HS & TİD & Both deaf\\
12 & Native & M & 0-3 & 26 & Prelingual & AD & TİD & Both deaf\\
13 & Native & M & 0-3 & 18 & Prelingual & HS & TİD & Both deaf\\
14 & Native & F & 0-3 & 21 & Prelingual & BA & TİD & Both deaf\\
15 & Native & M & 0-3 & 19 & Prelingual & HS & TİD & Both deaf\\
16 & Native & F & 0-3 & 25 & Prelingual & HS & TİD & Both deaf\\
17 & Late & F & 4-7 & 35 & Prelingual & AD & TİD, HS, TR & Both hearing\\
18 & Late & M & 13-17 & 50 & 4-7 years & MS & TR, HS & Both hearing\\
19 & Late & F & 8-12 & 49 & Prelingual & MS & HS, TR & Both hearing\\
20 & Late & M & 8-12 & 43 & Prelingual & BA & TR & Both hearing\\
21 & Late & M & 4-7 & 36 & 4-7 years & HS & TR, HS & Both hearing\\
22 & Late & M & 8-12 & 28 & Prelingual & BA & TR & Both hearing\\
23 & Late & M & 4-7 & 33 & Prelingual & HS & TR & Both hearing\\
24 & Late & M & 13-17 & 31 & Prelingual & BA & TR & Both hearing\\
25 & Late & F & 4-7 & 33 & Prelingual & HS & TR, HS & Both hearing\\
26 & Late & F & 8-12 & 24 & Prelingual & HS & HS & Both hearing\\
27 & Late & M & 8-12 & 33 & Prelingual & MS & HS & Both hearing\\
28 & Late & F & 4-7 & 31 & Prelingual & AD & HS & Both hearing\\
29 & Late & M & 4-7 & 35 & Prelingual & HS & HS & Both hearing\\
\bottomrule
\addlinespace
\end{tabular*}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, HS = high school, AD = associate degree, MS = middle school, BA/BS = Bachelor of Arts/Science, TR = (Spoken) Turkish.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\endgroup

\end{document}