### 2022-08-05 manuscript_v2.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Journal of Pragmatics} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% Pandoc header
\usepackage{floatrow}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{float}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\floatsetup[figure]{capposition=top}
\floatsetup[table]{capposition=top}
\usepackage{times}
\usepackage{smartdiagram}
\usepackage{ragged2e}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{./sty/langsci-gb4e}
\usepackage{caption}
\usepackage{multicol}
\setlength{\columnsep}{-3cm}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Reference Tracking Strategies of Native and Late Signers in
Turkish Sign Language}

  \begin{abstract}
  We investigated the reference tracking strategies among native and
  late adult deaf signers of Turkish Sign Language. Consistent with the
  theories of saliency and referential accessibility, regardless of
  their acquisiton groups, signers mainly used nominals and extension
  classifiers for introducing referents. To maintain a referent in the
  previous clause, signers chose null marking (e.g., constructed action,
  agreement and plain verbs). For topic shifts or re-introduced
  contexts, nominals and pronouns were chiefly favored although we
  observed very little pronominal use. As for age of acquisition
  effects, we only report limited over-redundancy for late signers who
  used null marking less compared to native signers especially for
  introduced and maintained contexts. Given our limited support for
  redundancy by late signers, we conclude that native signers might
  utilize the spatial affordances of the visual-spatial modality better
  than late signers only to a certain extent.
  \end{abstract}
   \begin{keyword} Referential accessibility, signed narratives,
pragmatic competence, delayed language acquisition, Turkish Sign
Language\end{keyword}
 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Narratives are extensive discourses that construct an important part of
human communication. In order to create a systematic and coherent
organization of experiences, narrators must choose appropriate
linguistic forms to refer to entities (Ariel, 1990; Givon, 1983; Gundel
et al., 1993): Certain forms such as nominal expressions are commonly
used for the first mention of a referent in a discourse, and other forms
such as pronouns and null marking may track previously mentioned
entities which could either be maintained across multiple clauses or
re-introduced back into the discourse. The selection of these referring
expressions (henceforth REs) is contingent on cognitive abilities, such
that individuals must rely on the activation of the working memory to
keep in mind previously referred expressions and long strings of
information (Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play
a pivotal role in the tracking of referents. The form with the least
amount of redundant information must be selected based on the
accessibility levels of the referent (i.e., how active a referent is in
the addressee's mind given a point in discourse) (Ahn, 2019; Ariel,
1990). Because of these cognitive and pragmatic necessities, it is not
surprising to observe that producing a coherent narrative is one of the
latest milestones in language acquisition among children (Hickmann et
al., 1996). As can be expected, there is a plethora of research
examining how narrative competence is achieved in both monolingual and
bilingual acquisition contexts among children and adults. Although
narrative competence and referent selection have been studied
extensively for typically developing speakers (Aksu-Koç and
Nicolopoulou, 2015; Hickmann et al., 1996; Williams, 1988), deaf
individuals who acquired a sign language in atypical learning contexts
(i.e., deaf signers with linguistic deprivation) have remained mostly
understudied with an exception of handful of studies (Becker, 2009;
Cormier et al., 2013; Gür, 2018). Nevertheless, it is paramount that we
investigate and describe the effects of language deprivation among deaf
adults on narrative competence more in depth. This will help us
understand whether the loss of systematic and frequent language input in
the early years of life affect how deaf signers produce a narrative in
their adulthood. Given this, the present study will focus on deaf adult
signers with language deprivation and examine their reference tracking
strategies in signed narratives.

Sign Languages are natural languages of the Deaf communities all around
the world. They are independent languages, which do not derive from
majority spoken language(s) but emerge naturally when deaf people form a
community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and has
been the officially recognized language of the Deaf community in Turkey
since 2005 (see Nuhbalaoglu, 2018 for a brief review of the historical
development of the language). Deaf or hearing children acquiring a SL
within a family where there is one or more deaf caregivers display the
same developmental milestones as hearing children acquiring a spoken
language in a hearing family (Chamberlain et al., 1999; Meier and
Newport, 1990). However, reports indicate that over 90\% of all deaf
individuals are born into hearing families who do not know how to sign
(i.e., those with language deprivation) (Woll, 2013). This means that
the auditory modality used in the household is not accessible to many
deaf children. That is why the age of acquisition of a first functional
language varies greatly among deaf children, with many of whom being
deprived of their first languages typically until after age 7 when they
start formal education in a deaf school. First language deprivation or
delay has been found to have negative effects on certain cognitive and
linguistic measures among late signing children. To name a few, delayed
exposure to a first functional language is reported to negatively affect
deaf children's morphosyntactic production and processing (Boudreault
and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et al., 2012;
Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport, 1990) and
narrative abilities (e.g., Becker, 2009). In our overall research, we
are concerned with the effects of this late exposure in the ultimate
adult forms of TİD users.

\hypertarget{referring-expression-re-selection}{%
\subsection{Referring Expression (RE)
Selection}\label{referring-expression-re-selection}}

Natural languages can be systematically used to keep track of previous
referents in the discourse and mark the changing cognitive statuses of
the REs (i.e., from being active to less active in the addressee's mind)
to create a more coherent narrative. Among the theoretical frameworks
that aim to explain that different linguistic forms are selected in
accordance with the addressee's mental representations of the referents
in the discourse is the Givenness Hierarchy (Gundel et al., 1993).
According to this model, given or familiar information is expressed in
more attenuated forms such as pronominal use and null marking since the
mental status of the expressed referent is already at the center of
attention, as indicated by the present contextual information (Chafe,
1976). The Givenness Hierarchy is akin to other previously proposed
models of reference control within the topicality and saliency paradigm
such as the Accessibility Hierarchy (Ariel, 1990), Topic Continuity
(Givon, 1983), and Familiarity Scale (Prince, 1981). Gundel et al.
(1993) argued that the distribution of forms is not random but abides by
the Gricean Maxim of Quantity (Grice, 1975). That is to say, individuals
must make the most informative and succinct choice possible to retain
economy of form. When multiple expressions are available for the same
referent as in an utterance like (1), phonetically simplest form is
selected with the least redundant information are selected based on an
economy principle. In (1), REs for ``Eddie'' are given in bold. All
three REs (nominal, pronominal, and null marking) refer to the same
entity; however, the quantity of marking decreases as the referent's
accessibility increases. On this ground, for highly accessible referents
(i.e., those that are highly active in the mind), linguistic forms with
the least phonetic content are selected for the least amount of
processing effort.

\begin{exe}
    \ex
    So \textbf{Eddie} turned around. \hfill Williams (1988; p. 343, ex. 7) \\
\textbf{He} said, "youse got a problem?" \\
"Yeah, we want you," they say.
So-\textbf{$\emptyset$} walked right up to them \\
\end{exe}

It is also possible to categorize REs into three according to their
discourse status (Gullberg, 2006). A referent can be introduced or
mentioned for the first time, maintained, that is, continued across two
or more clauses, or re-introduced into the discourse following a topic
shift. As a result, interlocutors mark some REs differently than others
based on the principles of reference accessibility. For spoken
languages, this observation is well established (Ahrenholz, 2005; Ariel,
1990; Azar and Özyürek, 2015; Chafe, 1976; Debreslioska et al., 2013;
Givon, 1983; Gundel et al., 1993; Hickmann et al., 1996). These research
indicate that the introduction and re-introduction of a referent are
often accomplished by full noun phrases which can be commonly modified
by an adjective or demonstrative. Maintenance, on the other hand, is
typically marked with pronominals and null markers since the continued
referent is still highly accessible to the addressee.

However, this particular observation is not peculiar to spoken
languages. There is an emerging body of literature that reports the
distribution of REs in signed narratives. Building on Gundel et al.
(1993)'s work, Swabey (2002) demonstrated that, like other spoken
languages such as Russian, Chinese, and Japanese, American Sign Language
(ASL) signers use bare nouns for introduced and reintroduced discourse
contexts whereas null marking is used for referents which are
``in-focus'' or very accessible to the addressee. Overt REs like
pronouns or nouns are disprefered for such referents. Swabey (2002)
concluded that ``native'' signers in her study abided by the Gricean
maxim of quantity and were thus economic in their reference tracking.
Similar findings have been reported for German Sign Language (DGS)
(Perniss and Özyürek, 2015) and British Sign Language (BSL) (Morgan,
2005). It is important to note, however, that these research either
collapsed native and late signers into a single group and labeled them
as ``native-like'' or explicitly compared native deaf signers with late
deaf signers (with language deprivation) in terms of narrative
competence.

\hypertarget{signed-vs.-spoken-res}{%
\subsubsection{Signed vs.~Spoken REs}\label{signed-vs.-spoken-res}}

Commonly in both spoken and sign languages nouns and noun phrases are
used as referring expressions. Spoken languages use pronouns, determines
and demonstratives where sign languages use a pointing sign, usually
glossed as IX, which assumes similar functions (Lillo-Martin and Klima,
1990). In addition, the pointing sign is used to referentially anchor a
referent to a location in signing space. For instance, when the signer
introduces \textsc{elvan} for the first time, s/he can assign her to the
right. This referential locus can later be used by agreement verbs and
classifier predicates (more below). For instance, in (2a) there are two
referents, \textsc{elvan}, who is referentially-assigned to the right
side of the signer, represented with the ``a'' subscript and
\textsc{demet}, who is referentially-assigned to the left side of the
signer, represented with the ``b'' subscript. When the signer makes a
contrast in (2b), as a pronominal use, she points to the right side to
refer to ELVAN and to the left to refer to \textsc{demet}. We
schematically represents the referential locus assigned to the referents
in (3).

\begin{exe}
    \ex 
\begin{xlist}
    \ex
    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
    \\
    \ex
    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
    \end{xlist}
    \end{exe}

    \begin{exe}
    \ex Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Curve Lines [id:da7006577339157659] 
\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
%Straight Lines [id:da8428827309782937] 
\draw    (112.9,128.8) -- (210.21,128.8) ;
%Curve Lines [id:da39290455666055124] 
\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
%Curve Lines [id:da9909903731270634] 
\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;

% Text Node
\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
% Text Node
\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
% Text Node
\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
% Text Node
\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
% Text Node
\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
\end{tikzpicture}
    \end{exe}

Grammatical gender as in Romance languages or noun classes as in Bantu
languages are some reference tracking tools that are familiar from
spoken languages. A somehow less known reference-tracking tool in spoken
languages is \emph{verbal classifiers}, which morphologically attach to
the verb and often cross-reference a core argument in a clause
(Aikhenvald, 2017). Classifiers denote sets of objects based on semantic
criteria such as size, shape, body-part, texture and sometimes function.
Usually, the verbal classifier and the cross-referenced noun co-occur in
the same clause when the argument is introduced into discourse. The
classifier can track the argument on its own on subsequent occurrences.
In sign languages, a verbal classifier morpheme is encoded by a
\emph{handshape} that is articulated phonologically co-temporally with a
movement, which encodes the verbal meaning (Benedicto and Brentari,
2004; Supalla, 1990). These two infused morphemes form a complex
predicate. Some classifier handshapes and the set of objects that they
denote are below. Some classifier handshapes and the set of objects that
they denote are presented below in (4).

\begin{exe}
\ex 
\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
\end{exe}

In (5), in the first clause, a referential locus, i.e.~locus-a, is
assigned to the locative source argument \textsc{cage}. In the second
clause, the core argument \textsc{bird} is cross-referenced with the
classifier handshape, i.e.~small-legged animal. It flies from the
referential locus of the \textsc{cage} (locus-a) to the referential
locus of the goal \textsc{tree} (locus-b).

\begin{exe}
    \ex 
\begin{xlist}
    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
    \textsc{garden bird cage-a open} \\
    ‘In the garden, there is a cage and it is open.’
    \\

    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
    \textsc{bird fly[manner]-a tree-b land. CL-b}\\
    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
    \end{xlist}
    \end{exe}
    \hfill (Adapted from Gökgöz and Sevgi 2020:318)

\hypertarget{linguistic-tools-for-tracking-referents-in-signed-narratives}{%
\subsection{Linguistic Tools for Tracking Referents in Signed
Narratives}\label{linguistic-tools-for-tracking-referents-in-signed-narratives}}

The linguistic inventory for discourse referents can include nominal
constructions (bare nouns, fingerspelled nouns, and modified nouns),
null markers (plain verb, agreement verb, and constructed action),
classifier predicates (whole entity classifiers, body part classifiers,
handling classifiers, and extension classifiers), and overt pronouns
(pronominal index or IX) (Frederiksen and Mayberry, 2016). This interim
section describes such constructions attested for sign languages under
four main headings (nominals, pronouns, classifiers, and null marking).

\clearpage

\begin{multicols}{2}
\begin{exe}
\ex
\begin{xlist}
\ex Full noun \\\glll
\includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
\ex Noun modified by classifier \\\glll
\includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
\ex Fingerspelled noun\\\glll 
\includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
\ex Noun modified by index (IX)\\\glll 
\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ IX \textsc{cat} \\ \\
\end{xlist}
\end{exe}
\end{multicols}
\ea 
\ea Pronominal \\\glll
\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ IX \textsc{sick} \\ \\
\z \z
\begin{multicols}{4}
\begin{exe}
\ex
\begin{xlist}
\ex Whole entity \\\glll 
\includegraphics[width=50pt]{pictures/3e.png} \\ \textsc{fish.fall}\\ \\
\ex Body part \\\glll
\includegraphics[width=50pt]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
\ex Handling \\\glll 
\includegraphics[width=50pt]{pictures/3c.png} \\ \textsc{hold} \\ \\
\ex Extension \\\glll 
\includegraphics[width=50pt]{pictures/3d_1.png} \includegraphics[width=50pt]{pictures/3a.png}\\ \textsc{plank} {\textsc{dh: mouse}} \\ {} {\textsc{ndh: tunnel}}\\ \\
\end{xlist}
\end{exe}
\end{multicols}
\begin{multicols}{3}
\begin{exe}
\ex
\begin{xlist}
\ex Constructed action \\\glll 
\includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\

\ex Plain verb  \\\glll
\includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\

\ex Agreement verb  \\\glll 
\includegraphics[width=50pt]{pictures/4b_1.png} 
\includegraphics[width=50pt]{pictures/4b_2.png} \\ {\textsc{$_{1}$hit$_{3}$}} {} \\ \\ 
\end{xlist}
\end{exe}
\end{multicols}
\footnotetext{We use small capitals to indicate sign glosses.}

\hypertarget{nominals}{%
\subsubsection{Nominals}\label{nominals}}

In TİD and many other sign languages, it is possible to use bare noun
phrases (\textsc{bare nom}) as in (1a) as well as fingerspelled nouns
(\textsc{fs nom}) as in (1c). Fingerspelling refers to the use of the
manual alphabet which is derived or borrowed from a surrounding written
language (Kubuş, 2008). Nouns can also be modified (\textsc{mod nom}) by
classifiers as in (1b) (Frederiksen and Mayberry, 2016) or by a
prenominal or postnominal IX functioning as a determiner or a
demonstrative as in (1d) (Maclaughlin, 1997).

\hypertarget{pronouns}{%
\subsubsection{Pronouns}\label{pronouns}}

The IX sign (extended index finger) has various functions in sign
languages, one of which is the pronominal use which refers to
established nominals by pointing to a previously assigned abstract or
physical locus in the signing space when re-introducing a topic in
discourse (Emmorey, 1996). The same IX function has also been identified
for TİD (Sevinç, 2006). The example in (2) illustrates a
spatially-anchored pronominal reference for topic re-introduction. These
signs are thus rarely used for the first mention of an entity (Perniss
and Özyürek, 2015). Other interpretations consider pointing signs
without noun phrases as examples of demonstratives and not personal
pronouns (Koulidobrova and Lillo-Martin, 2016). It is important to
mention here that in some spoken languages (e.g., German and Dutch),
demonstratives can also refer to persons (Bosch and Umbach, 2007) and
behave similarly to personal pronouns. However, for the purposes of the
present study, we will refer to IX constructions as pronominal IX and
particularly examine their non-deictic or personal pronoun realizations
as a possible discourse strategy to mainly mark topic shift among native
and late signers of TİD.

\hypertarget{classifiers}{%
\subsubsection{Classifiers}\label{classifiers}}

Classifier predicates are highly complex morphological constructions
which are very common in sign languages (Supalla, 1986). These are
manually signed iconic representations of animate or inanimate objects,
and they embody several properties pertaining to the nominal referent
including its location in the signing space, movement, path, size, and
shape (Talmy, 2003). As in most sign languages hitherto studied,
classifier constructions are also common and found in TİD (see Dikyuva
et al., 2015 for a review). Following the classification of
Engberg-Pedersen (1993), we distinguish between four types of classifier
constructions. Whole entity classifiers (WECL) represent the entirety of
any referent whereas body part classifiers (BPCL) only reflect a
specific part of an animate entity (e.g., limbs, head, tongue, teeth,
and so forth). Handling classifiers (HCL) occur with transitive verbs
and depict how an object is handled or manipulated by agent. On this
ground, an HCL\footnote{As discussed before, an HCL is transitive in
  nature and depicts the handling action performed by ananimate agent.
  This is why most research tend to consider HCLs as an extension to
  constructed action (Cormier et al., 2013; Dudis, 2004).} depicts the
manner of the referring action while the former two constructions encode
the location of the entity as well as its motion in space. Extension
classifiers (ExtCL), on the other hand, denote the size and shape of an
entity by following a trajectory to trace its shape. We exemplify these
constructions in (3).

\hypertarget{null-marking}{%
\subsubsection{Null marking}\label{null-marking}}

Parallel to spoken languages, many sign languages have been reported to
allow null subjects and/or objects and make use of null arguments when
referring to highly accessible referents which are still active in the
mental representation of the addressee. In signed discourse, this can be
achieved by employing certain verbal constructions. To identify the
predicate types in ASL, Padden (1986) suggested a triadic categorization
of verbs in sign languages: plain verbs, agreement verbs, and spatial
verbs. According to this classification, plain verbs (plain) (4b) do not
agree with the subject or the object. This means that they are not
inflected for number or person. This contrasts with agreeing verbs
(agreement) (4c) the starting or the ending point of which can be
modified to specific loci for person and number agreement while spatial
verbs (as in \textsc{ix book put} `I placed the book there') agree with
locative arguments but like plain verbs, they are not inflected for
number or person. TİD too licenses null arguments. Following this
classification, Sevinç (2006) argued that although it is common in TİD
to drop arguments, there were certain limitations on the null marking of
the arguments for agreement and plain verbs.

\hypertarget{learning-how-to-track-referents-in-sign-languages}{%
\subsection{Learning How to Track Referents in Sign
Languages}\label{learning-how-to-track-referents-in-sign-languages}}

In both communication modalities (visual and auditory), children start
to produce more complex and coherent stories with appropriate and
informative linguistic forms to indicate anaphoric or cataphoric
references after age 10 (Hickmann et al., 1996; Rathmann et al., 2007).
Prior to this age, both signing and speaking children fail to refer to
entities consistently since their pragmatic abilities are yet to
develop. The manipulation of linguistic forms for discourse functions to
signal topic continuance or shift is difficult to master until the
beginning of formal education. These findings are also comparable across
different sign languages including British Sign Language (BSL) (Morgan,
2005), ASL (Loew, 1984), and German Sign Language (Deutsche
Gebärdensprache - DGS) (Becker, 2009), pointing out to the fact that the
mastery of reference tracking in narratives is one of the most
protracted achievements in the acquisition of signed discourse.

Like in first language development, narrative skills have been
extensively studied among second language learners (see Bel et al.,
2015; Frederiksen and Mayberry, 2019). Comparing 11 deaf adults signers
and 13 hearing advanced learners of Catalan Sign Language (Llengua de
Signes Catalana - LSC), Bel et al. (2015) found that hearing signers
resort to more overt forms (nominals and pronouns) than native signers
to avoid potential ambiguities among referents. Nevertheless, their
results did not reach statistical significant for the interaction of
discourse status and acquisition group. A similar line of research came
from Frederiksen and Mayberry (2019) who found a statistically
significant difference between both groups, that is native signers and
hearing L2 learners of ASL, only for classifier use especially in
re-introduced contexts. L2 learners in the study used more nominals and
fewer instances of null marking in number, indicative of a limited
effect of overexplicitness among second language learners. However,
these studies treated late and native deaf signers as a single group.
That is, the control group consisted of both native signers who had one
or more deaf parents and also native-like (Cormier et al., 2012) late
deaf signers who had hearing parents and their age of acquisition
varied. Given the preliminary evidence that topic continuity and
salience marking is also difficult for second language learners
regardless of modality, it has been suggested under the Interface
Hypothesis that the syntax-discourse interface is vulnerable and
especially challenging to acquire (Sorace, 2011) for second language
learners. Similarly, Sorace and Serratrice (2009) discuss that this
interface entails much more than morphosyntactic knowledge but also
includes cognitive and pragmatic resources that help provide discourse
coherence. They propose that the second language discourse is
over-redundant. This means that second language learners produce more
overt pronouns or fuller forms of language for highly accessible
referents possibly as a compensatory strategy in an attempt to reduce
the increasing load on cognitive processing.

Still, only very few research examined whether linguistically deprived
signers (i.e., deaf signers with delayed exposure to a sign language)
also display redundancy in their reference tracking. Becker (2009)
examined 6 DGS signing deaf children (3 native and 3 late)`s narrative
abilities. She reported that late signing children generated more
explicit forms such as nominals when referring to the characters in the
story compared to native learners, concluding that first language
deprivation results in an inadequate representation of the addressee and
less efficient use of morphosyntax. Comparably, Cormier et al. (2013)
investigated the use of constructed action in 15 BSL among deaf native
and late adult signers and found that late signers were redundant in
topic continuity. In other words, the late signers produced more
explicit forms like nominals, and fewer instances of null marking when
maintaining previous referents. Most relevantly to the present study,
Gür (2018) studied the development of narrative skills among 10 native
and 10 late deaf children and deaf adult signers of TİD. She found that
native-signing children and adults had more instances of explicit
referent introduction, set the scene more frequently, and employed
constructed action in an appropriate manner more than corresponding
late-signing children and adults. Although this study presented
preliminary findings on age of acquisition effects on narrative
competence, it does not investigate how native and late signers employ
different linguistic tools based on the accessibility of the referent.
Therefore, to our knowledge, there is a lack of research directly
examining the effects of first language deprivation on deaf adult
signers' selection of REs in signed narratives.

Considering this, it is not very clear whether the negative effects on
narrative abilities observed among late signing deaf children hold when
they become adults. Understanding the impact of language deprivation or
delayed first language learning on narrative competence is crucial for
three reasons. First, studying this population will shed light on the
possible effect of lack of early language experience on future cognitive
and pragmatic abilities since the narrative competence is dependent on
such faculties. Second, examining reference tracking strategies in
particular will also inform about the linguistic inventory that native
and late deaf signers use to refer to entities. Finally, any possible
difference between the two acquisition groups in terms of narrative
competence will have further practical implications for the importance
of early access to a first functional sign language to retain future
linguistic and cognitive abilities.

\hypertarget{present-study}{%
\subsection{Present Study}\label{present-study}}

In an attempt at addressing this gap in the literature, the present
study aims to investigate the reference tracking strategies of native
and late adult signers in Turkish Sign Language (Türk İşaret Dili - TİD)
narratives with a story-telling paradigm. We explore the forms available
for referent tracking in TİD and examine the nature of RE selection as a
possible indicator of pragmatic competence between our acquisition
groups.

Overall, considering what has been suggested for the distribution of REs
in signed narratives, we expect more explicit forms like nominals for
referent introduction and and less explicit forms such as null marking,
pronominal and/or classifier use for referent maintainance. For the
re-introduction contexts, where referents which are discontinued across
one or more clauses are brought back into the discourse, we expect a
less clear picture. Some re-introduced referents will have higher
accessibility and will thus be tracked with more implicit REs, while
others that have lower accessibility will be tracked with more explicit
REs Also based on previous research, we hypothesize that late signers
will become more over-redundant in their reference tracking compared to
native signers. This would translate to a preference by late signers for
nominals over null marking especially in maintenance and re-introduction
contexts. In contrast, we would expect that native signers use more null
marking to track referents.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We tested 29 deaf adult signers, 15 with exposure to TİD from birth
(i.e., native signers), 14 with exposure between ages 3 and 17 (i.e.,
late signers) (\emph{M}\textsubscript{Age of acquisition} = 7.7 years,
\emph{SD} = 3.3 years). The criterion to be classified as a late signer
was to have hearing and non-signing caregivers. All the signers
satisfying this criterion were regarded as late signers
\footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers. Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3. In order to capture in-group variation among late signers, we run an additional supplementary analysis}.
All participants indicated that their preferred language of
communication was TİD which they used for most contexts. Detailed
information about the linguistic and educational background of the
participants can be found in \ref{section:a}.

All participating native signers (7 males, 8 females) were born into
deaf families and started acquiring TİD from at least one of their
caregivers after birth. The parents of 14 participating native signers
were deaf. One native signer had a deaf father and a hearing mother. In
addition to TİD, the home language for some native signers (n = 4) in
early childhood included the use of home
sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers. Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}.
The age of the native participants ranged between 18 and 35 (\emph{M} =
26.4 years, \emph{SD} = 5.0 years). This also corresponded to the total
average years of TİD use for native signers. All native signers were
pre-lingually deaf, that is they were either born deaf or they had
become deaf before the age of 3, with the exception of one signer who
became deaf between the ages of 4 and 7. Native signers reported to
complete at least high school education and their level of education
(LOE) was between 12 and 16 years (\emph{M} = 12.7 years, \emph{SD} =
1.5 years). In addition, all native participants attended at least two
deaf schools, one being primary and the other middle school. 10 native
participants also attended a deaf school for their high school
education. Based on a self-reported 5-point scale of proficiency (1:
poor, 5: proficient), native participants declared that they have high
receptive and productive skills in TİD (\emph{M} = 5 points, \emph{SD} =
0 points). As for their proficiency in (spoken) Turkish, their
self-reported rankings were on average 3.7 (\emph{SD} = 0.7), 3.7
(\emph{SD} = 0.7), and 2.6 (\emph{SD} = 1.2) for reading, writing, and
speaking abilities, respectively.

Comparably, late learners (8 males, 6 females) were all born into
hearing families and started learning TİD later in life after age 3,
usually with the start of formal education. The age of sign language
acquisition for late signers ranged between years 3-17 (\emph{M} = 8.7
years, \emph{SD} = 3.5 years). The parents of all participating late
signers were hearing, and the reported home languages used in childhood
mainly consisted of Turkish and home sign. The average age of late
signers was 34.6 years (\emph{SD} = 8 years) and ranged between 24 and
50. The derived mean number of years for TİD use following its
acquisition was comparable to native signers with 25.9 years (\emph{SD}
= 7.9 years). Except for two signers who became deaf between the ages of
4 and 7, all the other late signers were pre-lingually deaf. All late
signers completed middle school education and the average year of
schooling was 12.3 years (\emph{SD} = 2.8 years). All late participants
attended at least one deaf school. While 8 participants attended two
deaf schools (either primary and middle school, or middle and high
school), 3 participants reported attend three deaf schools. Using the
same scale, late signers self-rated their abilities in TİD and (spoken)
Turkish. Both of their receptive and productive TİD abilities rated on
average 4.9 points (\emph{SD} = 0.3 points). Late signers' mean rating
for their reading abilities in Turkish was 3.5 points (\emph{SD} = 0.7
points) while their writing and speaking abilities rated on average 3.4
(\emph{SD} = 0.5 points) and 3.1 points (\emph{SD} = 0.7 points),
respectively.

Native and late participants' sex, average years of schooling, TİD use
following its acquisition, and their self-reported receptive and
productive TİD ratings, and their reading, writing, and speaking ratings
for Turkish did not statistically differ; Bayesian \emph{p}'s
\textgreater{} 0.1. However, participants' ages statistically differed
between the two acquisition groups; \emph{t}(27) = 3.4, Bayesian
\emph{p} \textless{} 0.01. Overall, the native group consisted of
younger participants. This was to ensure that the average years of TİD
use among native signers approximated those of late signers.

\hypertarget{stimuli}{%
\subsection{Stimuli}\label{stimuli}}

We used a story-telling paradigm to examine how native and late signers
of TİD track reference in signed discourse. Participants were presented
with 10 short video clips extracted from the cartoon \emph{Tom and
Jerry} and asked to retell them. Each clip had six to ten episodes
(\emph{M} = 8 episodes). The duration of the clips ranged from 17
seconds to 28 seconds (\emph{M} = 22.7 seconds). The narratives included
on average 2 recurring animate characters for which the present context
required the introduction, maintenance, and re-introduction of
referents. In addition, in each narrative, there were on average 2
inanimate objects with which the animate referents interacted. The
detailed description of the episodes in each narrative can be found in
\ref{section:b}.

\begin{table}[p]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:tab-2}Annotation Tiers with their Tags and Descriptions}

\footnotesize{

\begin{tabular}{p{3.2cm}p{3.8cm}p{3.8cm}}
\toprule
Tier & \multicolumn{1}{c}{Tag} & \multicolumn{1}{c}{Description}\\
\midrule
Clause Boundary &  & The boundary between clauses indicated by the presence of predicates and/or certain prosodic cues (e.g., head nod)\\
Discourse Status & Introduced & A first mention of an entity in a given discourse regardless of any syntactic role\\
 & Maintained & A subject referent continued across two or more immediate clauses\\
 & Reintroduced & A subject referent discontinued across at least one clause\\
REAT Type & NOM: Bare Noun & A non-modified noun phrase (e.g., CAT, MOUSE)\\
 & NOM: Modified Noun & A noun phrase modified by an index sign or an adjective\\
 & NOM: Fingerspelled Noun & Borrowed nouns from Turkish spelled by the manual alphabet (e.g., K-E-D-İ ‘cat’)\\
 & PRO: Pronominal IX & An index sign used in isolation to refer to an entity in discourse\\
 & CL: Whole Entity Classifier & A manual construction that refers to the entirety of an entity\\
 & CL: Body Part Classifier & A manual construction that refers to a part of an entity\\
 & CL: Handling Classifier & A manual construction that refers to the handling of an item\\
 & CL: Extension Classifier & A manual construction that denotes the size and shape of an entity\\
 & NULL: Constructed Action & A multifunctional construction where the signers take on the role of a referent and perform their actions\\
 & NULL: Plain Verb & A non-agreeing verb (e.g., LOVE)\\
 & NULL: Agreement Verb & A verbal construction the beginning and ending of which agree with person and number\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{task-procedure}{%
\subsection{Task Procedure}\label{task-procedure}}

Participants sat in front of a laptop computer which had the stimulus
stories. The instructions were all given by a deaf research assistant
who was present in all sessions and ensured that the task requirements
were clear. The informed consent form was translated into TİD and
explained to the participants by the same assistant before data
recording. Participants filled in the consent form and a background
questionnaire form before the onset of the trials. Following this step,
signers were seated across a hearing researcher who is an advanced late
learner of TİD. We instructed the participants to narrate as they would
to a deaf friend. All trials were recorded by a digital camera placed in
front of the signers. The participants were instructed to watch the
narrative clip as many times as they needed, and narrate the content of
each clip either to the experimenter or the recording digital
camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons. First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter. Therefore, most signers preferred to narrate to the camera. Furhermore, some signers (4 in number) did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available. We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers in Appendix. The results imply that there is not much of a difference.}.

\hypertarget{data-annotation}{%
\subsection{Data Annotation}\label{data-annotation}}

We coded the discourse status (Introduction, Maintenance,
Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null). We
used ELAN Linguistic Annotation Software (Crasborn and Sloetjes, 2008)
for data annotation. First Author who is a hearing signer of TİD
annotated all the data and translated the narratives into Turkish
together with a native deaf signer of TİD.

Following Berman and Slobin (2013), we first identified the clause
boundaries in each sentence indicated by the presence of predicates and
informative non-manual markers (e.g., head nod). On the second tier, we
coded the discourse status for each referent. Having adapted the
classification of Gullberg (2006), we considered all referents mentioned
for the first time as introduced regardless of their syntactic function
in the sentence. Only subjects were coded for maintenance and
re-introduction. We coded subject referents as maintained only if they
had been referred to in the previous clause. To this end, the referents
which had been discontinued across one or more sentences were all
regarded re-introduced into the discourse. On a third tier, we had the
RE type for a single reference with the main categories nominal,
pronominal, null marking, and classifier. Under nominal, we identified
bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we
distinguished among three types of constructions: WECL, BPCL, and ExtCL.
Finally, we had CA, agreement, plain tags for null constructions. We
were unable to observe any instance of HCLs occurring in isolation
(i.e., not part of constructed action). The annotation tiers are
presented in Table \ref{tab:tab-2}.

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

In our analysis, we use Bayesian regression models with the help of the
brms package (Bürkner, 2018) in R. In a Bayesian regression the prior
and the likelihood is used to calculate the posterior. Instead of a
single filter of significance, we get a posterior probability
distribution. We do not define our own priors and use the brms package
defaults. We are using Bayesian regression to avoid some common pitfalls
of other approaches (Kruschke, 2011; Wagenmakers, 2007).

For this particular type of experimental data, we fit the responses to a
multinomial distribution. Most of the time, experimental results have
yes/no-true/false responses which are fit to binomial distributions,
reading/response times which are fit to log-normal distributions, and
ratings which are fit to cumulative distributions. In these cases, the
researcher is not particularly interested in the `intercept' coefficient
which shows the mean value of the response. The researcher is usually
interested in the remaining coefficients which give information about
predictor level comparisons.

In a multinomial distribution, the responses are in a categorical
relationship where the levels can be A,B,C, and more. These levels are
not ordered in time, they are not cumulative, and they are not additive.
If we were to make an analogy, they are like countries on Earth which
can have differences and similarities, but they don't have to. In a
multinomial distribution, one of the responses becomes the level of
comparison and other levels are compared against it. In a way, there are
multiple binomial distributions. However, the intercepts in a
multinomial distribution end up having more of an interest for the
researcher. This is because the intercept is a comparison of two levels
of the response to the exclusion of the remaining response levels. There
is another consideration that should go into interpreting the results of
a multinomial distribution. Different from the intercepts, the remaining
coefficients are predictor level comparisons within the grand comparison
of the response levels. To indicate which predictor level comparison
belongs to which response comparison, we name our coefficients with
prefixes followed by an underscore.

In our model plots, we report median estimates as points, 50\% credible
intervals as a thick line and 95\% credible intervals as a thin line.
These numbers are relatively arbitrary, and they are intended to show
the posterior probability distribution of the coefficients. We use
interval plots to save space and have a more compact plot. If a
considerable amount (\textgreater90\%) of the posterior probability
distribution is towards a sign (-/+) we interpret it as a categorial
effect, if a good amount (\textgreater50\%) of the distribution is
towards a sign (-/+) we interpret it as a tendency, and we see the
remaining distributions as indifference. In addition to these, we look
at how the distribution skewed in the extremes or how `wide' it is from
the median estimate to inform us mainly about uncertainty and variation.

\hypertarget{results}{%
\section{Results}\label{results}}

We collected 4142 data points as a result of the experiment. Each data
point is a referent tracked by the participant. Each of the REs is
classified in terms of its Discourse Status (Introduction, Maintenance,
Re-introduction), its Type (Nominal, Pronominal, Classifer,
Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null). However, we allude to the numerical distribution of the subtypes in this and following sections},
and Participant Group (Native,
Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17). Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}.

Across our acquisition groups, we observed 1066 (0.26) instances of
referent introduction, 1651 (0.4) instances of referent maintenance, and
1425 (0.34) instances of referent re-introduction. For the introduction
of a referent, the majority of the REs were nominal constructions (Table
\ref{tab:my-table}). However, we observed a sporadic use of classifiers
and null marking respectively. When the discourse status was maintained,
signers extensively used null markers to refer to the entities in the
immediately previous clause. Classifier constructions followed null
markers in frequency in maintained contexts. We observed only a few
instances of nominal use for referent maintenance. In re-introduction
contexts, signers used a greater proportion of nominal markings which
were followed by null, classifier, and pronominal use.

\begin{table}
\small
\caption{Distribution of all discourse status by referring expression types}
\label{tab:my-table}
\begin{tabular}{llll}
\hline
                     & Introduced    & Maintained     & Re-introduced \\ \hline
\textit{Nominal:}    & 92.78\% (989) & 5.63\% (93)    & 63.65\% (907) \\
\textit{Pronominal:} & 0.09\% (1)    & 0.24\% (4)     & 2.46\% (35)   \\
\textit{Classifier:} & 5.35\% (57)   & 18.29\% (302)  & 8\% (114)     \\
\textit{Null:}       & 1.78\% (19)   & 75.83\% (1252) & 25.89\% (369) \\ \hline
\end{tabular}%
\end{table}

Table \ref{tab:my-table2} shows that for all discourse statuses the most
numerous nominal type was \textsc{bare nom} which was followed by nouns
modified by index or classifiers. In number, signers used
\textsc{mod nom} more in introduction and re-introduction compared to
maintenance settings. Signers produced a total number of 40 pronouns
(Table \ref{tab:my-table3}). A great majority of the pronominal
constructions were used to re-introduce a referent into the discourse.
We observed only a very small amount of pronoun use for referent
introduction and maintenance.

\begin{table}
\small
\caption{Distribution of nominal referring expressions by discourse status}
\label{tab:my-table2}
\begin{tabular}{lllll}
\hline
                        & Bare Noun  & Modified Noun & FS Noun \\ \hline
\textit{Introduced:}    & 48.93\% (842) & 53.56\% (139)     & 100\% (8) \\
\textit{Maintained:}    & 4.82\% (83)  & 3.85\% (10)      & 0\% (0) \\
\textit{Re-introduced:} & 46.25\% (796) & 42.69\% (111)     & 0\% (0) \\ \hline
\end{tabular}
\end{table}

\begin{table}
\small
\caption{Distribution of pronominal referring expressions by discourse status}
\label{tab:my-table3}
\begin{tabular}{ll}
\hline
                        & Pronominal  \\ \hline
\textit{Introduced:}    & 2.5\% (1)   \\
\textit{Maintained:}    & 10\% (4)    \\
\textit{Re-introduced:} & 87.5\% (35) \\ \hline
\end{tabular}
\end{table}

As for the classifier constructions (Table \ref{tab:my-table4}), WECL
was the most common type, followed by BPCL and ExtCL. For the first
mention of a referent, signers mostly preferred to use ExtCL. However,
for referent maintenance and re-introduction WECL was used more compared
to BPCL and ExtCL.

Among the null markers, CA was most frequently employed, and it was
followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
Overall, signers used very few null markings for the introduction of a
referent and the majority of the existing ones were CA. Additionally,
signers used CA more for referent re-introduction than maintenance. We
observed a similar instances of agreement verbs for maintained and
re-introduced contexts. In contrast, plain verbs were favored more in
the maintenance setting than re-introduction.

\begin{table}
\small
\caption{Distribution of classifier referring expressions by discourse status}
\label{tab:my-table4}
\begin{tabular}{llll}
\hline
                        & WECL          & BPCL        & ExtCL         \\ \hline
\textit{Introduced:}    & 2.14\% (6)    & 2.4\% (3)   & 71.64\% (48) \\
\textit{Maintained:}    & 67.26\% (189) & 77.6\% (97) & 23.88\% (16) \\
\textit{Re-introduced:} & 30.6\% (86)   & 20\% (25)   & 4.48\% (3)   \\ \hline
\end{tabular}
\end{table}

Across the discourse statuses, there is also a difference between
participant groups in the use of null marking for reference tracking.
Native signers of TİD used it more (0.23) compared to the late signers
(0.16). Comparably, late signers used nominals more frequently (0.25)
than native signers (0.23). Figure 1 shows the numerical distribution of
the responses by RE type and Group.

\begin{table}
\small
\caption{Distribution of null referring expressions by discourse status}
\label{tab:my-table5}
\begin{tabular}{llll}
\hline
                        & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
\textit{Introduced:}    & 1.58\% (17)        & 0.28\% (1)    & 0.5\% (1)      \\
\textit{Maintained:}    & 71.64\% (773)      & 89.97\% (323) & 77.23\% (156)  \\
\textit{Re-introduced:} & 26.78\% (289)      & 9.75\% (35)   & 22.28\% (45)   \\ \hline
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{manuscript_v2_files/figure-latex/fig:count-plot-1.pdf}
\caption{Referent counts by Discourse Status (Introduction, Maintanence,
Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and
Participant Group (Native, Late)}
\end{figure}

For more inference on the data and having pairwise comparisons, we fit a
regression model to RE Type using Discourse Status and Group as
predictors. We used the brms package (Bürkner, 2018) in R, with a
categorical distribution on the responses and sum contrasts for the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late). As the overall number of produced
pronouns was very little (40 data points), we excluded pronominal use
from the following analyses. For the current analysis, we did not define
priors ourselves but used the preset uninformed priors. Figure
2\footnote{We ran another model where we used the Nativeness predictor
  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
  model did not converg properly. We provide the model results as a plot
  and table in the Appendix.} shows the posterior probability
distributions where the dot is the median estimate, the thick line is
the 50\% credible intervals, and the thin line is the 95\% credible
intervals. There are 3 responses we are considering in this model, and
they are Classifier, Nominative, and Null. Classifier response is the
level that other responses Nominative and Null are compared against.

\begin{figure}
\centering
\includegraphics{manuscript_v2_files/figure-latex/fig:model-gram-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Null) with the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late)}
\end{figure}

In a multinomial distribution, the model takes one of the responses as a
reference and compares all other responses to it. For example, the first
coefficient, Nom\_Intercept (nominal intercept), in Figure 2 is the
comparison of a nominal versus a classifier. It means that participants
produced more nominals than classifiers in general regardless of
participant group and discourse status. The second coefficient shows
that null markings were also used more than classifiers by the
participants in general, but not as much as nominals.

In the comparison of nominals and classifiers, being a native acquirer
of TİD did not have an effect. Nominals were used more when the
discourse status was introduction compared to re-introduction. Again,
the whole comparison is made against a classifier. In contrast to the
introduction status, nominals were used less in maintenance contexts
compared to the re-introduction contexts within the greater comparison
to classifier referents. There does not seem to be an interaction
between the introduction status of discourse and being a native acquirer
of TİD. However, there seems to be an interaction between maintaining a
referent and being a native acquirer of TİD. In that, there were fewer
nominal referents when it was a native acquirer of TİD and the discourse
context was maintenance.

Compared to classifiers, being a native acquirer of TİD increased the
odds of using null marking in general. Null constructions were used less
in introduction contexts than in re-introduction. This contrasts with
the nominals (4th coefficient) as they were used more in referent
introduction versus re-introduction. In contrast to introduction
contexts, null markers were used more in referent maintenance compared
to re-introduction, yet again within the greater comparison to the
classifiers. There seems to be an interaction between introducing a
referent and being a native acquirer of TİD. In that, the odds of using
null marking increases if the participant is a native acquirer of TİD
and the referent is used in an introduction context. There is another
interaction with regards to referent maintenance, contrasting the first
interaction. In that, the odds of using null markers decrease if the
participant is a native acquirer of TİD and the referent is maintained.
This does not mean that late signers used more null markers in
maintenance contexts. It means that the use of null markers did not
linearly increase as the addition of using a null marker instead of a
classifier as a native plus using null markers instead of a classifier
for maintaining a referent versus re-introducing it.

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

In our study, we aimed to (i) describe how signers track referents in
signed discourse in TİD and (ii) examine whether language deprivation
results in more over-redundant referent selection. To this end, we
investigated the distribution of referring expressions (REs) that the
native and late signers select when employing referents in different
discourse statuses (introduced, maintained, and re-introduced).

In our first set of hypotheses regarding referent tracking strategies,
we had the general expectation that signers would use more explicit
forms (mainly nominals) for referent introduction and less implicit
forms such as null marking, pronominal use, and classifiers. In
re-introduced contexts, we predicted that signers would use both
explicit and implicit markers depending on the accessibility of the
tracked referent.

For age of acquisition effects, based on what is proposed for child and
adult deaf signers with language deprivation in terms of their
morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
Cormier et al., 2013), we expected that late signers would be more
over-redundant in their referent tracking, akin to second language
learners (Bel et al., 2015; Frederiksen and Mayberry, 2019). That is, we
had the expectation that late signers would use more explicit markers
like nouns and fewer implicit markers like null marking even for highly
accessible referents (when the discourse status is maintained or even
re-introduced in same cases) than the control group (i.e., native
signers).

\hypertarget{tracking-referents-in-tid}{%
\subsection{Tracking referents in TİD}\label{tracking-referents-in-tid}}

The results of the present study partly support our hypotheses. In line
with our expectations, regardless of their acquisition groups (native or
late) (i) signers mainly used nominals to introduce a referent into the
discourse, (ii) used null marking and classifiers but not pronouns for
maintained contexts, and (iii) employed both nominals and less explicit
markers (classifiers, null markers and pronouns) to re-introduce a
referent.

However, the proportion of the observed discourse statuses (26\% for
introduction, 40\% for maintenance, 34\% for re-introduction across
acquisition groups) in this study differed from that of some other
studies. For a comparison, while the percentage of introduced referents
were similar in both research (26\% versus 24\%), the participants in
the present study brought back referents into the discourse more
frequently compared to the participants in Frederiksen and Mayberry
(2019) and Frederiksen and Mayberry (2016) which reported that
re-introduction contexts accounted for only 7\% of the referents in ASL.
On the other hand, maintenance percentages were lower in the present
study compared to Frederiksen and Mayberry (2016) (40\% versus 69\%).
Investigating the distribution of REs across discourse statuses in ASL,
a similar study by Czubek (2017) also reported contrasting numbers (17\%
for introduction, 64\% for maintenance, and 19\% for re-introduction).
We suggest that the discrepancy might be due to the nature of the
stimuli used in these research. In Frederiksen and Mayberry (2016), the
authors used 4-6-episode-long picture story with 2 animate characters
and 1 inanimate object. In contrast, the stimuli in the present study
consisted of 8-episode-long video clips on average with a mean number of
2 animate characters and 3 inanimate objects. Although the number of
animate characters and inanimate characters in Czubek (2017)'s stimuli
matched that of the present study, their narrative structure (a
6-episode-long picture story) was still shorter compared to our
narrative stimuli. We believe fewer possible referents and/or a simpler
narrative structure in these studies may have contributed to the lower
number of re-introduced contexts compared to the present study. Due to
the story length and composition in this study, the participants had to
track more recurring animate or inanimate referents for a longer period
of time. As a result, our signers engaged in referent re-introduction
more than what was reported in previous studies (Czubek, 2017;
Frederiksen and Mayberry, 2016).

As for the distribution of REs across different discourses statuses, we
observed that signers overwhelmingly preferred to use nominals over
other REs for the first mention of a referent and used null markers such
as CA, and agreement and plain verb for referents with high
accessibility (i.e., when the discourse status was maintained). This
finding parallels other accounts of referential accessibility suggested
for other sign languages {[}see Frederiksen and Mayberry (2016); Czubek
(2017); and Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al.
(2015); and Barberà and Massó (2009) for LSC. Our results also indicate
that modified nouns (e.g., nouns that are preceded or followed by a
classifier or IX) appeared more in introduced contexts compared to
re-introduced contexts. This brings along the assumption that modified
nouns are used more to track referents with lower accessibility,
probably due to their more explicit structure.

Interestingly and to our surprise, the signers in this study used
pronominal IX very sporadically with only 40 observed instances that
were mainly used to re-introduce referents rather than maintain them.
This finding is comparable with what has been reported for ASL (Czubek,
2017; Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC
(Bel et al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek,
2015), in which the signers extensively made overt pronominal references
(up to 30\%) to maintain entities. Comparable to our findings, It has
been reported that pronouns do not frequently occur in the local context
and instead follows a topic shift and thus rank higher than classifiers
and null markers different genre (see Barberà and Massó, 2009 for LSC;
and Nuhbalaoglu, 2018 for TİD). Considering this, the claim that the IX
sign functions like demonstratives that are also used to track referents
in re-introduced contexts has not been borne out (Koulidobrova and
Lillo-Martin, 2016). Similar to these research, our findings indicate
that the pronominal IX in TİD was used more concurrently with nominal
constructions and pronominal IX was used very infrequently in isolation.
The observation that overt pronouns are rare in ASL narratives comes
from studies that include a simple picture retelling task with one
protagonist (e.g., Frederiksen and Mayberry, 2016) as well as a more
complex narrative design as in (Czubek, 2017). However, Czubek (2017)'s
study also incorporated different genres and reported more instances of
overt pronouns in personal narratives (compared to picture or video
retelling). In fact, this finding suggests that the use of pronouns
might be genre-dependent. Moreover, another production study on referent
tracking in ASL reported object preference for pronominals (Frederiksen
and Mayberry, 2017). Nonetheless, we only coded subjects for maintained
and re-introduced referents. We argue that all of these factors
(genre-dependency as well as object preference) might contribute to the
explanation of why overt pronouns in this data are dispreferred in favor
of other markers for highly accessible referents such as null marking
and classifiers.

A deeper look into the allocation of classifier constructions revealed
that WECLs were the most common type of classifiers for the maintenance
of a referent followed by BPCLs and ExtCLs. We observed a similar
alignment for re-introduced contexts. In contrast, a great number of the
overall ExtCLs were used to introduce a new referent into the discourse.
That is to say, whereas WECLs were the preferred classifier types to
signal highly accessible referents, ExtCLs constructions were mainly
used for referents which were not easily accessible. We suggest that the
less iconic and transparent nature of WECLs allowed signers to continue
a previous referent without being redundant. As pointed out by Czubek
(2017), signers prefer to rely on ExtCLs for the first mention of an
entity because these constructions are informative and explicit by
nature, and they provide visual and descriptive cues about an entity and
its location in space. Nevertheless, we only observed 3 instances of
ExtCLs for re-introduced contexts, replicating Frederiksen and Mayberry
(2019)'s findings for ASL.

As for the distribution of the null markers, approximately 90\% of the
plain verbs were used to maintain a previous referent, making these
constructions least implicit REs used for TİD\footnote{As in other sign
  languages, in TİD, the lexical meaning of signs has an effect on the
  morphological grouping. For instance, usually (di)transitive agreement
  verbs denote some sort of physical or metaphorical transfer of an
  object or abstract notion (Meir 1999). As for plain verbs, both
  phonology of a sign and its semantics have an effect on morphology.
  For instance, \textsc{eat} in TİD is signed nearby the mouth and since
  it needs to remain in that phonological location, it is not suitable
  for agreement and thus it is regarded as a plain verb. Likewise the
  experiencer verb \textsc{love} is signed on the torso and as such it
  does not take part in agreement. These generalizations noted we should
  also be cautious in labelling a verb as categorially one type or the
  other. For instance, some intransitive plain verbs like \textsc{fall}
  or \textsc{jump} in TİD may display spatial person agreement with the
  single argument in the clause (Gökgöz and Sevgi, 2020). See also for a
  similar discussion in Brazilian Sign Language.\}.}. Agreement verbs
and CA were the following implicit markers of reference, 0.77 and 0.72
of which were used for maintained contexts, respectively. This contrasts
with some of the research (Czubek, 2017) in which agreement verbs ranked
lowest on scale, indicating that these constructions were more implicit
compared to plain verbs. However, there seems to be a common ground for
the position of CA markers which have been claimed to rank higher than
other null markers as in the present study (Barberà and Massó, 2009;
Czubek, 2017; Frederiksen and Mayberry, 2016). In Figure
\ref{fig:scale}, we report the distribution of REs in TİD as a function
of referent accessibility and compare our accessibility hierarchy with
that of (Frederiksen and Mayberry, 2016) for ASL. In the hierarchies,
the left most REs (e.g., modified and bare nouns for TİD) are used when
the accessibility is low. If the referent is already active in the
addressee's mind and its accessibility is high, then the right most
markers (e.g., agreement and plain verbs for TİD) are selected.

\begin{figure}
\caption{Comparing the present study and Frederikse and Mayberry (2016) in terms of proposed accessibility hierarchies}
\label{fig:scale}
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16) -- (214.92,115.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15) -- (314.89,115.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34) -- (414.89,115.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize M NOM}};
% Text Node
\draw (191,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize BARE NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node
\draw (285,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize WECL \& BPCL}};
% Text Node
\draw (359,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize CA}};
% Text Node
\draw (405,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR}};
% Text Node
\draw (452,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in the present study for TİD
    \end{minipage}

\vspace{1 cm}

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node% Text Node
\draw (333,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR \& CA}};
% Text Node
\draw (445,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in Frederiksen and Mayberry (2016) for ASL
    \end{minipage}
\end{figure}

Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals
and ExtCLs as markers of lowly accessibly referents. Likewise, in both
hierarchies, as markers of highly accessible referents, different types
of null markers are further distinguished. Plain verbs rank the lowest
on the scales, preceded by agreement verbs and constructed action. On a
similar ground, both scales lack pronouns, which occurred very
infrequently in the data. Although Frederiksen and Mayberry (2016) did
not include other classifier types in the hierarchy, we proposed that
WECLs and BPCLs must ranker higher than null markers but lower than
ExtCLs that were mainly used by interlocutors to introduce a referent
into the discourse context.

\hypertarget{age-of-acquisition-effects-on-referent-tracking}{%
\subsection{Age of acquisition effects on referent
tracking}\label{age-of-acquisition-effects-on-referent-tracking}}

The results in this study only somewhat support our second set of
hypotheses that we set for the effects of age of acquisition on
participants' RE selection. That is, while being a native acquirer of
TİD increased the odds of using null marking in reference tracking
across different discourse statuses, we failed to find such a difference
for overall nominal production. Taking a closer look at the interaction
terms, we observed that native signers, compared to late signers, were
less likely to use nominal constructions when the discourse status was
maintained. Native acquisition further increased the odds of null
markers when the discourse status was introduced. However, we found the
opposite alignment for the interaction of maintenance and native
acquisition. That is, the coefficient for this interaction term slightly
decreased the combined main effects of having a maintained context and
being a native signer. This means that these main effects were greater
than the interaction or added effects, pointing out to the implication
that the type of discourse status (introduced, maintained,
re-introduced) accounts for the distribution of REs in TİD narratives
better than age of acquisition (native, late) which, in comparison, had
a less salient effect.

While late signers were only partially over-redundant in maintained
contexts, they were more so in introduced contexts. In line with our
study, a tendency for limited over-redundancy has been suggested for
late signing children in DGS (Becker, 2009) and late signing adults in
BSL specifically for the use of constructed action (Cormier et al.,
2013). As for TİD, to our knowledge, there is only one study (Gür, 2018)
that directly aimed at examining age of acquisition effects on adult
signers' narrative skills. Gür (2018) observed that late signing deaf
adults and children set the scene less frequently and used constructed
action less accurately compared to their native counterparts. However,
this particular investigation did not present any data on the
distribution of REs between two acquisition groups. In the present
study, the finding that native signers preferred to use null arguments
even for the first mention of a referent (19 in total) is surprising
given some of the previous literature in the topicality and saliency
framework (Cormier et al., 2013; Swabey, 2002). These research reported
very little or no use of null marking by native signers for introduced
contexts.

Given the paucity of research directed at the late signing group, we now
turn to the referent tracking strategies observed for second language
learners of a sign language. Our study replicated some of the findings
of the line of research that investigated deaf or hearing bimodal
bilingual signers (Bel et al., 2015; Frederiksen and Mayberry,
2019).\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).}
The authors reported that more instances of null marking in the context
of referent introduction was attested for native signers than deaf or
hearing bimodal bilingual signers. Although using null marking to
introduce a referent into the discourse would normally result in
referential ambiguity, it is important to note that our stimuli
consisted of 10 interconnected narratives with 2 or more recurring
characters. It may be possible that native signers, having already
introduced the referent with an explicit nominal marker in a different
discourse context, found it felicitous to refer to the same entity with
more implicit markers in the following narratives even for introduced
contexts. Then, our results imply a difference in pragmatic sensitivity
to referential accessibility between native and late signers, which is
more robustly reflected in introduced contexts.

Since we only observed a small difference between the two acquisition
groups for the use of certain REs, the aforementioned findings allude to
the implication that the effect on language deprivation on pragmatic
factors should also be considered in different contexts. For one, both
groups of signers can be tested with different genres of narratives and
with a more complex story design. In order to capture possible variation
in the accessibility levels of re-introduced referents, a more
fine-tuned protocol (as in Toole, 1996) can be used for measuring
referent distance and the number of other possible competitors. Taken
altogether, our results accentuates the necessity of engaging in signed
narrative production early in life as well as the importance of early
intervention for late signing deaf children.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The present study aimed to investigate the reference tracking strategies
of native and late deaf adult signers in TİD narratives. In line with
our expectations, we found that the distribution of REs (nominal,
pronominal, classifier, and null marking) followed from previous
accounts of saliency and referential accessibility. For introduced
contexts, signers produced more nominals and ExtCLs, and to maintain a
previous referent, hey mainly used null arguments (plain verbs,
agreement verbs and constructed action) and certain classifiers (e.g.,
BPCL and WECL) to track highly accessible referents. We observed only
very few instances of pronouns (mostly produced by late signers), unlike
our expectations. The ones we observed were mostly reserved for
re-introduced contexts along with nominals. Based on our findings across
the two groups, we proposed a hierarchy of referent accessibility for
TİD. As for the age of acquisition effects, we observed limited
over-redundancy for signers with delayed language acquisition (i.e.,
late signers) who employed fewer implicit forms of reference tracking
compared to native signers. Native signers were less likely to use a
nominal for maintained contexts and more likely to use null marking for
introduced contexts than late signers. We conclude that native signers
might only partially adhere to pragmatic norms and leverage
accessibility better than late signers since they use less form when the
referent is more active in the mind.

\clearpage

\appendix

\hypertarget{participant-background-information}{%
\section{Participant background
information}\label{participant-background-information}}

\label{section:a}

\begin{table}[hbt!]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-2}Participant demographic information based on self-report}

\small{

\begin{tabular}{llllllll}
\toprule
No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{Deaf schools} & \multicolumn{1}{c}{Home language}\\
\midrule
2 & Native & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign\\
3 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign\\
13 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign\\
14 & Native & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign\\
18 & Native & M & 0-3 & 28 & 12 & PS, MS & TİD\\
20 & Native & M & 0-3 & 32 & 12 & PS, MS & TİD\\
22 & Native & M & 0-3 & 35 & 12 & PS, MS, HS & TİD\\
24 & Native & F & 0-3 & 23 & 12 & PS, MS, HS & TİD\\
26 & Native & F & 0-3 & 30 & 12 & PS, MS, HS & TİD\\
27 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
29 & Native & F & 0-3 & 24 & 14 & PS, MS, HS & TİD\\
30 & Native & M & 0-3 & 18 & 12 & PS, MS & TİD\\
33 & Native & F & 0-3 & 21 & 16 & PS, MS, HS & TİD\\
34 & Native & M & 0-3 & 19 & 16 & PS, MS, HS & TİD\\
38 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
10 & Late & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR\\
23 & Late & M & 4-7 & 36 & 12 & MS, HS & TR, home sign\\
4 & Late & M & 4-7 & 33 & 12 & PS, HS & TR\\
16 & Late & F & 4-7 & 33 & 12 & PS & TR, home sign\\
21 & Late & F & 4-7 & 31 & 14 & PS, MS & Home sign\\
25 & Late & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign\\
35 & Late & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR\\
36 & Late & M & 8-12 & 43 & 16 & PS, MS & TR\\
12 & Late & M & 8-12 & 28 & 16 & MS & TR\\
15 & Late & F & 8-12 & 24 & 12 & PS, MS, HS & TR, TİD\\
19 & Late & F & 8-12 & 24 & 8 & PS, MS & Home sign\\
28 & Late & M & 8-12 & 33 & 12 & PS, MS & Home sign\\
31 & Late & M & 13-17 & 50 & 8 & MS & TR, home sign\\
32 & Late & M & 13-17 & 31 & 16 & PS, MS, HS & TR\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\clearpage

\hypertarget{effect-of-narrating-to-the-camera-vs.-experimenter}{%
\section{Effect of narrating to the camera
vs.~experimenter}\label{effect-of-narrating-to-the-camera-vs.-experimenter}}

\label{section:b}

\begin{figure}
\centering
\includegraphics{manuscript_v2_files/figure-latex/fig:model-camera-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Null) with the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late)}
\end{figure}

\clearpage

\hypertarget{additional-model-with-3-levels-for-nativeness}{%
\section{Additional model with 3 levels for
Nativeness}\label{additional-model-with-3-levels-for-nativeness}}

\label{section:c}

\begin{table}[hbt!]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-4}Model results for Nativeness with 3 levels}

\small{

\begin{tabular}{llllll}
\toprule
Coefficients & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{EstError} & \multicolumn{1}{c}{lower95} & \multicolumn{1}{c}{upper95} & \multicolumn{1}{c}{Rhat}\\
\midrule
muNominal\_Intercept & 1.44 & 0.17 & 1.12 & 1.81 & 1.02\\
Null\_Intercept & -18.01 & 19.44 & -80.15 & -2.32 & 1.08\\
Nom\_Late1vsNative & 0.25 & 0.43 & -0.52 & 1.16 & 1.02\\
Nom\_Late2vsLate1 & -0.28 & 0.45 & -1.24 & 0.55 & 1.02\\
Nom\_Intro & 4.24 & 0.45 & 3.50 & 5.24 & 1.06\\
Nom\_Maintain & -5.33 & 0.30 & -5.97 & -4.79 & 1.05\\
Nom\_L1vsN*Intro & 2.56 & 1.29 & 0.49 & 5.68 & 1.07\\
Nom\_L2vsL1*Intro & -4.00 & 1.25 & -6.96 & -2.01 & 1.07\\
Nom\_L1vsN*Maintain & -0.33 & 0.83 & -2.17 & 1.13 & 1.06\\
Nom\_L2vsL1*Maintain & 0.47 & 0.83 & -1.01 & 2.29 & 1.07\\
Null\_Late1vsNative & -6.19 & 28.69 & -75.09 & 57.19 & 1.06\\
Null\_Late2vsLate1 & 10.31 & 26.94 & -34.28 & 88.61 & 1.06\\
Null\_Intro & -77.04 & 77.76 & -326.00 & -14.23 & 1.08\\
Null\_Maintain & 38.72 & 38.88 & 7.28 & 162.93 & 1.08\\
Null\_L1vsN*Intro & -25.29 & 114.78 & -300.34 & 227.46 & 1.06\\
Null\_L2vsL1*Intro & 40.13 & 107.78 & -137.50 & 352.79 & 1.06\\
Null\_L1vsN*Maintain & 12.85 & 57.38 & -114.06 & 150.55 & 1.06\\
Null\_L2vsL1*Maintain & -20.72 & 53.89 & -177.66 & 68.26 & 1.06\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Rhat values above 1 indicate that the regression model chains did not converge properly. There were 4 chains in our regression model.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\clearpage

\setlength{\columnsep}{10pt}
\twocolumn
\linespread{1}
\justifying
\footnotesize

\hypertarget{episodes-in-the-stimuli}{%
\section{Episodes in the stimuli}\label{episodes-in-the-stimuli}}

\label{section:b}

\textbf{Narration 1}

1. Tom chases Jerry, wearing a wooden plank

2. Tom and Jerry enter a tube

3. Tom falls down and Jerry runs away

4. Tube gets thinner and Tom chases Jerry

5. Jerry gets out and escapes

6. Tom follows Jerry in shape of a stick

\textbf{Narration 2}

1. Tom walks past a female cat sitting on a couch

2. Tom turns to and approaches female cat

3. Female cat blushes

4. Tom offers a fish in aquarium to female cat

5. Female cat rejects offer

6. Tom offers a bird in cage to female cat

7. Episode 5 is repeated

8. Tom leaves by walking in a flirtatious manner

\textbf{Narration 3}

1. Jerry falls down and tries to walk up stairs

2. Tom pulls carpet back

3. Jerry struggles

4. A piano falls down stairs

5. Jerry manages to escape

6. Tom sees piano and tries to run away

7. Piano smashes into Tom and flattens him

8. Tom falls on ground

\textbf{Narration 4}

1. Tom hides behind a cardboard house

2. Jerry walks to house with a toy mouse

3. Toy leaves Jerry who stops at entrance

4. Toy enters house

5. Toy gets smashed into pieces

6. Jerry is surprised

7. Tom hurts his stomach after swallowing toy

8. Tom looks at his broken teeth on a mirror

9. Tom gets angry and throws down mirror

\textbf{Narration 5}

1. Tom fakes sleeping and suddenly gets up

2. Tom pulls a rope inside a couch

3. Jerry appears holding a bar of soap

4. Jerry washes Tom's mouth with bar and flees

5. Tom spits soap out

6. Jerry escapes to nest with rope

7. Tom catches rope and stars pulling it

8. A mouse trap appears and catches Tom's nose

9. Tom yells and removes trap

\textbf{Narration 6}

1. Tom points a sword at Jerry and catches him

2. Tom looks at Jerry

3. Red spots appear on Jerry's face

4. Tom gets scared and drops Jerry

5. Tom rushes into bathroom

6. Tom washes his hands

7. Tom opens a drawer

8. Tom swallows pills one after another

9. Feeling ill, Tom looks out of window

10. Tom and Jerry look at each other

\textbf{Narration 7}

1. Two male cats chase one another

2. Tom sees a female cat and kisses her

3. Tom hides away when black cat returns

4. Tom pokes his head behind couch again

5. Episode 3 is repeated

6. black cat sees female cat and hugs her

7. Tom appears and black cat kisses Tom

8. Two male cats move away from one another

\textbf{Narration 8}

1. Jerry hits Tom's knee with a hammer

2. Tom suddenly jumps up

3. Tom yells and holds his knee

4. Jerry appraches Tom holding a thermometer

5. Jerry sticks thermometer into Tom's mouth

6. Jerry lights a fire under thermometer

7. Red liquid starts to rise up

8. Thermometer explodes

\textbf{Narration 9}

1. Tom sits on Jerry who tries to get loose

2. Jerry looks around and sees a hat ribbon

3. Jerry finds a needle attached to hat

4. Jerry takes needle and sticks Tom with it

5. Tom turns red and jumps

6. Jerry placesn eedle between two cushions

7. Tom dodges needle

\textbf{Narration 10}

1. Wearing a stethoscope, Tom catches Jerry

2. Jerry yells into stethoscope

3. Tom's head pops up

4. Jerry runs and enter a mouse nest

5. Tom inserts a gun into nest

6. Gun bends over toward Tom

7. Tom fires gun and his head explodes \clearpage

\onecolumn

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-ahn2019}{}}%
Ahn, Dorothy, 2019. The determinacy scale: A competition mechanism for
anaphoric expressions (PhD thesis).

\leavevmode\vadjust pre{\hypertarget{ref-ahrenholz2005}{}}%
Ahrenholz, Bernt, 2005. Reference to persons and objects in the function
of subject in learner varieties. The structure of learner varieties
1964.

\leavevmode\vadjust pre{\hypertarget{ref-aikhenvald2017}{}}%
Aikhenvald, Alexandra, 2017. A typology of noun categorization devices.

\leavevmode\vadjust pre{\hypertarget{ref-aksu-kouxe72015}{}}%
Aksu-Koç, Ayhan, Nicolopoulou, Ageliki, 2015. Character reference in
young children's narratives: A crosslinguistic comparison of English,
Greek, and Turkish. Lingua, Reference and referentiality in native and
learner grammars 155, 62--84.
doi:\href{https://doi.org/10.1016/j.lingua.2014.04.006}{10.1016/j.lingua.2014.04.006}

\leavevmode\vadjust pre{\hypertarget{ref-ariel1990}{}}%
Ariel, Mira, 1990. Accessing Noun-Phrase Antecedents. Routledge.

\leavevmode\vadjust pre{\hypertarget{ref-azar2015}{}}%
Azar, Zeynep, Özyürek, Aslı, 2015. Discourse management: Reference
tracking in speech and gesture in Turkish narratives. Dutch Journal of
Applied Linguistics 4, 222--240.
doi:\href{https://doi.org/10.1075/dujal.4.2.06aza}{10.1075/dujal.4.2.06aza}

\leavevmode\vadjust pre{\hypertarget{ref-bamberg1997}{}}%
Bamberg, Michael GW, 1997. Positioning between structure and
performance. Journal of narrative and life history 7, 335342.

\leavevmode\vadjust pre{\hypertarget{ref-barberuxe02009}{}}%
Barberà, Gemma, Massó, Guillem, 2009. Keeping track in sign language:
Coreference annotation using centering theory, in: Talk Presented at the
Sign Language Corpora: Linguistic Issues Workshop, University College
London.

\leavevmode\vadjust pre{\hypertarget{ref-becker2009}{}}%
Becker, Claudia, 2009. Narrative competences of Deaf children in German
Sign Language. Sign Language \& Linguistics 12, 113--160.
doi:\href{https://doi.org/10.1075/sll.12.2.02bec}{10.1075/sll.12.2.02bec}

\leavevmode\vadjust pre{\hypertarget{ref-bel2015}{}}%
Bel, Aurora, Ortells, Marta, Morgan, Gary, 2015. Reference control in
the narratives of adult sign language learners. International Journal of
Bilingualism 19, 608--624.
doi:\href{https://doi.org/10.1177/1367006914527186}{10.1177/1367006914527186}

\leavevmode\vadjust pre{\hypertarget{ref-benedicto2004}{}}%
Benedicto, Elena, Brentari, Diane, 2004. Where did all the arguments
go?: Argument-changing properties of classifiers in ASL. Natural
Language \& Linguistic Theory 22, 743810.

\leavevmode\vadjust pre{\hypertarget{ref-berman2013relating}{}}%
Berman, Ruth A, Slobin, Dan Isaac, 2013. Relating events in narrative: A
crosslinguistic developmental study. Psychology Press.

\leavevmode\vadjust pre{\hypertarget{ref-bosch2007}{}}%
Bosch, Peter, Umbach, Carla, 2007. Reference determination for
demonstrative pronouns. ZAS Papers in Linguistics 48, 39--51.
doi:\href{https://doi.org/10.21248/zaspil.48.2007.353}{10.21248/zaspil.48.2007.353}

\leavevmode\vadjust pre{\hypertarget{ref-boudreault2006}{}}%
Boudreault, Patrick, Mayberry, Rachel I., 2006. Grammatical processing
in american sign language: Age of first-language acquisition effects in
relation to syntactic structure. Language and Cognitive Processes 21,
608--635.
doi:\href{https://doi.org/10.1080/01690960500139363}{10.1080/01690960500139363}

\leavevmode\vadjust pre{\hypertarget{ref-buxfcrkner2018}{}}%
Bürkner, Paul-Christian, 2018. Advanced bayesian multilevel modeling
with the R package brms. The R Journal 10, 395.
doi:\href{https://doi.org/10.32614/RJ-2018-017}{10.32614/RJ-2018-017}

\leavevmode\vadjust pre{\hypertarget{ref-chafe1976}{}}%
Chafe, W., 1976.
\href{https:///paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}{Givenness,
contrastiveness, definiteness, subjects, topics, and point of view}.

\leavevmode\vadjust pre{\hypertarget{ref-chamberlain_language_1999}{}}%
Chamberlain, Charlene, Morford, Jill P., Mayberry, Rachel I. (Eds.),
1999. Language acquisition by eye. Psychology Press, Mahwah, N.J.

\leavevmode\vadjust pre{\hypertarget{ref-cheng2021}{}}%
Cheng, Qi, Mayberry, Rachel I., 2021. When event knowledge overrides
word order in sentence comprehension: Learning a first language after
childhood. Developmental Science.
doi:\href{https://doi.org/10.1111/desc.13073}{10.1111/desc.13073}

\leavevmode\vadjust pre{\hypertarget{ref-cormier2012}{}}%
Cormier, Kearsy, Schembri, Adam, Vinson, David, Orfanidou, Eleni, 2012.
First language acquisition differs from second language acquisition in
prelingually deaf signers: Evidence from sensitivity to grammaticality
judgement in British Sign Language. Cognition 124, 50--65.
doi:\href{https://doi.org/10.1016/j.cognition.2012.04.003}{10.1016/j.cognition.2012.04.003}

\leavevmode\vadjust pre{\hypertarget{ref-cormier2013}{}}%
Cormier, Kearsy, Smith, Sandra, Zwets, Martine, 2013. Framing
constructed action in British Sign Language narratives. Journal of
Pragmatics 55, 119--139.
doi:\href{https://doi.org/10.1016/j.pragma.2013.06.002}{10.1016/j.pragma.2013.06.002}

\leavevmode\vadjust pre{\hypertarget{ref-crasborn2008}{}}%
Crasborn, O., Sloetjes, H., 2008.
\href{https://www.semanticscholar.org/paper/Enhanced-ELAN-functionality-for-sign-language-Crasborn-Sloetjes/e1623ec6b2bc3532129a3682eb263cc2e7d5389c}{Enhanced
ELAN functionality for sign language corpora}. undefined.

\leavevmode\vadjust pre{\hypertarget{ref-czubek2017}{}}%
Czubek, Todd Alan, 2017. A comprehensive study of referring expressions
in ASL 244.

\leavevmode\vadjust pre{\hypertarget{ref-debreslioska2013}{}}%
Debreslioska, Sandra, Özyürek, Asli, Gullberg, Marianne, Perniss,
Pamela, 2013. Gestural viewpoint signals referent accessibility.
Discourse Processes 50, 431--456.
doi:\href{https://doi.org/10.1080/0163853X.2013.824286}{10.1080/0163853X.2013.824286}

\leavevmode\vadjust pre{\hypertarget{ref-dikyuva2015}{}}%
Dikyuva, Hasan, Makaroğlu, Bahtiyar, Arık, Engin, 2015. Türk i{ş}aret
dili dilbilgisi kitab{ı}. Aile ve Sosyal Politikalar Bakanl{ı}{ğ}{ı}
Yay{ı}nlar{ı}: Ankara.

\leavevmode\vadjust pre{\hypertarget{ref-dudis2004}{}}%
Dudis, Paul G., 2004. Body partitioning and real-space blends. Cognitive
Linguistics 15.
doi:\href{https://doi.org/10.1515/cogl.2004.009}{10.1515/cogl.2004.009}

\leavevmode\vadjust pre{\hypertarget{ref-emmorey1996}{}}%
Emmorey, Karen, 1996. The confluence of space and language in signed
languages, in: Language, Speech, and Communication. The MIT Press,
Cambridge, MA, US, pp. 171--209.

\leavevmode\vadjust pre{\hypertarget{ref-engberg-pedersen1993}{}}%
Engberg-Pedersen, Elisabeth, 1993. Space in danish sign language: The
semantics and morphosyntax of the use of space in a visual language.

\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2016}{}}%
Frederiksen, Anne Therese, Mayberry, Rachel I., 2016. Who's on First?
Investigating the referential hierarchy in simple native ASL narratives.
Lingua 180, 49--68.
doi:\href{https://doi.org/10.1016/j.lingua.2016.03.007}{10.1016/j.lingua.2016.03.007}

\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2017}{}}%
Frederiksen, Anne Therese, Mayberry, Rachel I., 2017. Pronominalizing
subjects: Referentially empty pronouns in ASL.

\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2019}{}}%
Frederiksen, Anne Therese, Mayberry, Rachel I., 2019. Reference tracking
in early stages of different modality L2 acquisition: Limited
over-explicitness in novice ASL signers{'} referring expressions. Second
Language Research 35, 253--283.
doi:\href{https://doi.org/10.1177/0267658317750220}{10.1177/0267658317750220}

\leavevmode\vadjust pre{\hypertarget{ref-givon1983}{}}%
Givon, T., 1983. \href{https://benjamins.com/catalog/tsl.3}{Topic
continuity in discourse}. John Benjamins Publishing Company.

\leavevmode\vadjust pre{\hypertarget{ref-grice1975}{}}%
Grice, Herbert P., 1975. Logic and conversation. Brill, p. 4158.

\leavevmode\vadjust pre{\hypertarget{ref-gullberg2006}{}}%
Gullberg, Marianne, 2006. Handling Discourse: Gestures, Reference
Tracking, and Communication Strategies in Early L2: Language Learning
Vol. 56, No. 1. Language Learning 56, 155--196.
doi:\href{https://doi.org/10.1111/j.0023-8333.2006.00344.x}{10.1111/j.0023-8333.2006.00344.x}

\leavevmode\vadjust pre{\hypertarget{ref-gundel1993}{}}%
Gundel, Jeanette K., Hedberg, Nancy, Zacharski, Ron, 1993. Cognitive
status and the form of referring expressions in discourse. Language 69,
274--307. doi:\href{https://doi.org/10.2307/416535}{10.2307/416535}

\leavevmode\vadjust pre{\hypertarget{ref-guxfcr2018}{}}%
Gür, Cansu, 2018. The Effects of Late Language Acquisition on the
Development of Narrative Skills: Insights from Turkish Sign Language
(PhD thesis). Erzurum.

\leavevmode\vadjust pre{\hypertarget{ref-hickmann1996}{}}%
Hickmann, Maya, Hendriks, Henriëtte, Roland, Françoise, Liang, James,
1996. The marking of new information in children's narratives: a
comparison of English, French, German and Mandarin Chinese*. Journal of
Child Language 23, 591--619.
doi:\href{https://doi.org/10.1017/S0305000900008965}{10.1017/S0305000900008965}

\leavevmode\vadjust pre{\hypertarget{ref-karaduxf6ller2017}{}}%
Karadöller, Dilay Z, Sümer, Beyza, Özyürek, Aslı, 2017. Effects of
Delayed Language Exposure on Spatial Language Acquisition by Signing
Children and Adults 5.

\leavevmode\vadjust pre{\hypertarget{ref-kayaba2022}{}}%
Kayabaşı, Demet, Gökgöz, Kadir, 2022. The causative-inchoative
alternation and age-of-acquisition effects on multi-predicate
constructions in turkish sign language. Language Learning and
Development 0, 1--26.
doi:\href{https://doi.org/10.1080/15475441.2022.2047690}{10.1080/15475441.2022.2047690}

\leavevmode\vadjust pre{\hypertarget{ref-koulidobrova2016}{}}%
Koulidobrova, Elena, Lillo-Martin, Diane, 2016. A {`}point{'}of inquiry:
The case of the (non-) pronominal IX in ASL. The impact of pronominal
form on interpretation 221, 250.

\leavevmode\vadjust pre{\hypertarget{ref-kruschke2011}{}}%
Kruschke, John K., 2011. Bayesian assessment of null values via
parameter estimation and model comparison. Perspectives on Psychological
Science 6, 299312.

\leavevmode\vadjust pre{\hypertarget{ref-kubu2008}{}}%
Kubuş, Okan, 2008. AN ANALYSIS OF TURKISH SIGN LANGUAGE (T{İ}D)
PHONOLOGY AND MORPHOLOGY (PhD thesis). Ankara.

\leavevmode\vadjust pre{\hypertarget{ref-lillo-martin1990}{}}%
Lillo-Martin, Diane, Klima, Edward S., 1990. Pointing out differences:
ASL pronouns in syntactic theory. Theoretical issues in sign language
research, Volume 1: Linguistics 191.

\leavevmode\vadjust pre{\hypertarget{ref-loew1984}{}}%
Loew, Ruth Carolyn, 1984.
\href{https:///paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}{Roles
and reference in American sign language : a developmental perspective}.

\leavevmode\vadjust pre{\hypertarget{ref-maclaughlin1997}{}}%
Maclaughlin, Dawn, 1997.
\href{http://lear.unive.it/jspui/handle/11707/7016}{The Structure of
Determiner Phrases: Evidence From American Sign Language.} (PhD thesis).

\leavevmode\vadjust pre{\hypertarget{ref-meier1990}{}}%
Meier, Richard P., Newport, Elissa L., 1990. Out of the hands of babes:
On a possible sign advantage in language acquisition. Language 66, 1.
doi:\href{https://doi.org/10.2307/415277}{10.2307/415277}

\leavevmode\vadjust pre{\hypertarget{ref-morgan2005}{}}%
Morgan, Gary, 2005.
\href{http://www.oxfordscholarship.com/view/10.1093/acprof:oso/9780195180947.001.0001/acprof-9780195180947}{The
Development of Narrative Skills in British Sign Language}, in: Schick,
Brenda, Marschark, Marc, Spencer, Patricia Elizabeth (Eds.),. Oxford
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-newport1990}{}}%
Newport, Elissa L., 1990. Maturational Constraints on Language Learning.
Cognitive Science 14, 11--28.
doi:\href{https://doi.org/10.1207/s15516709cog1401_2}{10.1207/s15516709cog1401\_2}

\leavevmode\vadjust pre{\hypertarget{ref-nuhbalaoglu2018}{}}%
Nuhbalaoglu, Derya, 2018. Comprehension and production of referential
expressions in German Sign Language and Turkish Sign Language: An
empirical approach (PhD thesis). Germany.

\leavevmode\vadjust pre{\hypertarget{ref-padden1986}{}}%
Padden, Carol, 1986. Verbs and role-shifting in american sign language.
National Association of the Deaf Silver Spring, MD, p. 57.

\leavevmode\vadjust pre{\hypertarget{ref-perniss2015}{}}%
Perniss, Pamela, Özyürek, Asli, 2015. Visible Cohesion: A Comparison of
Reference Tracking in Sign, Speech, and Co-Speech Gesture. Topics in
Cognitive Science 7, 36--60.
doi:\href{https://doi.org/10.1111/tops.12122}{10.1111/tops.12122}

\leavevmode\vadjust pre{\hypertarget{ref-prince1981}{}}%
Prince, EF, 1981. Toward a taxonomy of given - new information, in:
Cole, Peter (Ed.),. Academic Press, New York, pp. 223--255.

\leavevmode\vadjust pre{\hypertarget{ref-rathmann2007}{}}%
Rathmann, Christian, Mann, Wolfgang, Morgan, Gary, 2007. Narrative
Structure and Narrative Development in Deaf Children. Deafness \&
Education International 9, 187--196.
doi:\href{https://doi.org/10.1179/146431507790559932}{10.1179/146431507790559932}

\leavevmode\vadjust pre{\hypertarget{ref-sevinuxe72006}{}}%
Sevinç, Ayça Müge, 2006. Grammatical relations and word order in turkish
sign language (t{İ}d). Unpublished master's thesis, Middle East
Technical University, Turkey.

\leavevmode\vadjust pre{\hypertarget{ref-sorace2011}{}}%
Sorace, Antonella, 2011. Pinning down the concept of {``}interface{''}
in bilingualism. Linguistic approaches to bilingualism 1, 133.

\leavevmode\vadjust pre{\hypertarget{ref-sorace2009}{}}%
Sorace, Antonella, Serratrice, Ludovica, 2009.
\href{https://doi.org/10.1177/1367006909339810}{Internal and external
interfaces in bilingual language development: Beyond structural
overlap}. International Journal of Bilingualism 13, 195--210.

\leavevmode\vadjust pre{\hypertarget{ref-supalla1986}{}}%
Supalla, Ted, 1986. The classifier system in american sign language.
Noun classes and categorization 7, 181214.

\leavevmode\vadjust pre{\hypertarget{ref-supalla1990}{}}%
Supalla, Ted, 1990. Serial verbs of motion in ASL. Theoretical issues in
sign language research 1, 127152.

\leavevmode\vadjust pre{\hypertarget{ref-swabey2002}{}}%
Swabey, Laurie Anne, 2002. The Cognitive Status, Form and Distribution
of Referring Expressions in ASL and English Narratives (PhD thesis).
Minnepolis, USA.

\leavevmode\vadjust pre{\hypertarget{ref-talmy2003}{}}%
Talmy, Leonard, 2003. The representation of spatial structure in spoken
and signed language. Perspectives on classifier constructions in sign
language 169195.

\leavevmode\vadjust pre{\hypertarget{ref-toole1996}{}}%
Toole, Janine, 1996. The effect of genre on referential choice. p. 263.
doi:\href{https://doi.org/10.1075/pbns.38.16too}{10.1075/pbns.38.16too}

\leavevmode\vadjust pre{\hypertarget{ref-wagenmakers2007}{}}%
Wagenmakers, Eric-Jan, 2007. A practical solution to the pervasive
problems of p values. Psychonomic bulletin \& review 14, 779804.
doi:\href{https://doi.org/10.3758/bf03194105}{10.3758/bf03194105}

\leavevmode\vadjust pre{\hypertarget{ref-williams1988}{}}%
Williams, Jessica, 1988. Zero Anaphora in Second Language Acquisition: A
Comparison among Three Varieties of English. Studies in Second Language
Acquisition 10, 339--370.
doi:\href{https://doi.org/10.1017/S0272263100007488}{10.1017/S0272263100007488}

\leavevmode\vadjust pre{\hypertarget{ref-woll2013}{}}%
Woll, Bencie, 2013. Sign language and spoken language development in
young children: Measuring vocabulary by means of the CDI, in: Meurant,
Laurence, Sinte, Aurélie, Van Herreweghe, Mieke, Vermeerbergen, Myriam
(Eds.),. DE GRUYTER, Berlin, Boston, pp. 15--34.
doi:\href{https://doi.org/10.1515/9781614511472.15}{10.1515/9781614511472.15}

\end{CSLReferences}

\end{document}