### 2022-08-05 manuscript_v2.Rmd
updated

---
editor_options:
  chunk_output_type: console
title: Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language
abstract: |
  We investigated the reference tracking strategies among native and late adult deaf signers of Turkish Sign Language. Consistent with the theories of saliency and referential accessibility, regardless of their acquisiton groups, signers mainly used nominals and extension classifiers for introducing referents. To maintain a referent in the previous clause, signers chose null marking (e.g., constructed action, agreement and plain verbs). For topic shifts or re-introduced contexts, nominals and pronouns were chiefly favored although we observed very little pronominal use. As for age of acquisition effects, we only report limited over-redundancy for late signers who used null marking less compared to native signers especially for introduced and maintained contexts. Given our limited support for redundancy by late signers, we conclude that native signers might utilize the spatial affordances of the visual-spatial modality better than late signers only to a certain extent.
keywords: "Referential accessibility, signed narratives, pragmatic competence, delayed language acquisition, Turkish Sign Language"
journal: "Journal of Pragmatics"
date: "`r Sys.Date()`"
bibliography: ./bibliography/references.bib
times: true
numbersections: true
classoption: "review"
csl: ./bibliography/elsevier-harvard.csl
output: rticles::elsevier_article
header-includes:
   - \usepackage{floatrow}
   - \usepackage{hyperref}
   - \usepackage{enumitem}  
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \floatplacement{table}{H}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{times}
   - \usepackage{smartdiagram}
   - \usepackage{ragged2e}
   - \usepackage{tikz}
   - \usepackage{xcolor}
   - \usepackage[utf8]{inputenc}
   - \usepackage[T1]{fontenc}
   - \usepackage{graphicx}
   - \usepackage{./sty/langsci-gb4e}
   - \usepackage{caption}
   - \usepackage{multicol}
   - \setlength{\columnsep}{-3cm}
   - \usepackage{setspace}
---

# Introduction

Narratives are extensive discourses that construct an important part of human communication. In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities [@ariel1990; @givon1983; @gundel1993]: Certain forms such as nominal expressions are commonly used for the first mention of a referent in a discourse, and other forms such as pronouns and null marking may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse. The selection of these referring expressions (henceforth REs) is contingent on cognitive abilities, such that individuals must rely on the activation of the working memory to keep in mind previously referred expressions and long strings of information [@morgan2005; @bamberg1997]. Pragmatic abilities too play a pivotal role in the tracking of referents. The form with the least amount of redundant information must be selected based on the accessibility levels of the referent (i.e., how active a referent is in the addressee's mind given a point in discourse) [@ahn2019; @ariel1990]. Because of these cognitive and pragmatic necessities, it is not surprising to observe that producing a coherent narrative is one of the latest milestones in language acquisition among children [@hickmann1996]. As can be expected, there is a plethora of research examining how narrative competence is achieved in both monolingual and bilingual acquisition contexts among children and adults. Although narrative competence and referent selection have been studied extensively for typically developing speakers [@aksu-koc2015; @hickmann1996; @williams1988], deaf individuals who acquired a sign language in atypical learning contexts (i.e., deaf signers with linguistic deprivation) have remained mostly understudied with an exception of handful of studies [@cormier2013; @gur2018; @becker2009]. Nevertheless, it is paramount that we investigate and describe the effects of language deprivation among deaf adults on narrative competence more in depth. This will help us understand whether the loss of systematic and frequent language input in the early years of life affect how deaf signers produce a narrative in their adulthood. Given this, the present study will focus on deaf adult signers with language deprivation and examine their reference tracking strategies in signed narratives.

Sign Languages are natural languages of the Deaf communities all around the world. They are independent languages, which do not derive from majority spoken language(s) but emerge naturally when deaf people form a community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and has been the officially recognized language of the Deaf community in Turkey since 2005 [see @nuhbalaoglu2018 for a brief review of the historical development of the language]. Deaf or hearing children acquiring a SL within a family where there is one or more deaf caregivers display the same developmental milestones as hearing children acquiring a spoken language in a hearing family [@chamberlain_language_1999; @meier1990]. However, reports indicate that over 90% of all deaf individuals are born into hearing families who do not know how to sign (i.e., those with language deprivation) [@woll2013]. This means that the auditory modality used in the household is not accessible to many deaf children. That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school. First language deprivation or delay has been found to have negative effects on certain cognitive and linguistic measures among late signing children. To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic production and processing [@kayabasi2020; @boudreault2006; @newport1990; @cheng2021; @karadoller2017; @cormier2012] and narrative abilities [e.g., @becker2009]. In our overall research, we are concerned with the effects of this late exposure in the ultimate adult forms of TİD users.

## Referring Expression (RE) Selection

Natural languages can be systematically used to keep track of previous referents in the discourse and mark the changing cognitive statuses of the REs (i.e., from being active to less active in the addressee's mind) to create a more coherent narrative. Among the theoretical frameworks that aim to explain that different linguistic forms are selected in accordance with the addressee's mental representations of the referents in the discourse is the Givenness Hierarchy [@gundel1993]. According to this model, given or familiar information is expressed in more attenuated forms such as pronominal use and null marking since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information [@chafe1976]. The Givenness Hierarchy is akin to other previously proposed models of reference control within the topicality and saliency paradigm such as the Accessibility Hierarchy [@ariel1990], Topic Continuity [@givon1983], and Familiarity Scale [@prince1981]. @gundel1993 argued that the distribution of forms is not random but abides by the Gricean Maxim of Quantity [@grice1975]. That is to say, individuals must make the most informative and succinct choice possible to retain economy of form. When multiple expressions are available for the same referent as in an utterance like (\ref{ex:1}), phonetically simplest form is selected with the least redundant information are selected based on an economy principle. In the utterance, REs for "Eddie" are given in bold. All three REs (nominal, pronominal, and null marking) refer to the same entity; however, the quantity of marking decreases as the referent's accessibility increases. On this ground, for highly accessible referents (i.e., those that are highly active in the mind), linguistic forms with the least phonetic content are selected for the least amount of processing effort.

```{=tex}
\begin{exe}
    \ex \label{ex:1}
    So \textbf{Eddie} turned around. \hfill Williams (1988; p. 343, ex. 7) \\
\textbf{He} said, "youse got a problem?" \\
"Yeah, we want you," they say.
So-$\pmb{\emptyset}$ walked right up to them \\
\end{exe}
```
It is also possible to categorize REs into three according to their discourse status [@gullberg2006]. A referent can be introduced or mentioned for the first time, maintained, that is, continued across two or more clauses, or re-introduced into the discourse following a topic shift. As a result, interlocutors mark some REs differently than others based on the principles of reference accessibility. For spoken languages, this observation is well established [@ariel1990; @givon1983; @gundel1993; @debreslioska2013; @ahrenholz2005; @chafe1976; @hickmann1996; @azar2015]. These research indicate that the introduction and re-introduction of a referent are often accomplished by full noun phrases which can be commonly modified by an adjective or demonstrative. Maintenance, on the other hand, is typically marked with pronominals and null markers since the continued referent is still highly accessible to the addressee.

However, this particular observation is not peculiar to spoken languages. There is an emerging body of literature that reports the distribution of REs in signed narratives. Building on @gundel1993's work, @swabey2002 demonstrated that, like other spoken languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and re-introduced discourse contexts whereas null marking is used for referents which are "in-focus" or very accessible to the addressee. Overt REs like pronouns or nouns are disprefered for such referents. @swabey2002 concluded that "native" signers in her study abided by the Gricean maxim of quantity and were thus economic in their reference tracking. Similar findings have been reported for German Sign Language (Deutsche Gebärdensprache - DGS) [@perniss2015] and British Sign Language (BSL) [@morgan2005]. It is important to note, however, that these research either collapsed native and late signers into a single group and labeled them as "native-like" or explicitly compared native deaf signers with late deaf signers (with language deprivation) in terms of narrative competence.

### Signed vs. Spoken REs

Commonly in both spoken and sign languages nouns and noun phrases are used as REs. Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990]. In addition, the pointing sign is used to referentially anchor a referent to a location in signing space. For instance, when the signer introduces \textsc{elvan} for the first time, s/he can assign her to the right. This referential locus can later be used by agreement verbs and classifier predicates (more below). For instance, in (\ref{ex:2a}) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "a" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "b" subscript. When the signer makes a contrast in (\ref{ex:2b}), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}. We schematically represents the referential locus assigned to the referents in (\ref{ex:3}).

```{=tex}
\begin{exe}
    \ex 
\begin{xlist}
    \ex \label{ex:2a}
    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
    \\
    \ex \label{ex:2b}
    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
    \end{xlist}
    \end{exe}

    \begin{exe}
    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Curve Lines [id:da7006577339157659] 
\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
%Straight Lines [id:da8428827309782937] 
\draw    (112.9,128.8) -- (210.21,128.8) ;
%Curve Lines [id:da39290455666055124] 
\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
%Curve Lines [id:da9909903731270634] 
\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;

% Text Node
\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
% Text Node
\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
% Text Node
\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
% Text Node
\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
% Text Node
\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
\end{tikzpicture}
    \end{exe}
```
Grammatical gender as in Romance languages or noun classes as in Bantu languages are some reference tracking tools that are familiar from spoken languages. A somehow less known reference-tracking tool in spoken languages is *verbal classifiers*, which morphologically attach to the verb and often cross-reference a core argument in a clause [@aikhenvald2017]. Classifiers denote sets of objects based on semantic criteria such as size, shape, body-part, texture and sometimes function. Usually, the verbal classifier and the cross-referenced noun co-occur in the same clause when the argument is introduced into discourse. The classifier can track the argument on its own on subsequent occurrences. In sign languages, a verbal classifier morpheme is encoded by a *handshape* that is articulated phonologically co-temporally with a movement, which encodes the verbal meaning [@supalla1990; @benedicto2004]. These two infused morphemes form a complex predicate. Some classifier handshapes and the set of objects that they denote are below. Some classifier handshapes and the set of objects that they denote are presented below in (\ref{ex:4}).

```{=tex}
\begin{exe}
\ex \label{ex:4}
\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
\end{exe}
```
In (\ref{ex:5}), in the first clause, a referential locus, i.e. locus-a, is assigned to the locative source argument \textsc{cage}. In the second clause, the core argument \textsc{bird} is cross-referenced with the classifier handshape, i.e. small-legged animal. It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-b).

```{=tex}
\begin{exe}
    \ex \label{ex:5}
\begin{xlist}
    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
    \textsc{garden bird cage-a open} \\
    ‘In the garden, there is a cage and it is open.’
    \\

    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
    \textsc{bird fly[manner]-a tree-b land. CL-b}\\
    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
    \end{xlist}
    \end{exe}
    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
```
## Linguistic Tools for Tracking Referents in Signed Narratives

The linguistic inventory for discourse referents can include nominal constructions (bare nouns, fingerspelled nouns, and modified nouns), null markers (plain verb, agreement verb, and constructed action), classifier predicates (whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX) [@frederiksen2016]. This interim section describes such constructions attested for sign languages under four main headings (nominals, pronouns, classifiers, and null marking).

\clearpage

```{=tex}
\begin{multicols}{2}
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:6a} Full noun \\\glll
\includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
\ex \label{ex:6b} Noun modified by classifier \\\glll
\includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
\ex \label{ex:6c} Fingerspelled noun\\\glll 
\includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
\ex \label{ex:6d} Noun modified by index (\textsc{IX})\\\glll 
\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
\end{xlist}
\end{exe}
\end{multicols}
\ea 
\ea \label{ex:7} Pronominal \\\glll
\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
\z \z
\begin{multicols}{4}
\begin{exe}
\ex \label{ex:8}
\begin{xlist}
\ex Whole entity \\\glll 
\includegraphics[width=50pt]{pictures/3e.png} \\ \textsc{fish.fall}\\ \\
\ex Body part \\\glll
\includegraphics[width=50pt]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
\ex Handling \\\glll 
\includegraphics[width=50pt]{pictures/3c.png} \\ \textsc{hold} \\ \\
\ex Extension \\\glll 
\includegraphics[width=50pt]{pictures/3d_1.png} \includegraphics[width=50pt]{pictures/3a.png}\\ \textsc{plank} {\textsc{dh: mouse}} \\ {} {\textsc{ndh: tunnel}}\\ \\
\end{xlist}
\end{exe}
\end{multicols}
\begin{multicols}{3}
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:9a} Constructed action \\\glll 
\includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\

\ex \label{ex:9b} Plain verb  \\\glll
\includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\

\ex \label{ex:9c} Agreement verb  \\\glll 
\includegraphics[width=50pt]{pictures/4b_1.png} 
\includegraphics[width=50pt]{pictures/4b_2.png} \\ {\textsc{$_{1}$hit$_{3}$}} {} \\ \\ 
\end{xlist}
\end{exe}
\end{multicols}
\footnotetext{We use small capitals to indicate sign glosses.}
```
### Nominals

In TİD and many other sign languages, it is possible to use bare noun phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in (\ref{ex:6c}). Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubus2008]. Nouns can also be modified by classifiers as in (\ref{ex:6b}) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (\ref{ex:6d}) [@maclaughlin1997].

### Pronouns

The IX sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996]. The same IX function has also been identified for TİD [@sevinc2006]. The example in (\ref{ex:7}) illustrates a spatially-anchored pronominal reference for topic re-introduction. These signs are thus rarely used for the first mention of an entity [@perniss2015]. Other interpretations consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016]. It is important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns. However, for the purposes of the present study, we will refer to IX constructions as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.

### Classifiers

Classifier predicates are highly complex morphological constructions which are very common in sign languages [@supalla1986]. These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003]. As in most sign languages hitherto studied, classifier constructions are also common and found in TİD [see @dikyuva2015 for a review]. Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions. Whole entity classifiers (WECL) represent the entirety of any referent whereas body part classifiers (BPCL) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth). Handling classifiers (HCL) occur with transitive verbs and depict how an object is handled or manipulated by agent. On this ground, an HCL[^1] depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space. Extension classifiers (ExtCL), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape. We exemplify these constructions in (\ref{ex:8}).

[^1]: As discussed before, an HCL is transitive in nature and depicts the handling action performed by ananimate agent. This is why most research tend to consider HCLs as an extension to constructed action [@dudis2004; @cormier2013].

### Null marking

Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null arguments when referring to highly accessible referents which are still active in the mental representation of the addressee. In signed discourse, this can be achieved by employing certain verbal constructions or constructed action (CA). The latter construction exemplified in (\ref{ex:9a}) refers to cases in which the signers engage with the space in front of themselves as if they are interacting with the objects and characters in the story. This is often achieved with the signer taking on the role of a referent in the narrative and performing the actions [@metzger1995]. CA is often marked by certain non-manual cues such as the facial expression and the breaking of the eye contact. This is to say, the narrator's head, face, torso, arms, and as an optional marker, hands represent those of the associated referent(s) [@smith2014]. As for the verbal constructions, to identify the predicate types in ASL, @padden1986 suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs. According to this classification, plain verbs (\ref{ex:9b}) do not agree with the subject or the object. This means that they are not inflected for number or person. This contrasts with agreeing verbs (\ref{ex:9c}) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person. TİD too licenses null arguments. Following this classification, @sevinc2006 argued that although it is common in TİD to drop arguments, there were certain limitations on the null marking of the arguments for agreement and plain verbs.

## Learning How to Track Referents in Sign Languages

In both communication modalities (visual and auditory), children start to produce more complex and coherent stories with appropriate and informative linguistic forms to indicate anaphoric or cataphoric references after age 10 [@hickmann1996; @rathmann2007]. Prior to this age, both signing and speaking children fail to refer to entities consistently since their pragmatic abilities are yet to develop. The manipulation of linguistic forms for discourse functions to signal topic continuance or shift is difficult to master until the beginning of formal education. These findings are also comparable across different sign languages including BSL [@morgan2005], ASL [@loew1984], and DGS [@becker2009], pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.

Like in first language development, narrative skills have been extensively studied among second language learners [see @bel2015; @frederiksen2019]. Comparing 11 deaf adults signers and 13 hearing advanced learners of Catalan Sign Language (Llengua de Signes Catalana - LSC), @bel2015 found that hearing signers resort to more overt forms (nominals and pronouns) than native signers to avoid potential ambiguities among referents. Nevertheless, their results did not reach statistical significant for the interaction of discourse status and acquisition group. A similar line of research came from @frederiksen2019 who found a statistically significant difference between both groups, that is native signers and hearing L2 learners of ASL, only for classifier use especially in re-introduced contexts. L2 learners in the study used more nominals and fewer instances of null marking in number, indicative of a limited effect of overexplicitness among second language learners. However, these studies treated late and native deaf signers as a single group. That is, the control group consisted of both native signers who had one or more deaf parents and also native-like [@cormier2012] late deaf signers who had hearing parents and their age of acquisition varied. Given the preliminary evidence that topic continuity and salience marking is also difficult for second language learners regardless of modality, it has been suggested under the Interface Hypothesis that the syntax-discourse interface is vulnerable and especially challenging to acquire [@sorace2011] for second language learners. Similarly, @sorace2009 discuss that this interface entails much more than morphosyntactic knowledge but also includes cognitive and pragmatic resources that help provide discourse coherence. They propose that the second language discourse is over-redundant. This means that second language learners produce more overt pronouns or fuller forms of language for highly accessible referents possibly as a compensatory strategy in an attempt to reduce the increasing load on cognitive processing.

Still, only very few research examined whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking. @becker2009 examined 6 DGS signing deaf children (3 native and 3 late)'s narrative abilities. She reported that late signing children generated more explicit forms such as nominals when referring to the characters in the story compared to native learners, concluding that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax. Comparably, @cormier2013 investigated the use of CA in 15 BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity. In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents. Most relevantly to the present study, @gur2018 studied the development of narrative skills among 10 native and 10 late deaf children and deaf adult signers of TİD. She found that native-signing children and adults had more instances of explicit referent introduction, set the scene more frequently, and employed CA in an appropriate manner more than corresponding late-signing children and adults. Although this study presented preliminary findings on age of acquisition effects on narrative competence, it does not investigate how native and late signers employ different linguistic tools based on the accessibility of the referent. Therefore, to our knowledge, there is a lack of research directly examining the effects of first language deprivation on deaf adult signers' selection of REs in signed narratives.

Considering this, it is not very clear whether the negative effects on narrative abilities observed among late signing deaf children hold when they become adults. Understanding the impact of language deprivation or delayed first language learning on narrative competence is crucial for three reasons. First, studying this population will shed light on the possible effect of lack of early language experience on future cognitive and pragmatic abilities since the narrative competence is dependent on such faculties. Second, examining reference tracking strategies in particular will also inform about the linguistic inventory that native and late deaf signers use to refer to entities. Finally, any possible difference between the two acquisition groups in terms of narrative competence will have further practical implications for the importance of early access to a first functional sign language to retain future linguistic and cognitive abilities.

## Present Study

In an attempt at addressing this gap in the literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in TİD narratives with a story-telling paradigm. We explore the forms available for referent tracking in TİD and examine the nature of RE selection as a possible indicator of pragmatic competence between our acquisition groups.

Overall, considering what has been suggested for the distribution of REs in signed narratives, we expect more explicit forms like nominals for referent introduction and and less explicit forms such as null marking, pronominal and/or classifier use for referent maintainance. For the re-introduction contexts, where referents which are discontinued across one or more clauses are brought back into the discourse, we expect a less clear picture. Some re-introduced referents will have higher accessibility and will thus be tracked with more implicit REs, while others that have lower accessibility will be tracked with more explicit REs Also based on previous research, we hypothesize that late signers will become more over-redundant in their reference tracking compared to native signers. This would translate to a preference by late signers for nominals over null marking especially in maintenance and re-introduction contexts. In contrast, we would expect that native signers use more null marking to track referents.

# Method

## Participants

We tested 29 deaf adult signers, 15 with exposure to TİD from birth (i.e., native signers), 14 with exposure between ages 3 and 17 (i.e., late signers) (*M*\textsubscript{Age of acquisition} = 7.7 years, *SD* = 3.3 years). The criterion to be classified as a late signer was to have hearing and non-signing caregivers. All the signers satisfying this criterion were regarded as late signers \footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers. Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3. In order to capture in-group variation among late signers, we run an additional supplementary analysis}. All participants indicated that their preferred language of communication was TİD which they used for most contexts. Detailed information about the linguistic and educational background of the participants can be found in \ref{section:a}.

All participating native signers (7 males, 8 females) were born into deaf families and started acquiring TİD from at least one of their caregivers after birth. The parents of 14 participating native signers were deaf. One native signer had a deaf father and a hearing mother. In addition to TİD, the home language for some native signers (n = 4) in early childhood included the use of home sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers. Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}. The age of the native participants ranged between 18 and 35 (*M* = 26.4 years, *SD* = 5.0 years). This also corresponded to the total average years of TİD use for native signers. All native signers were pre-lingually deaf, that is they were either born deaf or they had become deaf before the age of 3, with the exception of one signer who became deaf between the ages of 4 and 7. Native signers reported to complete at least high school education and their level of education (LOE) was between 12 and 16 years (*M* = 12.7 years, *SD* = 1.5 years). In addition, all native participants attended at least two deaf schools, one being primary and the other middle school. 10 native participants also attended a deaf school for their high school education. Based on a self-reported 5-point scale of proficiency (1: poor, 5: proficient), native participants declared that they have high receptive and productive skills in TİD (*M* = 5 points, *SD* = 0 points). As for their proficiency in (spoken) Turkish, their self-reported rankings were on average 3.7 (*SD* = 0.7), 3.7 (*SD* = 0.7), and 2.6 (*SD* = 1.2) for reading, writing, and speaking abilities, respectively.

Comparably, late learners (8 males, 6 females) were all born into hearing families and started learning TİD later in life after age 3, usually with the start of formal education. The age of sign language acquisition for late signers ranged between years 3-17 (*M* = 8.7 years, *SD* = 3.5 years). The parents of all participating late signers were hearing, and the reported home languages used in childhood mainly consisted of Turkish and home sign. The average age of late signers was 34.6 years (*SD* = 8 years) and ranged between 24 and 50. The derived mean number of years for TİD use following its acquisition was comparable to native signers with 25.9 years (*SD* = 7.9 years). Except for two signers who became deaf between the ages of 4 and 7, all the other late signers were pre-lingually deaf. All late signers completed middle school education and the average year of schooling was 12.3 years (*SD* = 2.8 years). All late participants attended at least one deaf school. While 8 participants attended two deaf schools (either primary and middle school, or middle and high school), 3 participants reported attend three deaf schools. Using the same scale, late signers self-rated their abilities in TİD and (spoken) Turkish. Both of their receptive and productive TİD abilities rated on average 4.9 points (*SD* = 0.3 points). Late signers' mean rating for their reading abilities in Turkish was 3.5 points (*SD* = 0.7 points) while their writing and speaking abilities rated on average 3.4 (*SD* = 0.5 points) and 3.1 points (*SD* = 0.7 points), respectively.

Native and late participants' sex, average years of schooling, TİD use following its acquisition, and their self-reported receptive and productive TİD ratings, and their reading, writing, and speaking ratings for Turkish did not statistically differ; Bayesian *p*'s \> 0.1. However, participants' ages statistically differed between the two acquisition groups; *t*(27) = 3.4, Bayesian *p* \< 0.01. Overall, the native group consisted of younger participants. This was to ensure that the average years of TİD use among native signers approximated those of late signers.

## Stimuli

We used a story-telling paradigm to examine how native and late signers of TİD track reference in signed discourse. Participants were presented with 10 short video clips extracted from the cartoon *Tom and Jerry* and asked to retell them. Each clip had six to ten episodes (*M* = `r mean(c(6, 8, 8, 9, 9, 10, 8, 8, 7, 7))` episodes). The duration of the clips ranged from 17 seconds to 28 seconds (*M* = `r mean(c(19, 23, 28, 24, 22, 26, 17, 22, 23, 23))` seconds). The narratives included on average `r round(mean(c(2, 4, 2, 3, 2, 2, 3, 2, 2, 2)))` recurring animate characters for which the present context required the introduction, maintenance, and re-introduction of referents. In addition, in each narrative, there were on average `r round(mean(c(1, 1, 3, 2, 4, 4, 1, 3, 3, 3)))` inanimate objects with which the animate referents interacted. The detailed description of the episodes in each narrative can be found in \ref{section:b}.

```{r tab-2, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
options(scipen=1, digits=2)

df_table_2 <- read.csv("./appendix/tables/table_tier.csv", sep = ";")
apa_table(df_table_2,
          caption ="Annotation Tiers with their Tags and Descriptions",
          note = "NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.",
          escape = FALSE,
          align = c("p{3.2cm}", rep("p{3.8cm}", ncol(df_table_2))),
          placement = "p",
          font_size = "footnotesize",
          landscape = FALSE
)
```

## Task Procedure

Participants sat in front of a laptop computer which had the stimulus stories. The instructions were all given by a deaf research assistant who was present in all sessions and ensured that the task requirements were clear. The informed consent form was translated into TİD and explained to the participants by the same assistant before data recording. Participants filled in the consent form and a background questionnaire form before the onset of the trials. Following this step, signers were seated across a hearing researcher who is an advanced late learner of TİD. We instructed the participants to narrate as they would to a deaf friend. All trials were recorded by a digital camera placed in front of the signers. The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons. First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter. Therefore, most signers preferred to narrate to the camera. Furhermore, some signers did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available. We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers (see \ref{section:c} The results imply that there is not much of a difference.}.

```{r echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
# generic libraries
library(readxl) #read excel
library(gdata) # reorder.factor
library(magrittr) # pipe
library(tidyverse) # tidy
library(ggstance)
library(janitor)
library(MASS)
theme_set(theme_bw()) #set ggplot theme
library(extrafont)
loadfonts()

# statistics libraries
library(brms) # bayesian modeling
library(bayesplot) # plot options for bayesian models
bayesplot_theme_set(new = theme_bw()) # set plot theme
library(tidybayes)
library(bayestestR)
theme_apa <- function(base_size = 10, base_family = "serif", box = FALSE) {
  adapted_theme <- ggplot2::theme_bw(base_size, base_family) +
    ggplot2::theme(legend.position = "bottom",
                   plot.title = ggplot2::element_text(size = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, ggplot2::rel(14), 0))
                   # , axis.title = ggplot2::element_text(size = ggplot2::rel(1.1))
                   , axis.title.x = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(ggplot2::rel(12), 0, 0, 0))
                   , axis.title.y = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, ggplot2::rel(12), 0, 0))
                   , axis.title.y.right = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(12)))
                   , axis.ticks.length = ggplot2::unit(ggplot2::rel(6), "points")
                   , axis.text = ggplot2::element_text(size = ggplot2::rel(0.9))
                   , axis.text.x = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(ggplot2::rel(6), 0, 0, 0))
                   , axis.text.y = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, ggplot2::rel(6), 0, 0))
                   , axis.text.y.right = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(6)))
                   , axis.line = ggplot2::element_line()
                   # , axis.line.x = ggplot2::element_line()
                   # , axis.line.y = ggplot2::element_line()
                   , legend.title = ggplot2::element_text()
                   , legend.key = ggplot2::element_rect(fill = NA, color = NA)
                   , legend.key.width = ggplot2::unit(ggplot2::rel(20), "points")
                   , legend.key.height = ggplot2::unit(ggplot2::rel(20), "points")
                   , legend.margin = ggplot2::margin(
                     t = ggplot2::rel(16)
                     , r = ggplot2::rel(16)
                     , b = ggplot2::rel(16)
                     , l = ggplot2::rel(16)
                     , unit = "points"
                   )
                   , panel.spacing = ggplot2::unit(ggplot2::rel(14), "points")
                   , panel.grid.major.x = ggplot2::element_blank()
                   , panel.grid.minor.x = ggplot2::element_blank()
                   , panel.grid.major.y = ggplot2::element_blank()
                   , panel.grid.minor.y = ggplot2::element_blank()
                   , strip.background = ggplot2::element_rect(fill = NA, color = NA)
                   , strip.text.x = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, ggplot2::rel(10), 0))
                   , strip.text.y = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(10)))
    )
  if(box) {
    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_rect(color = "black"))
  } else {
    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_blank())
  }
  adapted_theme
}

# Read df and drop NAs
df_narration <- read.csv("data/narration_data.csv", sep = ";") %>% tidyr::drop_na()

df_misc <- read.csv("./appendix/appendix_a_table.csv", sep=";") %>% clean_names() %>% 
  dplyr::mutate(Subject = as.integer(no),
                Group = case_when(aoa == "0-3" ~ "Native",
                                  aoa == "4-7" ~ "Late1",
                                  T ~ "Late2") %>% as.factor()) %>% dplyr::select(Subject, Group)
df_narration %<>% left_join(df_misc)

# Coding nominal, Zero, Classifier, and Pronominal categories
df_narration$ref_type <- with(df_narration, case_when(
  GTR1 %in% c("Bare noun", "Point noun", "Noun Point",
              "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
  GTR1 %in%  c("Whole entity CL", "Body part CL",
               "SASS - Tracing", "SASS - Static") ~ "Classifier",
  T ~ "Pronominal"))

# Coding nominal, Zero, Classifier, and Pronominal categories
df_narration$GTR1_plot <- with(df_narration, case_when(
  GTR1 %in% c("Bare noun") ~ "BARE NOM",
  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "M NOM",
  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "FS NOM",
  GTR1 %in% c("Constructed action", "Constructed Action") ~ "CA",
  GTR1 %in% c("Plain verb") ~ "PLAIN",
  GTR1 %in% c("Person agreement verb") ~ "AGR",
  GTR1 %in% c("Whole entity CL") ~ "WECL",
  GTR1 %in%  c("Body part CL")  ~ "BPCL",
  GTR1 %in%  c("SASS - Tracing", "SASS - Static")  ~ "SASS",
  T ~ "Pronominal"))

# Encode vector types
df_narration$Narrative %<>% as.integer() 
df_narration$Subject %<>% as.integer()
df_narration$GTR1 %<>% as.factor() 
df_narration$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
df_narration$Discourse %<>% as.factor()
df_narration$ref_type %<>% as.factor()

# Summary of the data
summary_df <- df_narration %>% 
  group_by(Nativeness, ref_type) %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))

# Descriptive plots
df_narration$ref_type2 <- df_narration$ref_type %>% recode('Classifier' = 'CL', 'Null' = 'NULL',
                                                           'Pronominal' = 'PRO', 'Nominal' = 'NOM')
# dropping Pronominal references because they are too few.
df_model1 <- df_narration %>% subset(ref_type !="Pronominal") %>% drop.levels()

# Contrast coding
df_model1$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
contrasts(df_model1$Nativeness) <- contr.sum(2)/2
contrasts(df_model1$Discourse) <- contr.sum(3)/2
df_model1$Group %<>% reorder.factor(new.order = c("Late2","Late1","Native"))
contrasts(df_model1$Group) <- contr.sdif(3)

# counts of ref types
ref_count <- df_narration %>% group_by(ref_type) %>% summarise(count=n())

# Model for reference type
model_1 <- brm(ref_type ~ Nativeness*Discourse + (1|Subject), 
               data = df_model1, family = categorical(link = "logit"), 
               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_1.rds")

# Posterior plot for reference type
df_model_1 <- mcmc_intervals_data(model_1, pars = vars(starts_with("b_")))
df_model_1$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `b_muNull_Intercept` = "Null_Intercept",
                                        `b_muNominal_Nativeness1` = "Nom_Native",
                                        `b_muNull_Nativeness1` = "Null_Native",
                                        `b_muNominal_Discourse1` = "Nom_Intro",
                                        `b_muNominal_Discourse2` = "Nom_Maintain",
                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
                                        `b_muNull_Discourse1` = "Null_Intro",
                                        `b_muNull_Discourse2` = "Null_Maintain")
saveRDS(df_model_1, "./rds/df_model_1.rds")

# Supplementary model with 3 groups of acquisition
model_3 <- brm(ref_type ~ Group*Discourse + (1|Subject), 
               data = df_model1, family = categorical(link = "logit"), 
               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_3.rds")

# Posterior plot for reference type
df_model_3 <- mcmc_intervals_data(model_3, pars = vars(starts_with("b_")))
df_model_3$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `b_muNull_Intercept` = "Null_Intercept",
                                        `b_muNominal_Group2M1` = "Nom_Late1-Late2",
                                        `b_muNominal_Group3M2` = "Nom_Native-Late1",
                                        `b_muNominal_Discourse1` = "Nom_Intro",
                                        `b_muNominal_Discourse2` = "Nom_Maintain",
                                        `b_muNominal_Group2M1:Discourse1` = 'Nom_L1-L2*Intro',
                                        `b_muNominal_Group3M2:Discourse1` = 'Nom_N-L1*Intro',
                                        `b_muNominal_Group2M1:Discourse2` = 'Nom_L1-L2*Maintain',
                                        `b_muNominal_Group3M2:Discourse2` = 'Nom_N-L1*Maintain',
                                        `b_muNull_Group2M1` = "Null_Late1-Late2",
                                        `b_muNull_Group3M2` = "Null_Native-Late1",
                                        `b_muNull_Discourse1` = "Null_Intro",
                                        `b_muNull_Discourse2` = "Null_Maintain",
                                        `b_muNull_Group2M1:Discourse1` = 'Null_L1-L2*Intro',
                                        `b_muNull_Group3M2:Discourse1` = 'Null_N-L1*Intro',
                                        `b_muNull_Group2M1:Discourse2` = 'Null_L1-L2*Maintain',
                                        `b_muNull_Group3M2:Discourse2` = 'Null_N-L1*Maintain')
saveRDS(df_model_3, "./rds/df_model_3.rds")
#df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))
```

## Data Annotation

We coded the discourse status (Introduction, Maintenance, Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null). We used ELAN Linguistic Annotation Software [@crasborn2008] for data annotation. First Author who is a hearing signer of TİD annotated all the data and translated the narratives into Turkish together with a native deaf signer of TİD.

Following @berman2013relating, we first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod). On the second tier, we coded the discourse status for each referent. Having adapted the classification of @gullberg2006, we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence. Only subjects were coded for maintenance and re-introduction. We coded subject referents as maintained only if they had been referred to in the previous clause. To this end, the referents which had been discontinued across one or more sentences were all regarded re-introduced into the discourse. On a third tier, we had the RE type for a single reference with the main categories nominal, pronominal, null marking, and classifier. Under nominal, we identified bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we distinguished among three types of constructions: WECL, BPCL, and ExtCL. Finally, we had CA, agreement, plain tags for null constructions. We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of CA). The annotation tiers are presented in Table \ref{tab:tab-2}.

## Data Analysis

In our analysis, we use Bayesian regression models with the help of the brms package [@burkner2018] in R. In a Bayesian regression the prior and the likelihood is used to calculate the posterior. Instead of a single filter of significance, we get a posterior probability distribution. We do not define our own priors and use the brms package defaults. We are using Bayesian regression to avoid some common pitfalls of other approaches [@wagenmakers2007; @kruschke2011].

For this particular type of experimental data, we fit the responses to a multinomial distribution. Most of the time, experimental results have yes/no-true/false responses which are fit to binomial distributions, reading/response times which are fit to log-normal distributions, and ratings which are fit to cumulative distributions. In these cases, the researcher is not particularly interested in the \`intercept' coefficient which shows the mean value of the response. The researcher is usually interested in the remaining coefficients which give information about predictor level comparisons.

In a multinomial distribution, the responses are in a categorical relationship where the levels can be A,B,C, and more. These levels are not ordered in time, they are not cumulative, and they are not additive. If we were to make an analogy, they are like countries on Earth which can have differences and similarities, but they don't have to. In a multinomial distribution, one of the responses becomes the level of comparison and other levels are compared against it. In a way, there are multiple binomial distributions. However, the intercepts in a multinomial distribution end up having more of an interest for the researcher. This is because the intercept is a comparison of two levels of the response to the exclusion of the remaining response levels. There is another consideration that should go into interpreting the results of a multinomial distribution. Different from the intercepts, the remaining coefficients are predictor level comparisons within the grand comparison of the response levels. To indicate which predictor level comparison belongs to which response comparison, we name our coefficients with prefixes followed by an underscore.

In our model plots, we report median estimates as points, 50% credible intervals as a thick line and 95% credible intervals as a thin line. These numbers are relatively arbitrary, and they are intended to show the posterior probability distribution of the coefficients. We use interval plots to save space and have a more compact plot. If a considerable amount (\>90%) of the posterior probability distribution is towards a sign (-/+) we interpret it as a categorial effect, if a good amount (\>50%) of the distribution is towards a sign (-/+) we interpret it as a tendency, and we see the remaining distributions as indifference. In addition to these, we look at how the distribution skewed in the extremes or how \`wide' it is from the median estimate to inform us mainly about uncertainty and variation.

# Results

We collected `r nrow(df_narration)` data points as a result of the experiment. Each data point is a referent tracked by the participant. Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null). However, we allude to the numerical distribution of the subtypes in this and following sections}, and Participant Group (Native, Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17). Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}.

Across our acquisition groups, we observed `r sum(df_narration$Discourse == "Introduction")` (`r sum(df_narration$Discourse == "Introduction")/nrow(df_narration)`) instances of referent introduction, `r sum(df_narration$Discourse == "Maintenance")` (`r sum(df_narration$Discourse == "Maintenance")/nrow(df_narration)`) instances of referent maintenance, and `r sum(df_narration$Discourse == "Re-introduction")` (`r sum(df_narration$Discourse == "Re-introduction")/nrow(df_narration)`) instances of referent re-introduction. For the introduction of a referent, the majority of the REs were nominal constructions (Table \ref{tab:my-table}). However, we observed a sporadic use of classifiers and null marking respectively. When the discourse status was maintained, signers extensively used null markers to refer to the entities in the immediately previous clause. Classifier constructions followed null markers in frequency in maintained contexts. We observed only a few instances of nominal use for referent maintenance. In re-introduction contexts, signers used a greater proportion of nominal markings which were followed by null, classifier, and pronominal use.

```{=tex}
\begin{table}
\small
\caption{Distribution of all discourse status by RE types}
\label{tab:my-table}
\begin{tabular}{llll}
\hline
                     & Introduced    & Maintained     & Re-introduced \\ \hline
\textit{Nominal:}    & 92.78\% (989) & 5.63\% (93)    & 63.65\% (907) \\
\textit{Pronominal:} & 0.09\% (1)    & 0.24\% (4)     & 2.46\% (35)   \\
\textit{Classifier:} & 5.35\% (57)   & 18.29\% (302)  & 8\% (114)     \\
\textit{Null:}       & 1.78\% (19)   & 75.83\% (1252) & 25.89\% (369) \\ \hline
\end{tabular}%
\end{table}
```
Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was \textsc{bare nom} which was followed by nouns modified by index or classifiers. In number, signers used \textsc{mod nom} more in introduction and re-introduction compared to maintenance settings. Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}). A great majority of the pronominal constructions were used to re-introduce a referent into the discourse. We observed only a very small amount of pronoun use for referent introduction and maintenance.

```{=tex}
\begin{table}
\small
\caption{Distribution of nominals by discourse status}
\label{tab:my-table2}
\begin{tabular}{lllll}
\hline
                        & Bare Noun  & Modified Noun & FS Noun \\ \hline
\textit{Introduced:}    & 48.93\% (842) & 53.56\% (139)     & 100\% (8) \\
\textit{Maintained:}    & 4.82\% (83)  & 3.85\% (10)      & 0\% (0) \\
\textit{Re-introduced:} & 46.25\% (796) & 42.69\% (111)     & 0\% (0) \\ \hline
\end{tabular}
\end{table}

\begin{table}
\small
\caption{Distribution of pronominals by discourse status}
\label{tab:my-table3}
\begin{tabular}{ll}
\hline
                        & Pronominal  \\ \hline
\textit{Introduced:}    & 2.5\% (1)   \\
\textit{Maintained:}    & 10\% (4)    \\
\textit{Re-introduced:} & 87.5\% (35) \\ \hline
\end{tabular}
\end{table}
```
As for the classifier constructions (Table \ref{tab:my-table4}), WECL was the most common type, followed by BPCL and ExtCL. For the first mention of a referent, signers mostly preferred to use ExtCL. However, for referent maintenance and re-introduction WECL was used more compared to BPCL and ExtCL.

Among the null markers, CA was most frequently employed, and it was followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}). Overall, signers used very few null markings for the introduction of a referent and the majority of the existing ones were CA. Additionally, signers used CA more for referent re-introduction than maintenance. We observed a similar instances of agreement verbs for maintained and re-introduced contexts. In contrast, plain verbs were favored more in the maintenance setting than re-introduction.

```{=tex}
\begin{table}
\small
\caption{Distribution of classifiers by discourse status}
\label{tab:my-table4}
\begin{tabular}{llll}
\hline
                        & WECL          & BPCL        & ExtCL         \\ \hline
\textit{Introduced:}    & 2.14\% (6)    & 2.4\% (3)   & 71.64\% (48) \\
\textit{Maintained:}    & 67.26\% (189) & 77.6\% (97) & 23.88\% (16) \\
\textit{Re-introduced:} & 30.6\% (86)   & 20\% (25)   & 4.48\% (3)   \\ \hline
\end{tabular}
\end{table}
```
Across the discourse statuses, there is also a difference between participant groups in the use of null marking for reference tracking. Native signers of TİD used it more (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`) compared to the late signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`). Comparably, late signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`) than native signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`). Figure 1 shows the numerical distribution of the responses by RE type and Group.

```{=tex}
\begin{table}
\small
\caption{Distribution of null markers by discourse status}
\label{tab:my-table5}
\begin{tabular}{llll}
\hline
                        & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
\textit{Introduced:}    & 1.58\% (17)        & 0.28\% (1)    & 0.5\% (1)      \\
\textit{Maintained:}    & 71.64\% (773)      & 89.97\% (323) & 77.23\% (156)  \\
\textit{Re-introduced:} & 26.78\% (289)      & 9.75\% (35)   & 22.28\% (45)   \\ \hline
\end{tabular}
\end{table}
```

```{r count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))

p1 <- df_narration %>%
  dplyr::select(Nativeness, ref_type2, Discourse) %>%
  group_by(Nativeness, Discourse) %>%
  ggplot(aes(ref_type2, fill = Nativeness)) + geom_bar(position = "dodge") +
  stat_count(geom = "text", aes(color = Nativeness, label = ..count..), vjust = 0,
             position = position_dodge(width = 0.85), size = 2) +
  facet_grid(~Discourse) + ylab("Count") + xlab("REATs") +
  scale_color_grey() +
  scale_fill_grey() +
  theme_apa() 
p1
```

For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors. We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late). As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses. For the current analysis, we did not define priors ourselves and used the preset uninformed priors. \@ref{fig:model-gram}[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals. There are 3 responses we are considering in this model. They are; Classifier, Nominative, and Null. Classifier response is the level that other responses Nominative and Null are compared against.

[^2]: We ran another model where we used the Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our model did not converge properly. We provide the model results as a plot and table in the \ref{section:d}.

```{r model-gram, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
p3 <- df_model_1 %>% 
  ggplot(aes(m, y=factor(parameter,levels = rev(levels(factor(parameter)))))) +
  geom_point(size = 2.5) +
  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5) +
  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0)) + vline_0(linetype = "dashed") +
  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa()
p3
```

In a multinomial distribution, the model takes one of the responses as a reference and compares all other responses to it. For example, the first coefficient, Nom_Intercept (nominal intercept), in Figure 2 is the comparison of a nominal versus a classifier. It means that participants produced more nominals than classifiers in general regardless of participant group and discourse status. The second coefficient shows that null markings were also used more than classifiers by the participants in general, but not as much as nominals.

In the comparison of nominals and classifiers, being a native acquirer of TİD did not have an effect. Nominals were used more when the discourse status was introduction compared to re-introduction. Again, the whole comparison is made against a classifier. In contrast to the introduction status, nominals were used less in maintenance contexts compared to the re-introduction contexts within the greater comparison to classifier referents. There seems to be an interaction between the introduction status of discourse and being a native acquirer of TİD. When the acquirer is a native speaker and the discourse function is introduction, the chances of observing a nominal referent gets even higher. Also, there seems to be an interaction between maintaining a referent and being a native acquirer of TİD. When it was a native acquirer of TİD and the discourse context was maintenance, the chances of observing a nominal referent gets even lower.

Compared to classifiers, being a native acquirer of TİD increased the odds of using null marking in general. Null constructions were used less in introduction contexts than in re-introduction. This contrasts with the nominals (4th coefficient) as they were used more in referent introduction versus re-introduction. In contrast to introduction contexts, null markers were used more in referent maintenance compared to re-introduction, yet again within the greater comparison to the classifiers. There seems to be an interaction between introducing a referent and being a native acquirer of TİD. In that, the odds of using null marking increases if the participant is a native acquirer of TİD and the referent is used in an introduction context. There is another interaction with regards to referent maintenance, contrasting the first interaction. In that, the odds of using null markers decrease if the participant is a native acquirer of TİD and the referent is maintained. This does not mean that late signers used more null markers in maintenance contexts. It means that the use of null markers did not linearly increase as the addition of using a null marker instead of a classifier as a native plus using null markers instead of a classifier for maintaining a referent versus re-introducing it.

# General Discussion

In our study, we aimed to (i) describe how signers track referents in signed discourse in TİD and (ii) examine whether language deprivation results in more over-redundant referent selection. To this end, we investigated the distribution of referring expressions (REs) that the native and late signers select when employing referents in different discourse statuses (introduced, maintained, and re-introduced).

In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as null marking, pronominal use, and classifiers. In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.

For age of acquisition effects, based on what is proposed for child and adult deaf signers with language deprivation in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayabasi2020] and narrative development [@becker2009; @cormier2013], we expected that late signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015]. That is, we had the expectation that late signers would use more explicit markers like nouns and fewer implicit markers like null marking even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than the control group (i.e., native signers).

## Tracking referents in TİD

The results of the present study partly support our hypotheses. In line with our expectations, regardless of their acquisition groups (native or late) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used null marking and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, null markers and pronouns) to re-introduce a referent.

However, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies. For a comparison, while the percentage of introduced referents were similar in both research (26% versus 24%), the participants in the present study brought back referents into the discourse more frequently compared to the participants in @frederiksen2019 and @frederiksen2016 which reported that re-introduction contexts accounted for only 7% of the referents in ASL. On the other hand, maintenance percentages were lower in the present study compared to @frederiksen2016 (40% versus 69%). Investigating the distribution of REs across discourse statuses in ASL, a similar study by @czubek2017 also reported contrasting numbers (17% for introduction, 64% for maintenance, and 19% for re-introduction). We suggest that the discrepancy might be due to the nature of the stimuli used in these research. In @frederiksen2016, the authors used 4-6-episode-long picture story with 2 animate characters and 1 inanimate object. In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects. Although the number of animate characters and inanimate characters in @czubek2017's stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli. We believe fewer possible referents and/or a simpler narrative structure in these studies may have contributed to the lower number of re-introduced contexts compared to the present study. Due to the story length and composition in this study, the participants had to track more recurring animate or inanimate referents for a longer period of time. As a result, our signers engaged in referent re-introduction more than what was reported in previous studies [@czubek2017; @frederiksen2016].

As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used null markers such as CA, and agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained). This finding parallels other accounts of referential accessibility suggested for other sign languages [see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barbera2009 for LSC. Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts compared to re-introduced contexts. This brings along the assumption that modified nouns are used more to track referents with lower accessibility, probably due to their more explicit structure.

Interestingly and to our surprise, the signers in this study used pronominal IX very sporadically with only 40 observed instances that were mainly used to re-introduce referents rather than maintain them. This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities. Comparable to our findings, It has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers different genre [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD]. Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016]. Similar to these research, our findings indicate that the pronominal IX in TİD was used more concurrently with nominal constructions and pronominal IX was used very infrequently in isolation. The observation that overt pronouns are rare in ASL narratives comes from studies that include a simple picture retelling task with one protagonist [e.g., @frederiksen2016] as well as a more complex narrative design as in [@czubek2017]. However, @czubek2017's study also incorporated different genres and reported more instances of overt pronouns in personal narratives (compared to picture or video retelling). In fact, this finding suggests that the use of pronouns might be genre-dependent. Moreover, another production study on referent tracking in ASL reported object preference for pronominals [@frederiksen2017]. Nonetheless, we only coded subjects for maintained and re-introduced referents. We argue that all of these factors (genre-dependency as well as object preference) might contribute to the explanation of why overt pronouns in this data are dispreferred in favor of other markers for highly accessible referents such as null marking and classifiers.

A deeper look into the allocation of classifier constructions revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs. We observed a similar alignment for re-introduced contexts. In contrast, a great number of the overall ExtCLs were used to introduce a new referent into the discourse. That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCLs constructions were mainly used for referents which were not easily accessible. We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without being redundant. As pointed out by @czubek2017, signers prefer to rely on ExtCLs for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space. Nevertheless, we only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating @frederiksen2019's findings for ASL.

As for the distribution of the null markers, approximately 90% of the plain verbs were used to maintain a previous referent, making these constructions least implicit REs used for TİD[^3]. Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively. This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on scale, indicating that these constructions were more implicit compared to plain verbs. However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other null markers as in the present study [@frederiksen2016; @czubek2017; @barbera2009]. In Figure \ref{fig:scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL. In the hierarchies, the left most REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low. If the referent is already active in the addressee's mind and its accessibility is high, then the right most markers (e.g., agreement and plain verbs for TİD) are selected.

[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping. For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion (Meir 1999). As for plain verbs, both phonology of a sign and its semantics have an effect on morphology. For instance, \textsc{eat} in TİD is signed nearby the mouth and since it needs to remain in that phonological location, it is not suitable for agreement and thus it is regarded as a plain verb. Likewise the experiencer verb \textsc{love} is signed on the torso and as such it does not take part in agreement. These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other. For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause (Gökgöz and Sevgi, 2020). See also for a similar discussion in Brazilian Sign Language.}.

```{=tex}
\begin{figure}
\caption{Comparing the present study and Frederikse and Mayberry (2016) in terms of proposed accessibility hierarchies}
\label{fig:scale}
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16) -- (214.92,115.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15) -- (314.89,115.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34) -- (414.89,115.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize M NOM}};
% Text Node
\draw (191,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize BARE NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node
\draw (285,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize WECL \& BPCL}};
% Text Node
\draw (359,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize CA}};
% Text Node
\draw (405,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR}};
% Text Node
\draw (452,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in the present study for TİD
    \end{minipage}

\vspace{1 cm}

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node% Text Node
\draw (333,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR \& CA}};
% Text Node
\draw (445,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in Frederiksen and Mayberry (2016) for ASL
    \end{minipage}
\end{figure}
```
Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals and ExtCLs as markers of lowly accessibly referents. Likewise, in both hierarchies, as markers of highly accessible referents, different types of null markers are further distinguished. Plain verbs rank the lowest on the scales, preceded by agreement verbs and CA. On a similar ground, both scales lack pronouns, which occurred very infrequently in the data. Although @frederiksen2016 did not include other classifier types in the hierarchy, we proposed that WECLs and BPCLs must ranker higher than null markers but lower than ExtCLs that were mainly used by interlocutors to introduce a referent into the discourse context.

## Age of acquisition effects on referent tracking

The results in this study only somewhat support our second set of hypotheses that we set for the effects of age of acquisition on participants' RE selection. That is, while being a native acquirer of TİD increased the odds of using null marking in reference tracking across different discourse statuses, we failed to find such a difference for overall nominal production. Taking a closer look at the interaction terms, we observed that native signers, compared to late signers, were less likely to use nominal constructions when the discourse status was maintained. Native acquisition further increased the odds of null markers when the discourse status was introduced. However, we found the opposite alignment for the interaction of maintenance and native acquisition. That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a native signer. This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REs in TİD narratives better than age of acquisition (native, late) which, in comparison, had a less salient effect.

While late signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts. In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA[@cormier2013]. As for TİD, to our knowledge, there is only one study [@gur2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills. @gur2018 observed that late signing deaf adults and children set the scene less frequently and used CA less accurately compared to their native counterparts. However, this particular investigation did not present any data on the distribution of REs between two acquisition groups. In the present study, the finding that native signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002]. These research reported very little or no use of null marking by native signers for introduced contexts.

Given the paucity of research directed at the late signing group, we now turn to the referent tracking strategies observed for second language learners of a sign language. Our study replicated some of the findings of the line of research that investigated deaf or hearing bimodal bilingual signers [@bel2015; @frederiksen2019].\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).} The authors reported that more instances of null marking in the context of referent introduction was attested for native signers than deaf or hearing bimodal bilingual signers. Although using null marking to introduce a referent into the discourse would normally result in referential ambiguity, it is important to note that our stimuli consisted of 10 interconnected narratives with 2 or more recurring characters. It may be possible that native signers, having already introduced the referent with an explicit nominal marker in a different discourse context, found it felicitous to refer to the same entity with more implicit markers in the following narratives even for introduced contexts. Then, our results imply a difference in pragmatic sensitivity to referential accessibility between native and late signers, which is more robustly reflected in introduced contexts.

Since we only observed a small difference between the two acquisition groups for the use of certain REs, the aforementioned findings allude to the implication that the effect on language deprivation on pragmatic factors should also be considered in different contexts. For one, both groups of signers can be tested with different genres of narratives and with a more complex story design. In order to capture possible variation in the accessibility levels of re-introduced referents, a more fine-tuned protocol [as in @toole1996] can be used for measuring referent distance and the number of other possible competitors. Taken altogether, our results accentuates the necessity of engaging in signed narrative production early in life as well as the importance of early intervention for late signing deaf children.

# Conclusion

The present study aimed to investigate the reference tracking strategies of native and late deaf adult signers in TİD narratives. In line with our expectations, we found that the distribution of REs (nominal, pronominal, classifier, and null marking) followed from previous accounts of saliency and referential accessibility. For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, hey mainly used null arguments (CA, and plain and agreement verbs) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents. We observed only very few instances of pronouns (mostly produced by late signers), unlike our expectations. The ones we observed were mostly reserved for re-introduced contexts along with nominals. Based on our findings across the two groups, we proposed a hierarchy of referent accessibility for TİD. As for the age of acquisition effects, we observed limited over-redundancy for signers with delayed language acquisition (i.e., late signers) who employed fewer implicit forms of reference tracking compared to native signers. Native signers were less likely to use a nominal for maintained contexts and more likely to use null marking for introduced contexts than late signers. We conclude that native signers might only partially adhere to pragmatic norms and leverage accessibility better than late signers since they use less form when the referent is more active in the mind.

\clearpage

\appendix

# Participant background information

\label{section:a}

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
h1 = c("No", "Status", "Sex", "AOA", "Age", "LOE", "Deaf schools", "Home language", "Parents")

df_table_a <- read.csv("./appendix/appendix_a_table.csv", sep = ";")
names(df_table_a) <- h1
df_table_a %<>% dplyr::select(-Parents) 
apa_table(df_table_a,
          caption ="Participant demographic information based on self-report",
          note = "AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.",
          escape = FALSE,
          placement = "hbt!",
          font_size = "small",
          landscape = FALSE
)
```

\clearpage

```{=tex}
\setlength{\columnsep}{10pt}
\twocolumn
\linespread{1}
\justifying
\footnotesize
```

# Episodes in the stimuli

\label{section:b}

**Narration 1**

1\. Tom chases Jerry, wearing a wooden plank

2\. Tom and Jerry enter a tube

3\. Tom falls down and Jerry runs away

4\. Tube gets thinner and Tom chases Jerry

5\. Jerry gets out and escapes

6\. Tom follows Jerry in shape of a stick

**Narration 2**

1\. Tom walks past a female cat sitting on a couch

2\. Tom turns to and approaches female cat

3\. Female cat blushes

4\. Tom offers a fish in aquarium to female cat

5\. Female cat rejects offer

6\. Tom offers a bird in cage to female cat

7\. Episode 5 is repeated

8\. Tom leaves by walking in a flirtatious manner

**Narration 3**

1\. Jerry falls down and tries to walk up stairs

2\. Tom pulls carpet back

3\. Jerry struggles

4\. A piano falls down stairs

5\. Jerry manages to escape

6\. Tom sees piano and tries to run away

7\. Piano smashes into Tom and flattens him

8\. Tom falls on ground

**Narration 4**

1\. Tom hides behind a cardboard house

2\. Jerry walks to house with a toy mouse

3\. Toy leaves Jerry who stops at entrance

4\. Toy enters house

5\. Toy gets smashed into pieces

6\. Jerry is surprised

7\. Tom hurts his stomach after swallowing toy

8\. Tom looks at his broken teeth on a mirror

9\. Tom gets angry and throws down mirror

**Narration 5**

1\. Tom fakes sleeping and suddenly gets up

2\. Tom pulls a rope inside a couch

3\. Jerry appears holding a bar of soap

4\. Jerry washes Tom's mouth with bar and flees

5\. Tom spits soap out

6\. Jerry escapes to nest with rope

7\. Tom catches rope and stars pulling it

8\. A mouse trap appears and catches Tom's nose

9\. Tom yells and removes trap

**Narration 6**

1\. Tom points a sword at Jerry and catches him

2\. Tom looks at Jerry

3\. Red spots appear on Jerry's face

4\. Tom gets scared and drops Jerry

5\. Tom rushes into bathroom

6\. Tom washes his hands

7\. Tom opens a drawer

8\. Tom swallows pills one after another

9\. Feeling ill, Tom looks out of window

10\. Tom and Jerry look at each other

**Narration 7**

1\. Two male cats chase one another

2\. Tom sees a female cat and kisses her

3\. Tom hides away when black cat returns

4\. Tom pokes his head behind couch again

5\. Episode 3 is repeated

6\. black cat sees female cat and hugs her

7\. Tom appears and black cat kisses Tom

8\. Two male cats move away from one another

**Narration 8**

1\. Jerry hits Tom's knee with a hammer

2\. Tom suddenly jumps up

3\. Tom yells and holds his knee

4\. Jerry appraches Tom holding a thermometer

5\. Jerry sticks thermometer into Tom's mouth

6\. Jerry lights a fire under thermometer

7\. Red liquid starts to rise up

8\. Thermometer explodes

**Narration 9**

1\. Tom sits on Jerry who tries to get loose

2\. Jerry looks around and sees a hat ribbon

3\. Jerry finds a needle attached to hat

4\. Jerry takes needle and sticks Tom with it

5\. Tom turns red and jumps

6\. Jerry placesn eedle between two cushions

7\. Tom dodges needle

**Narration 10**

1\. Wearing a stethoscope, Tom catches Jerry

2\. Jerry yells into stethoscope

3\. Tom's head pops up

4\. Jerry runs and enter a mouse nest

5\. Tom inserts a gun into nest

6\. Gun bends over toward Tom

7\. Tom fires gun and his head explodes

\clearpage

\onecolumn

# Effect of narrating to the camera vs. experimenter

\label{section:c}

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}

# df for appendix (those who narrated to the experimenter are removed)
df_only_camera <- df_narration %>% filter(!Subject %in% c(16, 20, 21, 24))

# Coding nominal, Zero, Classifier, and Pronominal categories
df_only_camera$ref_type <- with(df_only_camera, case_when(
  GTR1 %in% c("Bare noun", "Point noun", "Noun Point",
              "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
  GTR1 %in%  c("Whole entity CL", "Body part CL",
               "SASS - Tracing", "SASS - Static") ~ "Classifier",
  T ~ "Pronominal"))

# Encode vector types
df_only_camera$Narrative %<>% as.integer() 
df_only_camera$Subject %<>% as.integer()
df_only_camera$GTR1 %<>% as.factor() 
df_only_camera$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
df_only_camera$Discourse %<>% as.factor()
df_only_camera$ref_type %<>% as.factor()

# dropping Pronominal references because they are too few.
df_model_camera <- df_only_camera %>% subset(ref_type !="Pronominal") %>% drop.levels()

# Contrast coding
df_model_camera$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
contrasts(df_model_camera$Nativeness) <- contr.sum(2)/2
contrasts(df_model_camera$Discourse) <- contr.sum(3)/2

# counts of ref types
ref_count <- df_only_camera %>% group_by(ref_type) %>% summarise(count=n())

# Model for reference type
model_camera <- brm(ref_type ~ Nativeness*Discourse + (1|Subject), 
               data = df_model_camera, family = categorical(link = "logit"), 
               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_camera.rds")

# Posterior plot for reference type
df_model_camera <- mcmc_intervals_data(model_camera, pars = vars(starts_with("b_")))
df_model_camera$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `b_muNull_Intercept` = "Null_Intercept",
                                        `b_muNominal_Nativeness1` = "Nom_Native",
                                        `b_muNull_Nativeness1` = "Null_Native",
                                        `b_muNominal_Discourse1` = "Nom_Intro",
                                        `b_muNominal_Discourse2` = "Nom_Maintain",
                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
                                        `b_muNull_Discourse1` = "Null_Intro",
                                        `b_muNull_Discourse2` = "Null_Maintain")
saveRDS(df_model_camera, "./rds/df_model_camera.rds")
```

```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
df_model_1 %<>% mutate(model = "All")
df_model_camera %<>% mutate(model = "Camera")
model_comparison <- rbind(df_model_1, df_model_camera)

p4 <- model_comparison %>% 
  ggplot(aes(m, y=factor(parameter, levels = rev(levels(factor(parameter)))), group = model, color = model)) +
  geom_point(size = 2.5, position = position_dodge(width = 0.6)) +
  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5,position = position_dodge(width = 0.6)) +
  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0), position = position_dodge(width = 0.6)) + vline_0(linetype = "dashed") +
  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_apa()
p4
```

\clearpage

# Additional model with 3 levels for Nativeness

\label{section:d}

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(janitor)
df <- summary(model_3) %>% .$fixed %>% as_tibble()
df_table_extra <- df %>% clean_names() %>% rownames_to_column("Coefficients") %>% dplyr::select(-7,-8)

names(df_table_extra) <- c("Coefficients", "Estimate", "EstError", "lower95", "upper95", "Rhat")

df_table_extra$Coefficients %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `muNull_Intercept` = "Null_Intercept",
                                        `muNominal_Group2M1` = "Nom_Late1vsNative",
                                        `muNominal_Group3M2` = "Nom_Late2vsLate1",
                                        `muNominal_Discourse1` = "Nom_Intro",
                                        `muNominal_Discourse2` = "Nom_Maintain",
                                        `muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
                                        `muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
                                        `muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
                                        `muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
                                        `muNull_Group2M1` = "Null_Late1vsNative",
                                        `muNull_Group3M2` = "Null_Late2vsLate1",
                                        `muNull_Discourse1` = "Null_Intro",
                                        `muNull_Discourse2` = "Null_Maintain",
                                        `muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
                                        `muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
                                        `muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
                                        `muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
df_table_extra %<>% mutate( across(.cols = "Coefficients",~str_replace( ., "_","\\\\_")))

apa_table(df_table_extra,
          format = "latex",
          booktabs = T,
          escape = FALSE,
          landscape = FALSE,
          font_size = "small",
          placement = "hbt!",
          caption ="Model results for Nativeness with 3 levels",
          note = "Rhat values above 1 indicate that the regression model chains did not converge properly. There were 4 chains in our regression model."
)
```

\clearpage

# References