### 2022-08-05 manuscript_v2.tex

\documentclass[review]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{Journal of Pragmatics} % Sets Journal name

\usepackage{lineno} % add

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{floatrow}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{float}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\floatsetup[figure]{capposition=top}
\floatsetup[table]{capposition=top}
\usepackage{times}
\usepackage{smartdiagram}
\usepackage{ragged2e}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{./sty/langsci-gb4e}
\usepackage{caption}
\usepackage{multicol}
\setlength{\columnsep}{-3cm}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\begin{frontmatter}

  \title{Reference Tracking Strategies of Native and Late Signers in
Turkish Sign Language}
      \cortext[cor1]{Corresponding author}

  \begin{abstract}
  We investigated the reference tracking strategies among native and
  late adult deaf signers of Turkish Sign Language. Consistent with the
  theories of saliency and referential accessibility, regardless of
  their acquisiton groups, signers mainly used nominals and extension
  classifiers for introducing referents. To maintain a referent in the
  previous clause, signers chose null marking (e.g., constructed action,
  agreement and plain verbs). For topic shifts or re-introduced
  contexts, nominals and pronouns were chiefly favored although we
  observed very little pronominal use. As for age of acquisition
  effects, we only report limited over-redundancy for late signers who
  used null marking less compared to native signers especially for
  introduced and maintained contexts. Given our limited support for
  redundancy by late signers, we conclude that native signers might
  utilize the spatial affordances of the visual-spatial modality better
  than late signers only to a certain extent.
  \end{abstract}
    \begin{keyword}
    Referential accessibility, signed narratives, pragmatic competence,
delayed language acquisition, Turkish Sign Language \sep 
    Referential accessibility, signed narratives, pragmatic competence,
delayed language acquisition, Turkish Sign Language
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Narratives are extensive discourses that construct an important part of
human communication. In order to create a systematic and coherent
organization of experiences, narrators must choose appropriate
linguistic forms to refer to entities
\citep{ariel1990, givon1983, gundel1993}: Certain forms such as nominal
expressions are commonly used for the first mention of a referent in a
discourse, and other forms such as pronouns and null marking may track
previously mentioned entities which could either be maintained across
multiple clauses or re-introduced back into the discourse. The selection
of these referring expressions (henceforth REs) is contingent on
cognitive abilities, such that individuals must rely on the activation
of the working memory to keep in mind previously referred expressions
and long strings of information \citep{morgan2005, bamberg1997}.
Pragmatic abilities too play a pivotal role in the tracking of
referents. The form with the least amount of redundant information must
be selected based on the accessibility levels of the referent (i.e., how
active a referent is in the addressee's mind given a point in discourse)
\citep{ahn2019, ariel1990}. Because of these cognitive and pragmatic
necessities, it is not surprising to observe that producing a coherent
narrative is one of the latest milestones in language acquisition among
children \citep{hickmann1996}. As can be expected, there is a plethora
of research examining how narrative competence is achieved in both
monolingual and bilingual acquisition contexts among children and
adults. Although narrative competence and referent selection have been
studied extensively for typically developing speakers
\citep{aksu-koc2015, hickmann1996, williams1988}, deaf individuals who
acquired a sign language in atypical learning contexts (i.e., deaf
signers with linguistic deprivation) have remained mostly understudied
with an exception of handful of studies
\citep{cormier2013, gur2018, becker2009}. Nevertheless, it is paramount
that we investigate and describe the effects of language deprivation
among deaf adults on narrative competence more in depth. This will help
us understand whether the loss of systematic and frequent language input
in the early years of life affect how deaf signers produce a narrative
in their adulthood. Given this, the present study will focus on deaf
adult signers with language deprivation and examine their reference
tracking strategies in signed narratives.

Sign Languages are natural languages of the Deaf communities all around
the world. They are independent languages, which do not derive from
majority spoken language(s) but emerge naturally when deaf people form a
community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and has
been the officially recognized language of the Deaf community in Turkey
since 2005 \citep[see][ for a brief review of the historical development
of the language]{nuhbalaoglu2018}. Deaf or hearing children acquiring a
SL within a family where there is one or more deaf caregivers display
the same developmental milestones as hearing children acquiring a spoken
language in a hearing family
\citep{chamberlain_language_1999, meier1990}. However, reports indicate
that over 90\% of all deaf individuals are born into hearing families
who do not know how to sign (i.e., those with language deprivation)
\citep{woll2013}. This means that the auditory modality used in the
household is not accessible to many deaf children. That is why the age
of acquisition of a first functional language varies greatly among deaf
children, with many of whom being deprived of their first languages
typically until after age 7 when they start formal education in a deaf
school. First language deprivation or delay has been found to have
negative effects on certain cognitive and linguistic measures among late
signing children. To name a few, delayed exposure to a first functional
language is reported to negatively affect deaf children's
morphosyntactic production and processing
\citep{kayabasi2020, boudreault2006, newport1990, cheng2021, karadoller2017, cormier2012}
and narrative abilities \citep[e.g.,][]{becker2009}. In our overall
research, we are concerned with the effects of this late exposure in the
ultimate adult forms of TİD users.

\hypertarget{referring-expression-re-selection}{%
\subsection{Referring Expression (RE)
Selection}\label{referring-expression-re-selection}}

Natural languages can be systematically used to keep track of previous
referents in the discourse and mark the changing cognitive statuses of
the REs (i.e., from being active to less active in the addressee's mind)
to create a more coherent narrative. Among the theoretical frameworks
that aim to explain that different linguistic forms are selected in
accordance with the addressee's mental representations of the referents
in the discourse is the Givenness Hierarchy \citep{gundel1993}.
According to this model, given or familiar information is expressed in
more attenuated forms such as pronominal use and null marking since the
mental status of the expressed referent is already at the center of
attention, as indicated by the present contextual information
\citep{chafe1976}. The Givenness Hierarchy is akin to other previously
proposed models of reference control within the topicality and saliency
paradigm such as the Accessibility Hierarchy \citep{ariel1990}, Topic
Continuity \citep{givon1983}, and Familiarity Scale \citep{prince1981}.
\citet{gundel1993} argued that the distribution of forms is not random
but abides by the Gricean Maxim of Quantity \citep{grice1975}. That is
to say, individuals must make the most informative and succinct choice
possible to retain economy of form. When multiple expressions are
available for the same referent as in an utterance like (\ref{ex:1}),
phonetically simplest form is selected with the least redundant
information are selected based on an economy principle. In the
utterance, REs for ``Eddie'' are given in bold. All three REs (nominal,
pronominal, and null marking) refer to the same entity; however, the
quantity of marking decreases as the referent's accessibility increases.
On this ground, for highly accessible referents (i.e., those that are
highly active in the mind), linguistic forms with the least phonetic
content are selected for the least amount of processing effort.

\begin{exe}
    \ex \label{ex:1}
    So \textbf{Eddie} turned around. \hfill Williams (1988; p. 343, ex. 7) \\
\textbf{He} said, "youse got a problem?" \\
"Yeah, we want you," they say.
So-$\pmb{\emptyset}$ walked right up to them \\
\end{exe}

It is also possible to categorize REs into three according to their
discourse status \citep{gullberg2006}. A referent can be introduced or
mentioned for the first time, maintained, that is, continued across two
or more clauses, or re-introduced into the discourse following a topic
shift. As a result, interlocutors mark some REs differently than others
based on the principles of reference accessibility. For spoken
languages, this observation is well established
\citep{ariel1990, givon1983, gundel1993, debreslioska2013, ahrenholz2005, chafe1976, hickmann1996, azar2015}.
These research indicate that the introduction and re-introduction of a
referent are often accomplished by full noun phrases which can be
commonly modified by an adjective or demonstrative. Maintenance, on the
other hand, is typically marked with pronominals and null markers since
the continued referent is still highly accessible to the addressee.

However, this particular observation is not peculiar to spoken
languages. There is an emerging body of literature that reports the
distribution of REs in signed narratives. Building on
\citet{gundel1993}'s work, \citet{swabey2002} demonstrated that, like
other spoken languages such as Russian, Chinese, and Japanese, American
Sign Language (ASL) signers use bare nouns for introduced and
re-introduced discourse contexts whereas null marking is used for
referents which are ``in-focus'' or very accessible to the addressee.
Overt REs like pronouns or nouns are disprefered for such referents.
\citet{swabey2002} concluded that ``native'' signers in her study abided
by the Gricean maxim of quantity and were thus economic in their
reference tracking. Similar findings have been reported for German Sign
Language (Deutsche Gebärdensprache - DGS) \citep{perniss2015} and
British Sign Language (BSL) \citep{morgan2005}. It is important to note,
however, that these research either collapsed native and late signers
into a single group and labeled them as ``native-like'' or explicitly
compared native deaf signers with late deaf signers (with language
deprivation) in terms of narrative competence.

\hypertarget{signed-vs.-spoken-res}{%
\subsubsection{Signed vs.~Spoken REs}\label{signed-vs.-spoken-res}}

Commonly in both spoken and sign languages nouns and noun phrases are
used as REs. Spoken languages use pronouns, determines and
demonstratives where sign languages use a pointing sign, usually glossed
as IX, which assumes similar functions \citep{lillo-martin1990}. In
addition, the pointing sign is used to referentially anchor a referent
to a location in signing space. For instance, when the signer introduces
\textsc{elvan} for the first time, s/he can assign her to the right.
This referential locus can later be used by agreement verbs and
classifier predicates (more below). For instance, in (\ref{ex:2a}) there
are two referents, \textsc{elvan}, who is referentially-assigned to the
right side of the signer, represented with the ``a'' subscript and
\textsc{demet}, who is referentially-assigned to the left side of the
signer, represented with the ``b'' subscript. When the signer makes a
contrast in (\ref{ex:2b}), as a pronominal use, she points to the right
side to refer to ELVAN and to the left to refer to \textsc{demet}. We
schematically represents the referential locus assigned to the referents
in (\ref{ex:3}).

\begin{exe}
    \ex 
\begin{xlist}
    \ex \label{ex:2a}
    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
    \\
    \ex \label{ex:2b}
    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
    \end{xlist}
    \end{exe}

    \begin{exe}
    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Curve Lines [id:da7006577339157659] 
\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
%Straight Lines [id:da8428827309782937] 
\draw    (112.9,128.8) -- (210.21,128.8) ;
%Curve Lines [id:da39290455666055124] 
\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
%Curve Lines [id:da9909903731270634] 
\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;

% Text Node
\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
% Text Node
\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
% Text Node
\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
% Text Node
\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
% Text Node
\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
\end{tikzpicture}
    \end{exe}

Grammatical gender as in Romance languages or noun classes as in Bantu
languages are some reference tracking tools that are familiar from
spoken languages. A somehow less known reference-tracking tool in spoken
languages is \emph{verbal classifiers}, which morphologically attach to
the verb and often cross-reference a core argument in a clause
\citep{aikhenvald2017}. Classifiers denote sets of objects based on
semantic criteria such as size, shape, body-part, texture and sometimes
function. Usually, the verbal classifier and the cross-referenced noun
co-occur in the same clause when the argument is introduced into
discourse. The classifier can track the argument on its own on
subsequent occurrences. In sign languages, a verbal classifier morpheme
is encoded by a \emph{handshape} that is articulated phonologically
co-temporally with a movement, which encodes the verbal meaning
\citep{supalla1990, benedicto2004}. These two infused morphemes form a
complex predicate. Some classifier handshapes and the set of objects
that they denote are below. Some classifier handshapes and the set of
objects that they denote are presented below in (\ref{ex:4}).

\begin{exe}
\ex \label{ex:4}
\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
\end{exe}

In (\ref{ex:5}), in the first clause, a referential locus, i.e.~locus-a,
is assigned to the locative source argument \textsc{cage}. In the second
clause, the core argument \textsc{bird} is cross-referenced with the
classifier handshape, i.e.~small-legged animal. It flies from the
referential locus of the \textsc{cage} (locus-a) to the referential
locus of the goal \textsc{tree} (locus-b).

\begin{exe}
    \ex \label{ex:5}
\begin{xlist}
    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
    \textsc{garden bird cage-a open} \\
    ‘In the garden, there is a cage and it is open.’
    \\

    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
    \textsc{bird fly[manner]-a tree-b land. CL-b}\\
    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
    \end{xlist}
    \end{exe}
    \hfill (Adapted from Gökgöz and Sevgi 2020:318)

\hypertarget{linguistic-tools-for-tracking-referents-in-signed-narratives}{%
\subsection{Linguistic Tools for Tracking Referents in Signed
Narratives}\label{linguistic-tools-for-tracking-referents-in-signed-narratives}}

The linguistic inventory for discourse referents can include nominal
constructions (bare nouns, fingerspelled nouns, and modified nouns),
null markers (plain verb, agreement verb, and constructed action),
classifier predicates (whole entity classifiers, body part classifiers,
handling classifiers, and extension classifiers), and overt pronouns
(pronominal index or IX) \citep{frederiksen2016}. This interim section
describes such constructions attested for sign languages under four main
headings (nominals, pronouns, classifiers, and null marking).

\clearpage

\begin{multicols}{2}
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:6a} Full noun \\\glll
\includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
\ex \label{ex:6b} Noun modified by classifier \\\glll
\includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
\ex \label{ex:6c} Fingerspelled noun\\\glll 
\includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
\ex \label{ex:6d} Noun modified by index (\textsc{IX})\\\glll 
\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
\end{xlist}
\end{exe}
\end{multicols}
\ea 
\ea \label{ex:7} Pronominal \\\glll
\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
\z \z
\begin{multicols}{4}
\begin{exe}
\ex \label{ex:8}
\begin{xlist}
\ex Whole entity \\\glll 
\includegraphics[width=50pt]{pictures/3e.png} \\ \textsc{fish.fall}\\ \\
\ex Body part \\\glll
\includegraphics[width=50pt]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
\ex Handling \\\glll 
\includegraphics[width=50pt]{pictures/3c.png} \\ \textsc{hold} \\ \\
\ex Extension \\\glll 
\includegraphics[width=50pt]{pictures/3d_1.png} \includegraphics[width=50pt]{pictures/3a.png}\\ \textsc{plank} {\textsc{dh: mouse}} \\ {} {\textsc{ndh: tunnel}}\\ \\
\end{xlist}
\end{exe}
\end{multicols}
\begin{multicols}{3}
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:9a} Constructed action \\\glll 
\includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\

\ex \label{ex:9b} Plain verb  \\\glll
\includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\

\ex \label{ex:9c} Agreement verb  \\\glll 
\includegraphics[width=50pt]{pictures/4b_1.png} 
\includegraphics[width=50pt]{pictures/4b_2.png} \\ {\textsc{$_{1}$hit$_{3}$}} {} \\ \\ 
\end{xlist}
\end{exe}
\end{multicols}
\footnotetext{We use small capitals to indicate sign glosses.}

\hypertarget{nominals}{%
\subsubsection{Nominals}\label{nominals}}

In TİD and many other sign languages, it is possible to use bare noun
phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in
(\ref{ex:6c}). Fingerspelling refers to the use of the manual alphabet
which is derived or borrowed from a surrounding written language
\citep{kubus2008}. Nouns can also be modified by classifiers as in
(\ref{ex:6b}) \citep{frederiksen2016} or by a prenominal or postnominal
IX functioning as a determiner or a demonstrative as in (\ref{ex:6d})
\citep{maclaughlin1997}.

\hypertarget{pronouns}{%
\subsubsection{Pronouns}\label{pronouns}}

The IX sign (extended index finger) has various functions in sign
languages, one of which is the pronominal use which refers to
established nominals by pointing to a previously assigned abstract or
physical locus in the signing space when re-introducing a topic in
discourse \citep{emmorey1996}. The same IX function has also been
identified for TİD \citep{sevinc2006}. The example in (\ref{ex:7})
illustrates a spatially-anchored pronominal reference for topic
re-introduction. These signs are thus rarely used for the first mention
of an entity \citep{perniss2015}. Other interpretations consider
pointing signs without noun phrases as examples of demonstratives and
not personal pronouns \citep{koulidobrova2016}. It is important to
mention here that in some spoken languages (e.g., German and Dutch),
demonstratives can also refer to persons \citep{bosch2007} and behave
similarly to personal pronouns. However, for the purposes of the present
study, we will refer to IX constructions as pronominal IX and
particularly examine their non-deictic or personal pronoun realizations
as a possible discourse strategy to mainly mark topic shift among native
and late signers of TİD.

\hypertarget{classifiers}{%
\subsubsection{Classifiers}\label{classifiers}}

Classifier predicates are highly complex morphological constructions
which are very common in sign languages \citep{supalla1986}. These are
manually signed iconic representations of animate or inanimate objects,
and they embody several properties pertaining to the nominal referent
including its location in the signing space, movement, path, size, and
shape \citep{talmy2003}. As in most sign languages hitherto studied,
classifier constructions are also common and found in TİD \citep[see][
for a review]{dikyuva2015}. Following the classification of
\citet{engberg-pedersen1993}, we distinguish between four types of
classifier constructions. Whole entity classifiers (WECL) represent the
entirety of any referent whereas body part classifiers (BPCL) only
reflect a specific part of an animate entity (e.g., limbs, head, tongue,
teeth, and so forth). Handling classifiers (HCL) occur with transitive
verbs and depict how an object is handled or manipulated by agent. On
this ground, an HCL\footnote{As discussed before, an HCL is transitive
  in nature and depicts the handling action performed by ananimate
  agent. This is why most research tend to consider HCLs as an extension
  to constructed action \citep{dudis2004, cormier2013}.} depicts the
manner of the referring action while the former two constructions encode
the location of the entity as well as its motion in space. Extension
classifiers (ExtCL), on the other hand, denote the size and shape of an
entity by following a trajectory to trace its shape. We exemplify these
constructions in (\ref{ex:8}).

\hypertarget{null-marking}{%
\subsubsection{Null marking}\label{null-marking}}

Parallel to spoken languages, many sign languages have been reported to
allow null subjects and/or objects and make use of null arguments when
referring to highly accessible referents which are still active in the
mental representation of the addressee. In signed discourse, this can be
achieved by employing certain verbal constructions or constructed action
(CA). The latter construction exemplified in (\ref{ex:9a}) refers to
cases in which the signers engage with the space in front of themselves
as if they are interacting with the objects and characters in the story.
This is often achieved with the signer taking on the role of a referent
in the narrative and performing the actions \citep{metzger1995}. CA is
often marked by certain non-manual cues such as the facial expression
and the breaking of the eye contact. This is to say, the narrator's
head, face, torso, arms, and as an optional marker, hands represent
those of the associated referent(s) \citep{smith2014}. As for the verbal
constructions, to identify the predicate types in ASL,
\citet{padden1986} suggested a triadic categorization of verbs in sign
languages: plain verbs, agreement verbs, and spatial verbs. According to
this classification, plain verbs (\ref{ex:9b}) do not agree with the
subject or the object. This means that they are not inflected for number
or person. This contrasts with agreeing verbs (\ref{ex:9c}) the starting
or the ending point of which can be modified to specific loci for person
and number agreement while spatial verbs (as in \textsc{ix book put} `I
placed the book there') agree with locative arguments but like plain
verbs, they are not inflected for number or person. TİD too licenses
null arguments. Following this classification, \citet{sevinc2006} argued
that although it is common in TİD to drop arguments, there were certain
limitations on the null marking of the arguments for agreement and plain
verbs.

\hypertarget{learning-how-to-track-referents-in-sign-languages}{%
\subsection{Learning How to Track Referents in Sign
Languages}\label{learning-how-to-track-referents-in-sign-languages}}

In both communication modalities (visual and auditory), children start
to produce more complex and coherent stories with appropriate and
informative linguistic forms to indicate anaphoric or cataphoric
references after age 10 \citep{hickmann1996, rathmann2007}. Prior to
this age, both signing and speaking children fail to refer to entities
consistently since their pragmatic abilities are yet to develop. The
manipulation of linguistic forms for discourse functions to signal topic
continuance or shift is difficult to master until the beginning of
formal education. These findings are also comparable across different
sign languages including BSL \citep{morgan2005}, ASL \citep{loew1984},
and DGS \citep{becker2009}, pointing out to the fact that the mastery of
reference tracking in narratives is one of the most protracted
achievements in the acquisition of signed discourse.

Like in first language development, narrative skills have been
extensively studied among second language learners
\citep[see][]{bel2015, frederiksen2019}. Comparing 11 deaf adults
signers and 13 hearing advanced learners of Catalan Sign Language
(Llengua de Signes Catalana - LSC), \citet{bel2015} found that hearing
signers resort to more overt forms (nominals and pronouns) than native
signers to avoid potential ambiguities among referents. Nevertheless,
their results did not reach statistical significant for the interaction
of discourse status and acquisition group. A similar line of research
came from \citet{frederiksen2019} who found a statistically significant
difference between both groups, that is native signers and hearing L2
learners of ASL, only for classifier use especially in re-introduced
contexts. L2 learners in the study used more nominals and fewer
instances of null marking in number, indicative of a limited effect of
overexplicitness among second language learners. However, these studies
treated late and native deaf signers as a single group. That is, the
control group consisted of both native signers who had one or more deaf
parents and also native-like \citep{cormier2012} late deaf signers who
had hearing parents and their age of acquisition varied. Given the
preliminary evidence that topic continuity and salience marking is also
difficult for second language learners regardless of modality, it has
been suggested under the Interface Hypothesis that the syntax-discourse
interface is vulnerable and especially challenging to acquire
\citep{sorace2011} for second language learners. Similarly,
\citet{sorace2009} discuss that this interface entails much more than
morphosyntactic knowledge but also includes cognitive and pragmatic
resources that help provide discourse coherence. They propose that the
second language discourse is over-redundant. This means that second
language learners produce more overt pronouns or fuller forms of
language for highly accessible referents possibly as a compensatory
strategy in an attempt to reduce the increasing load on cognitive
processing.

Still, only very few research examined whether linguistically deprived
signers (i.e., deaf signers with delayed exposure to a sign language)
also display redundancy in their reference tracking. \citet{becker2009}
examined 6 DGS signing deaf children (3 native and 3 late)`s narrative
abilities. She reported that late signing children generated more
explicit forms such as nominals when referring to the characters in the
story compared to native learners, concluding that first language
deprivation results in an inadequate representation of the addressee and
less efficient use of morphosyntax. Comparably, \citet{cormier2013}
investigated the use of CA in 15 BSL among deaf native and late adult
signers and found that late signers were redundant in topic continuity.
In other words, the late signers produced more explicit forms like
nominals, and fewer instances of null marking when maintaining previous
referents. Most relevantly to the present study, \citet{gur2018} studied
the development of narrative skills among 10 native and 10 late deaf
children and deaf adult signers of TİD. She found that native-signing
children and adults had more instances of explicit referent
introduction, set the scene more frequently, and employed CA in an
appropriate manner more than corresponding late-signing children and
adults. Although this study presented preliminary findings on age of
acquisition effects on narrative competence, it does not investigate how
native and late signers employ different linguistic tools based on the
accessibility of the referent. Therefore, to our knowledge, there is a
lack of research directly examining the effects of first language
deprivation on deaf adult signers' selection of REs in signed
narratives.

Considering this, it is not very clear whether the negative effects on
narrative abilities observed among late signing deaf children hold when
they become adults. Understanding the impact of language deprivation or
delayed first language learning on narrative competence is crucial for
three reasons. First, studying this population will shed light on the
possible effect of lack of early language experience on future cognitive
and pragmatic abilities since the narrative competence is dependent on
such faculties. Second, examining reference tracking strategies in
particular will also inform about the linguistic inventory that native
and late deaf signers use to refer to entities. Finally, any possible
difference between the two acquisition groups in terms of narrative
competence will have further practical implications for the importance
of early access to a first functional sign language to retain future
linguistic and cognitive abilities.

\hypertarget{present-study}{%
\subsection{Present Study}\label{present-study}}

In an attempt at addressing this gap in the literature, the present
study aims to investigate the reference tracking strategies of native
and late adult signers in TİD narratives with a story-telling paradigm.
We explore the forms available for referent tracking in TİD and examine
the nature of RE selection as a possible indicator of pragmatic
competence between our acquisition groups.

Overall, considering what has been suggested for the distribution of REs
in signed narratives, we expect more explicit forms like nominals for
referent introduction and and less explicit forms such as null marking,
pronominal and/or classifier use for referent maintainance. For the
re-introduction contexts, where referents which are discontinued across
one or more clauses are brought back into the discourse, we expect a
less clear picture. Some re-introduced referents will have higher
accessibility and will thus be tracked with more implicit REs, while
others that have lower accessibility will be tracked with more explicit
REs Also based on previous research, we hypothesize that late signers
will become more over-redundant in their reference tracking compared to
native signers. This would translate to a preference by late signers for
nominals over null marking especially in maintenance and re-introduction
contexts. In contrast, we would expect that native signers use more null
marking to track referents.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We tested 29 deaf adult signers, 15 with exposure to TİD from birth
(i.e., native signers), 14 with exposure between ages 3 and 17 (i.e.,
late signers) (\emph{M}\textsubscript{Age of acquisition} = 7.7 years,
\emph{SD} = 3.3 years). The criterion to be classified as a late signer
was to have hearing and non-signing caregivers. All the signers
satisfying this criterion were regarded as late signers
\footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers. Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3. In order to capture in-group variation among late signers, we run an additional supplementary analysis}.
All participants indicated that their preferred language of
communication was TİD which they used for most contexts. Detailed
information about the linguistic and educational background of the
participants can be found in \ref{section:a}.

All participating native signers (7 males, 8 females) were born into
deaf families and started acquiring TİD from at least one of their
caregivers after birth. The parents of 14 participating native signers
were deaf. One native signer had a deaf father and a hearing mother. In
addition to TİD, the home language for some native signers (n = 4) in
early childhood included the use of home
sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers. Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}.
The age of the native participants ranged between 18 and 35 (\emph{M} =
26.4 years, \emph{SD} = 5.0 years). This also corresponded to the total
average years of TİD use for native signers. All native signers were
pre-lingually deaf, that is they were either born deaf or they had
become deaf before the age of 3, with the exception of one signer who
became deaf between the ages of 4 and 7. Native signers reported to
complete at least high school education and their level of education
(LOE) was between 12 and 16 years (\emph{M} = 12.7 years, \emph{SD} =
1.5 years). In addition, all native participants attended at least two
deaf schools, one being primary and the other middle school. 10 native
participants also attended a deaf school for their high school
education. Based on a self-reported 5-point scale of proficiency (1:
poor, 5: proficient), native participants declared that they have high
receptive and productive skills in TİD (\emph{M} = 5 points, \emph{SD} =
0 points). As for their proficiency in (spoken) Turkish, their
self-reported rankings were on average 3.7 (\emph{SD} = 0.7), 3.7
(\emph{SD} = 0.7), and 2.6 (\emph{SD} = 1.2) for reading, writing, and
speaking abilities, respectively.

Comparably, late learners (8 males, 6 females) were all born into
hearing families and started learning TİD later in life after age 3,
usually with the start of formal education. The age of sign language
acquisition for late signers ranged between years 3-17 (\emph{M} = 8.7
years, \emph{SD} = 3.5 years). The parents of all participating late
signers were hearing, and the reported home languages used in childhood
mainly consisted of Turkish and home sign. The average age of late
signers was 34.6 years (\emph{SD} = 8 years) and ranged between 24 and
50. The derived mean number of years for TİD use following its
acquisition was comparable to native signers with 25.9 years (\emph{SD}
= 7.9 years). Except for two signers who became deaf between the ages of
4 and 7, all the other late signers were pre-lingually deaf. All late
signers completed middle school education and the average year of
schooling was 12.3 years (\emph{SD} = 2.8 years). All late participants
attended at least one deaf school. While 8 participants attended two
deaf schools (either primary and middle school, or middle and high
school), 3 participants reported attend three deaf schools. Using the
same scale, late signers self-rated their abilities in TİD and (spoken)
Turkish. Both of their receptive and productive TİD abilities rated on
average 4.9 points (\emph{SD} = 0.3 points). Late signers' mean rating
for their reading abilities in Turkish was 3.5 points (\emph{SD} = 0.7
points) while their writing and speaking abilities rated on average 3.4
(\emph{SD} = 0.5 points) and 3.1 points (\emph{SD} = 0.7 points),
respectively.

Native and late participants' sex, average years of schooling, TİD use
following its acquisition, and their self-reported receptive and
productive TİD ratings, and their reading, writing, and speaking ratings
for Turkish did not statistically differ; Bayesian \emph{p}'s
\textgreater{} 0.1. However, participants' ages statistically differed
between the two acquisition groups; \emph{t}(27) = 3.4, Bayesian
\emph{p} \textless{} 0.01. Overall, the native group consisted of
younger participants. This was to ensure that the average years of TİD
use among native signers approximated those of late signers.

\hypertarget{stimuli}{%
\subsection{Stimuli}\label{stimuli}}

We used a story-telling paradigm to examine how native and late signers
of TİD track reference in signed discourse. Participants were presented
with 10 short video clips extracted from the cartoon \emph{Tom and
Jerry} and asked to retell them. Each clip had six to ten episodes
(\emph{M} = 8 episodes). The duration of the clips ranged from 17
seconds to 28 seconds (\emph{M} = 22.7 seconds). The narratives included
on average 2 recurring animate characters for which the present context
required the introduction, maintenance, and re-introduction of
referents. In addition, in each narrative, there were on average 2
inanimate objects with which the animate referents interacted. The
detailed description of the episodes in each narrative can be found in
\ref{section:b}.

\begin{table}[p]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:tab-2}Annotation Tiers with their Tags and Descriptions}

\footnotesize{

\begin{tabular}{p{3.2cm}p{3.8cm}p{3.8cm}}
\toprule
Tier & \multicolumn{1}{c}{Tag} & \multicolumn{1}{c}{Description}\\
\midrule
Clause Boundary &  & The boundary between clauses indicated by the presence of predicates and/or certain prosodic cues (e.g., head nod)\\
Discourse Status & Introduced & A first mention of an entity in a given discourse regardless of any syntactic role\\
 & Maintained & A subject referent continued across two or more immediate clauses\\
 & Reintroduced & A subject referent discontinued across at least one clause\\
REAT Type & NOM: Bare Noun & A non-modified noun phrase (e.g., CAT, MOUSE)\\
 & NOM: Modified Noun & A noun phrase modified by an index sign or an adjective\\
 & NOM: Fingerspelled Noun & Borrowed nouns from Turkish spelled by the manual alphabet (e.g., K-E-D-İ ‘cat’)\\
 & PRO: Pronominal IX & An index sign used in isolation to refer to an entity in discourse\\
 & CL: Whole Entity Classifier & A manual construction that refers to the entirety of an entity\\
 & CL: Body Part Classifier & A manual construction that refers to a part of an entity\\
 & CL: Handling Classifier & A manual construction that refers to the handling of an item\\
 & CL: Extension Classifier & A manual construction that denotes the size and shape of an entity\\
 & NULL: Constructed Action & A multifunctional construction where the signers take on the role of a referent and perform their actions\\
 & NULL: Plain Verb & A non-agreeing verb (e.g., LOVE)\\
 & NULL: Agreement Verb & A verbal construction the beginning and ending of which agree with person and number\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{task-procedure}{%
\subsection{Task Procedure}\label{task-procedure}}

Participants sat in front of a laptop computer which had the stimulus
stories. The instructions were all given by a deaf research assistant
who was present in all sessions and ensured that the task requirements
were clear. The informed consent form was translated into TİD and
explained to the participants by the same assistant before data
recording. Participants filled in the consent form and a background
questionnaire form before the onset of the trials. Following this step,
signers were seated across a hearing researcher who is an advanced late
learner of TİD. We instructed the participants to narrate as they would
to a deaf friend. All trials were recorded by a digital camera placed in
front of the signers. The participants were instructed to watch the
narrative clip as many times as they needed, and narrate the content of
each clip either to the experimenter or the recording digital
camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons. First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter. Therefore, most signers preferred to narrate to the camera. Furhermore, some signers did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available. We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers (see \ref{section:c} The results imply that there is not much of a difference.}.

\hypertarget{data-annotation}{%
\subsection{Data Annotation}\label{data-annotation}}

We coded the discourse status (Introduction, Maintenance,
Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null). We
used ELAN Linguistic Annotation Software \citep{crasborn2008} for data
annotation. First Author who is a hearing signer of TİD annotated all
the data and translated the narratives into Turkish together with a
native deaf signer of TİD.

Following \citet{berman2013relating}, we first identified the clause
boundaries in each sentence indicated by the presence of predicates and
informative non-manual markers (e.g., head nod). On the second tier, we
coded the discourse status for each referent. Having adapted the
classification of \citet{gullberg2006}, we considered all referents
mentioned for the first time as introduced regardless of their syntactic
function in the sentence. Only subjects were coded for maintenance and
re-introduction. We coded subject referents as maintained only if they
had been referred to in the previous clause. To this end, the referents
which had been discontinued across one or more sentences were all
regarded re-introduced into the discourse. On a third tier, we had the
RE type for a single reference with the main categories nominal,
pronominal, null marking, and classifier. Under nominal, we identified
bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we
distinguished among three types of constructions: WECL, BPCL, and ExtCL.
Finally, we had CA, agreement, plain tags for null constructions. We
were unable to observe any instance of HCLs occurring in isolation
(i.e., not part of CA). The annotation tiers are presented in Table
\ref{tab:tab-2}.

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

In our analysis, we use Bayesian regression models with the help of the
brms package \citep{burkner2018} in R. In a Bayesian regression the
prior and the likelihood is used to calculate the posterior. Instead of
a single filter of significance, we get a posterior probability
distribution. We do not define our own priors and use the brms package
defaults. We are using Bayesian regression to avoid some common pitfalls
of other approaches \citep{wagenmakers2007, kruschke2011}.

For this particular type of experimental data, we fit the responses to a
multinomial distribution. Most of the time, experimental results have
yes/no-true/false responses which are fit to binomial distributions,
reading/response times which are fit to log-normal distributions, and
ratings which are fit to cumulative distributions. In these cases, the
researcher is not particularly interested in the `intercept' coefficient
which shows the mean value of the response. The researcher is usually
interested in the remaining coefficients which give information about
predictor level comparisons.

In a multinomial distribution, the responses are in a categorical
relationship where the levels can be A,B,C, and more. These levels are
not ordered in time, they are not cumulative, and they are not additive.
If we were to make an analogy, they are like countries on Earth which
can have differences and similarities, but they don't have to. In a
multinomial distribution, one of the responses becomes the level of
comparison and other levels are compared against it. In a way, there are
multiple binomial distributions. However, the intercepts in a
multinomial distribution end up having more of an interest for the
researcher. This is because the intercept is a comparison of two levels
of the response to the exclusion of the remaining response levels. There
is another consideration that should go into interpreting the results of
a multinomial distribution. Different from the intercepts, the remaining
coefficients are predictor level comparisons within the grand comparison
of the response levels. To indicate which predictor level comparison
belongs to which response comparison, we name our coefficients with
prefixes followed by an underscore.

In our model plots, we report median estimates as points, 50\% credible
intervals as a thick line and 95\% credible intervals as a thin line.
These numbers are relatively arbitrary, and they are intended to show
the posterior probability distribution of the coefficients. We use
interval plots to save space and have a more compact plot. If a
considerable amount (\textgreater90\%) of the posterior probability
distribution is towards a sign (-/+) we interpret it as a categorial
effect, if a good amount (\textgreater50\%) of the distribution is
towards a sign (-/+) we interpret it as a tendency, and we see the
remaining distributions as indifference. In addition to these, we look
at how the distribution skewed in the extremes or how `wide' it is from
the median estimate to inform us mainly about uncertainty and variation.

\hypertarget{results}{%
\section{Results}\label{results}}

We collected 4142 data points as a result of the experiment. Each data
point is a referent tracked by the participant. Each of the REs is
classified in terms of its Discourse Status (Introduction, Maintenance,
Re-introduction), its Type (Nominal, Pronominal, Classifer,
Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null). However, we allude to the numerical distribution of the subtypes in this and following sections},
and Participant Group (Native,
Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17). Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}.

Across our acquisition groups, we observed 1066 (0.26) instances of
referent introduction, 1651 (0.4) instances of referent maintenance, and
1425 (0.34) instances of referent re-introduction. For the introduction
of a referent, the majority of the REs were nominal constructions (Table
\ref{tab:my-table}). However, we observed a sporadic use of classifiers
and null marking respectively. When the discourse status was maintained,
signers extensively used null markers to refer to the entities in the
immediately previous clause. Classifier constructions followed null
markers in frequency in maintained contexts. We observed only a few
instances of nominal use for referent maintenance. In re-introduction
contexts, signers used a greater proportion of nominal markings which
were followed by null, classifier, and pronominal use.

\begin{table}
\small
\caption{Distribution of all discourse status by RE types}
\label{tab:my-table}
\begin{tabular}{llll}
\hline
                     & Introduced    & Maintained     & Re-introduced \\ \hline
\textit{Nominal:}    & 92.78\% (989) & 5.63\% (93)    & 63.65\% (907) \\
\textit{Pronominal:} & 0.09\% (1)    & 0.24\% (4)     & 2.46\% (35)   \\
\textit{Classifier:} & 5.35\% (57)   & 18.29\% (302)  & 8\% (114)     \\
\textit{Null:}       & 1.78\% (19)   & 75.83\% (1252) & 25.89\% (369) \\ \hline
\end{tabular}%
\end{table}

Table \ref{tab:my-table2} shows that for all discourse statuses the most
numerous nominal type was \textsc{bare nom} which was followed by nouns
modified by index or classifiers. In number, signers used
\textsc{mod nom} more in introduction and re-introduction compared to
maintenance settings. Signers produced a total number of 40 pronouns
(Table \ref{tab:my-table3}). A great majority of the pronominal
constructions were used to re-introduce a referent into the discourse.
We observed only a very small amount of pronoun use for referent
introduction and maintenance.

\begin{table}
\small
\caption{Distribution of nominals by discourse status}
\label{tab:my-table2}
\begin{tabular}{lllll}
\hline
                        & Bare Noun  & Modified Noun & FS Noun \\ \hline
\textit{Introduced:}    & 48.93\% (842) & 53.56\% (139)     & 100\% (8) \\
\textit{Maintained:}    & 4.82\% (83)  & 3.85\% (10)      & 0\% (0) \\
\textit{Re-introduced:} & 46.25\% (796) & 42.69\% (111)     & 0\% (0) \\ \hline
\end{tabular}
\end{table}

\begin{table}
\small
\caption{Distribution of pronominals by discourse status}
\label{tab:my-table3}
\begin{tabular}{ll}
\hline
                        & Pronominal  \\ \hline
\textit{Introduced:}    & 2.5\% (1)   \\
\textit{Maintained:}    & 10\% (4)    \\
\textit{Re-introduced:} & 87.5\% (35) \\ \hline
\end{tabular}
\end{table}

As for the classifier constructions (Table \ref{tab:my-table4}), WECL
was the most common type, followed by BPCL and ExtCL. For the first
mention of a referent, signers mostly preferred to use ExtCL. However,
for referent maintenance and re-introduction WECL was used more compared
to BPCL and ExtCL.

Among the null markers, CA was most frequently employed, and it was
followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
Overall, signers used very few null markings for the introduction of a
referent and the majority of the existing ones were CA. Additionally,
signers used CA more for referent re-introduction than maintenance. We
observed a similar instances of agreement verbs for maintained and
re-introduced contexts. In contrast, plain verbs were favored more in
the maintenance setting than re-introduction.

\begin{table}
\small
\caption{Distribution of classifiers by discourse status}
\label{tab:my-table4}
\begin{tabular}{llll}
\hline
                        & WECL          & BPCL        & ExtCL         \\ \hline
\textit{Introduced:}    & 2.14\% (6)    & 2.4\% (3)   & 71.64\% (48) \\
\textit{Maintained:}    & 67.26\% (189) & 77.6\% (97) & 23.88\% (16) \\
\textit{Re-introduced:} & 30.6\% (86)   & 20\% (25)   & 4.48\% (3)   \\ \hline
\end{tabular}
\end{table}

Across the discourse statuses, there is also a difference between
participant groups in the use of null marking for reference tracking.
Native signers of TİD used it more (0.23) compared to the late signers
(0.16). Comparably, late signers used nominals more frequently (0.25)
than native signers (0.23). Figure 1 shows the numerical distribution of
the responses by RE type and Group.

\begin{table}
\small
\caption{Distribution of null markers by discourse status}
\label{tab:my-table5}
\begin{tabular}{llll}
\hline
                        & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
\textit{Introduced:}    & 1.58\% (17)        & 0.28\% (1)    & 0.5\% (1)      \\
\textit{Maintained:}    & 71.64\% (773)      & 89.97\% (323) & 77.23\% (156)  \\
\textit{Re-introduced:} & 26.78\% (289)      & 9.75\% (35)   & 22.28\% (45)   \\ \hline
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{manuscript_v2_files/figure-latex/count-plot-1.pdf}
\caption{Referent counts by Discourse Status (Introduction, Maintanence,
Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and
Participant Group (Native, Late)}
\end{figure}

For more inference on the data and having pairwise comparisons, we fit a
regression model to RE Type using discourse status and acquisition group
as predictors. We used the brms package \citep{burkner2018} in R, with a
categorical distribution on the responses and sum contrasts for the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late). As the overall number of produced
pronouns was very little (40 data points), we excluded pronominal use
from the following analyses. For the current analysis, we did not define
priors ourselves and used the preset uninformed priors.
@ref\{fig:model-gram\}\footnote{We ran another model where we used the
  Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and
  7-above:Late2. However, our model did not converge properly. We
  provide the model results as a plot and table in the \ref{section:d}.}
shows the posterior probability distributions where the dot is the
median estimate, the thick line is the 50\% credible intervals, and the
thin line is the 95\% credible intervals. There are 3 responses we are
considering in this model. They are; Classifier, Nominative, and Null.
Classifier response is the level that other responses Nominative and
Null are compared against.

\begin{verbatim}
## Warning: Ignoring unknown aesthetics: height
\end{verbatim}

\begin{figure}
\centering
\includegraphics{manuscript_v2_files/figure-latex/model-gram-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Null) with the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late)}
\end{figure}

In a multinomial distribution, the model takes one of the responses as a
reference and compares all other responses to it. For example, the first
coefficient, Nom\_Intercept (nominal intercept), in Figure 2 is the
comparison of a nominal versus a classifier. It means that participants
produced more nominals than classifiers in general regardless of
participant group and discourse status. The second coefficient shows
that null markings were also used more than classifiers by the
participants in general, but not as much as nominals.

In the comparison of nominals and classifiers, being a native acquirer
of TİD did not have an effect. Nominals were used more when the
discourse status was introduction compared to re-introduction. Again,
the whole comparison is made against a classifier. In contrast to the
introduction status, nominals were used less in maintenance contexts
compared to the re-introduction contexts within the greater comparison
to classifier referents. There seems to be an interaction between the
introduction status of discourse and being a native acquirer of TİD.
When the acquirer is a native speaker and the discourse function is
introduction, the chances of observing a nominal referent gets even
higher. Also, there seems to be an interaction between maintaining a
referent and being a native acquirer of TİD. When it was a native
acquirer of TİD and the discourse context was maintenance, the chances
of observing a nominal referent gets even lower.

Compared to classifiers, being a native acquirer of TİD increased the
odds of using null marking in general. Null constructions were used less
in introduction contexts than in re-introduction. This contrasts with
the nominals (4th coefficient) as they were used more in referent
introduction versus re-introduction. In contrast to introduction
contexts, null markers were used more in referent maintenance compared
to re-introduction, yet again within the greater comparison to the
classifiers. There seems to be an interaction between introducing a
referent and being a native acquirer of TİD. In that, the odds of using
null marking increases if the participant is a native acquirer of TİD
and the referent is used in an introduction context. There is another
interaction with regards to referent maintenance, contrasting the first
interaction. In that, the odds of using null markers decrease if the
participant is a native acquirer of TİD and the referent is maintained.
This does not mean that late signers used more null markers in
maintenance contexts. It means that the use of null markers did not
linearly increase as the addition of using a null marker instead of a
classifier as a native plus using null markers instead of a classifier
for maintaining a referent versus re-introducing it.

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

In our study, we aimed to (i) describe how signers track referents in
signed discourse in TİD and (ii) examine whether language deprivation
results in more over-redundant referent selection. To this end, we
investigated the distribution of referring expressions (REs) that the
native and late signers select when employing referents in different
discourse statuses (introduced, maintained, and re-introduced).

In our first set of hypotheses regarding referent tracking strategies,
we had the general expectation that signers would use more explicit
forms (mainly nominals) for referent introduction and less implicit
forms such as null marking, pronominal use, and classifiers. In
re-introduced contexts, we predicted that signers would use both
explicit and implicit markers depending on the accessibility of the
tracked referent.

For age of acquisition effects, based on what is proposed for child and
adult deaf signers with language deprivation in terms of their
morphosyntactic \citep{boudreault2006, cormier2012, kayabasi2020} and
narrative development \citep{becker2009, cormier2013}, we expected that
late signers would be more over-redundant in their referent tracking,
akin to second language learners \citep{frederiksen2019, bel2015}. That
is, we had the expectation that late signers would use more explicit
markers like nouns and fewer implicit markers like null marking even for
highly accessible referents (when the discourse status is maintained or
even re-introduced in same cases) than the control group (i.e., native
signers).

\hypertarget{tracking-referents-in-tid}{%
\subsection{Tracking referents in TİD}\label{tracking-referents-in-tid}}

The results of the present study partly support our hypotheses. In line
with our expectations, regardless of their acquisition groups (native or
late) (i) signers mainly used nominals to introduce a referent into the
discourse, (ii) used null marking and classifiers but not pronouns for
maintained contexts, and (iii) employed both nominals and less explicit
markers (classifiers, null markers and pronouns) to re-introduce a
referent.

However, the proportion of the observed discourse statuses (26\% for
introduction, 40\% for maintenance, 34\% for re-introduction across
acquisition groups) in this study differed from that of some other
studies. For a comparison, while the percentage of introduced referents
were similar in both research (26\% versus 24\%), the participants in
the present study brought back referents into the discourse more
frequently compared to the participants in \citet{frederiksen2019} and
\citet{frederiksen2016} which reported that re-introduction contexts
accounted for only 7\% of the referents in ASL. On the other hand,
maintenance percentages were lower in the present study compared to
\citet{frederiksen2016} (40\% versus 69\%). Investigating the
distribution of REs across discourse statuses in ASL, a similar study by
\citet{czubek2017} also reported contrasting numbers (17\% for
introduction, 64\% for maintenance, and 19\% for re-introduction). We
suggest that the discrepancy might be due to the nature of the stimuli
used in these research. In \citet{frederiksen2016}, the authors used
4-6-episode-long picture story with 2 animate characters and 1 inanimate
object. In contrast, the stimuli in the present study consisted of
8-episode-long video clips on average with a mean number of 2 animate
characters and 3 inanimate objects. Although the number of animate
characters and inanimate characters in \citet{czubek2017}'s stimuli
matched that of the present study, their narrative structure (a
6-episode-long picture story) was still shorter compared to our
narrative stimuli. We believe fewer possible referents and/or a simpler
narrative structure in these studies may have contributed to the lower
number of re-introduced contexts compared to the present study. Due to
the story length and composition in this study, the participants had to
track more recurring animate or inanimate referents for a longer period
of time. As a result, our signers engaged in referent re-introduction
more than what was reported in previous studies
\citep{czubek2017, frederiksen2016}.

As for the distribution of REs across different discourses statuses, we
observed that signers overwhelmingly preferred to use nominals over
other REs for the first mention of a referent and used null markers such
as CA, and agreement and plain verb for referents with high
accessibility (i.e., when the discourse status was maintained). This
finding parallels other accounts of referential accessibility suggested
for other sign languages {[}see \citet{frederiksen2016};
\citet{czubek2017}; and \citet{swabey2002}; for ASL, \citet{morgan2005};
for BSL \citet{bel2015}; and \citet{barbera2009} for LSC. Our results
also indicate that modified nouns (e.g., nouns that are preceded or
followed by a classifier or IX) appeared more in introduced contexts
compared to re-introduced contexts. This brings along the assumption
that modified nouns are used more to track referents with lower
accessibility, probably due to their more explicit structure.

Interestingly and to our surprise, the signers in this study used
pronominal IX very sporadically with only 40 observed instances that
were mainly used to re-introduce referents rather than maintain them.
This finding is comparable with what has been reported for ASL
\citep{swabey2002, frederiksen2016, czubek2017} but not with LSC
\citep{bel2015}, BSL \citep{morgan2005} and DGS \citep{perniss2015}, in
which the signers extensively made overt pronominal references (up to
30\%) to maintain entities. Comparable to our findings, It has been
reported that pronouns do not frequently occur in the local context and
instead follows a topic shift and thus rank higher than classifiers and
null markers different genre \citetext{\citealp[see][ for
LSC]{barbera2009}; \citealp[and][ for TİD]{nuhbalaoglu2018}}.
Considering this, the claim that the IX sign functions like
demonstratives that are also used to track referents in re-introduced
contexts has not been borne out \citep{koulidobrova2016}. Similar to
these research, our findings indicate that the pronominal IX in TİD was
used more concurrently with nominal constructions and pronominal IX was
used very infrequently in isolation. The observation that overt pronouns
are rare in ASL narratives comes from studies that include a simple
picture retelling task with one protagonist
\citep[e.g.,][]{frederiksen2016} as well as a more complex narrative
design as in \citep{czubek2017}. However, \citet{czubek2017}'s study
also incorporated different genres and reported more instances of overt
pronouns in personal narratives (compared to picture or video
retelling). In fact, this finding suggests that the use of pronouns
might be genre-dependent. Moreover, another production study on referent
tracking in ASL reported object preference for pronominals
\citep{frederiksen2017}. Nonetheless, we only coded subjects for
maintained and re-introduced referents. We argue that all of these
factors (genre-dependency as well as object preference) might contribute
to the explanation of why overt pronouns in this data are dispreferred
in favor of other markers for highly accessible referents such as null
marking and classifiers.

A deeper look into the allocation of classifier constructions revealed
that WECLs were the most common type of classifiers for the maintenance
of a referent followed by BPCLs and ExtCLs. We observed a similar
alignment for re-introduced contexts. In contrast, a great number of the
overall ExtCLs were used to introduce a new referent into the discourse.
That is to say, whereas WECLs were the preferred classifier types to
signal highly accessible referents, ExtCLs constructions were mainly
used for referents which were not easily accessible. We suggest that the
less iconic and transparent nature of WECLs allowed signers to continue
a previous referent without being redundant. As pointed out by
\citet{czubek2017}, signers prefer to rely on ExtCLs for the first
mention of an entity because these constructions are informative and
explicit by nature, and they provide visual and descriptive cues about
an entity and its location in space. Nevertheless, we only observed 3
instances of ExtCLs for re-introduced contexts, replicating
\citet{frederiksen2019}'s findings for ASL.

As for the distribution of the null markers, approximately 90\% of the
plain verbs were used to maintain a previous referent, making these
constructions least implicit REs used for TİD\footnote{As in other sign
  languages, in TİD, the lexical meaning of signs has an effect on the
  morphological grouping. For instance, usually (di)transitive agreement
  verbs denote some sort of physical or metaphorical transfer of an
  object or abstract notion (Meir 1999). As for plain verbs, both
  phonology of a sign and its semantics have an effect on morphology.
  For instance, \textsc{eat} in TİD is signed nearby the mouth and since
  it needs to remain in that phonological location, it is not suitable
  for agreement and thus it is regarded as a plain verb. Likewise the
  experiencer verb \textsc{love} is signed on the torso and as such it
  does not take part in agreement. These generalizations noted we should
  also be cautious in labelling a verb as categorially one type or the
  other. For instance, some intransitive plain verbs like \textsc{fall}
  or \textsc{jump} in TİD may display spatial person agreement with the
  single argument in the clause (Gökgöz and Sevgi, 2020). See also for a
  similar discussion in Brazilian Sign Language.\}.}. Agreement verbs
and CA were the following implicit markers of reference, 0.77 and 0.72
of which were used for maintained contexts, respectively. This contrasts
with some of the research \citep{czubek2017} in which agreement verbs
ranked lowest on scale, indicating that these constructions were more
implicit compared to plain verbs. However, there seems to be a common
ground for the position of CA markers which have been claimed to rank
higher than other null markers as in the present study
\citep{frederiksen2016, czubek2017, barbera2009}. In Figure
\ref{fig:scale}, we report the distribution of REs in TİD as a function
of referent accessibility and compare our accessibility hierarchy with
that of \citep{frederiksen2016} for ASL. In the hierarchies, the left
most REs (e.g., modified and bare nouns for TİD) are used when the
accessibility is low. If the referent is already active in the
addressee's mind and its accessibility is high, then the right most
markers (e.g., agreement and plain verbs for TİD) are selected.

\begin{figure}
\caption{Comparing the present study and Frederikse and Mayberry (2016) in terms of proposed accessibility hierarchies}
\label{fig:scale}
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16) -- (214.92,115.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15) -- (314.89,115.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34) -- (414.89,115.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize M NOM}};
% Text Node
\draw (191,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize BARE NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node
\draw (285,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize WECL \& BPCL}};
% Text Node
\draw (359,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize CA}};
% Text Node
\draw (405,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR}};
% Text Node
\draw (452,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in the present study for TİD
    \end{minipage}

\vspace{1 cm}

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node% Text Node
\draw (333,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR \& CA}};
% Text Node
\draw (445,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in Frederiksen and Mayberry (2016) for ASL
    \end{minipage}
\end{figure}

Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals
and ExtCLs as markers of lowly accessibly referents. Likewise, in both
hierarchies, as markers of highly accessible referents, different types
of null markers are further distinguished. Plain verbs rank the lowest
on the scales, preceded by agreement verbs and CA. On a similar ground,
both scales lack pronouns, which occurred very infrequently in the data.
Although \citet{frederiksen2016} did not include other classifier types
in the hierarchy, we proposed that WECLs and BPCLs must ranker higher
than null markers but lower than ExtCLs that were mainly used by
interlocutors to introduce a referent into the discourse context.

\hypertarget{age-of-acquisition-effects-on-referent-tracking}{%
\subsection{Age of acquisition effects on referent
tracking}\label{age-of-acquisition-effects-on-referent-tracking}}

The results in this study only somewhat support our second set of
hypotheses that we set for the effects of age of acquisition on
participants' RE selection. That is, while being a native acquirer of
TİD increased the odds of using null marking in reference tracking
across different discourse statuses, we failed to find such a difference
for overall nominal production. Taking a closer look at the interaction
terms, we observed that native signers, compared to late signers, were
less likely to use nominal constructions when the discourse status was
maintained. Native acquisition further increased the odds of null
markers when the discourse status was introduced. However, we found the
opposite alignment for the interaction of maintenance and native
acquisition. That is, the coefficient for this interaction term slightly
decreased the combined main effects of having a maintained context and
being a native signer. This means that these main effects were greater
than the interaction or added effects, pointing out to the implication
that the type of discourse status (introduced, maintained,
re-introduced) accounts for the distribution of REs in TİD narratives
better than age of acquisition (native, late) which, in comparison, had
a less salient effect.

While late signers were only partially over-redundant in maintained
contexts, they were more so in introduced contexts. In line with our
study, a tendency for limited over-redundancy has been suggested for
late signing children in DGS \citep{becker2009} and late signing adults
in BSL specifically for the use of CA\citep{cormier2013}. As for TİD, to
our knowledge, there is only one study \citep{gur2018} that directly
aimed at examining age of acquisition effects on adult signers'
narrative skills. \citet{gur2018} observed that late signing deaf adults
and children set the scene less frequently and used CA less accurately
compared to their native counterparts. However, this particular
investigation did not present any data on the distribution of REs
between two acquisition groups. In the present study, the finding that
native signers preferred to use null arguments even for the first
mention of a referent (19 in total) is surprising given some of the
previous literature in the topicality and saliency framework
\citep{cormier2013, swabey2002}. These research reported very little or
no use of null marking by native signers for introduced contexts.

Given the paucity of research directed at the late signing group, we now
turn to the referent tracking strategies observed for second language
learners of a sign language. Our study replicated some of the findings
of the line of research that investigated deaf or hearing bimodal
bilingual signers
\citep{bel2015, frederiksen2019}.\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).}
The authors reported that more instances of null marking in the context
of referent introduction was attested for native signers than deaf or
hearing bimodal bilingual signers. Although using null marking to
introduce a referent into the discourse would normally result in
referential ambiguity, it is important to note that our stimuli
consisted of 10 interconnected narratives with 2 or more recurring
characters. It may be possible that native signers, having already
introduced the referent with an explicit nominal marker in a different
discourse context, found it felicitous to refer to the same entity with
more implicit markers in the following narratives even for introduced
contexts. Then, our results imply a difference in pragmatic sensitivity
to referential accessibility between native and late signers, which is
more robustly reflected in introduced contexts.

Since we only observed a small difference between the two acquisition
groups for the use of certain REs, the aforementioned findings allude to
the implication that the effect on language deprivation on pragmatic
factors should also be considered in different contexts. For one, both
groups of signers can be tested with different genres of narratives and
with a more complex story design. In order to capture possible variation
in the accessibility levels of re-introduced referents, a more
fine-tuned protocol \citep[as in][]{toole1996} can be used for measuring
referent distance and the number of other possible competitors. Taken
altogether, our results accentuates the necessity of engaging in signed
narrative production early in life as well as the importance of early
intervention for late signing deaf children.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The present study aimed to investigate the reference tracking strategies
of native and late deaf adult signers in TİD narratives. In line with
our expectations, we found that the distribution of REs (nominal,
pronominal, classifier, and null marking) followed from previous
accounts of saliency and referential accessibility. For introduced
contexts, signers produced more nominals and ExtCLs, and to maintain a
previous referent, hey mainly used null arguments (CA, and plain and
agreement verbs) and certain classifiers (e.g., BPCL and WECL) to track
highly accessible referents. We observed only very few instances of
pronouns (mostly produced by late signers), unlike our expectations. The
ones we observed were mostly reserved for re-introduced contexts along
with nominals. Based on our findings across the two groups, we proposed
a hierarchy of referent accessibility for TİD. As for the age of
acquisition effects, we observed limited over-redundancy for signers
with delayed language acquisition (i.e., late signers) who employed
fewer implicit forms of reference tracking compared to native signers.
Native signers were less likely to use a nominal for maintained contexts
and more likely to use null marking for introduced contexts than late
signers. We conclude that native signers might only partially adhere to
pragmatic norms and leverage accessibility better than late signers
since they use less form when the referent is more active in the mind.

\clearpage

\appendix

\hypertarget{participant-background-information}{%
\section{Participant background
information}\label{participant-background-information}}

\label{section:a}

\begin{table}[hbt!]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-2}Participant demographic information based on self-report}

\small{

\begin{tabular}{llllllll}
\toprule
No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{Deaf schools} & \multicolumn{1}{c}{Home language}\\
\midrule
2 & Native & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign\\
3 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign\\
13 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign\\
14 & Native & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign\\
18 & Native & M & 0-3 & 28 & 12 & PS, MS & TİD\\
20 & Native & M & 0-3 & 32 & 12 & PS, MS & TİD\\
22 & Native & M & 0-3 & 35 & 12 & PS, MS, HS & TİD\\
24 & Native & F & 0-3 & 23 & 12 & PS, MS, HS & TİD\\
26 & Native & F & 0-3 & 30 & 12 & PS, MS, HS & TİD\\
27 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
29 & Native & F & 0-3 & 24 & 14 & PS, MS, HS & TİD\\
30 & Native & M & 0-3 & 18 & 12 & PS, MS & TİD\\
33 & Native & F & 0-3 & 21 & 16 & PS, MS, HS & TİD\\
34 & Native & M & 0-3 & 19 & 16 & PS, MS, HS & TİD\\
38 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
10 & Late & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR\\
23 & Late & M & 4-7 & 36 & 12 & MS, HS & TR, home sign\\
4 & Late & M & 4-7 & 33 & 12 & PS, HS & TR\\
16 & Late & F & 4-7 & 33 & 12 & PS & TR, home sign\\
21 & Late & F & 4-7 & 31 & 14 & PS, MS & Home sign\\
25 & Late & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign\\
35 & Late & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR\\
36 & Late & M & 8-12 & 43 & 16 & PS, MS & TR\\
12 & Late & M & 8-12 & 28 & 16 & MS & TR\\
15 & Late & F & 8-12 & 24 & 12 & PS, MS, HS & TR, TİD\\
19 & Late & F & 8-12 & 24 & 8 & PS, MS & Home sign\\
28 & Late & M & 8-12 & 33 & 12 & PS, MS & Home sign\\
31 & Late & M & 13-17 & 50 & 8 & MS & TR, home sign\\
32 & Late & M & 13-17 & 31 & 16 & PS, MS, HS & TR\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\clearpage

\setlength{\columnsep}{10pt}
\twocolumn
\linespread{1}
\justifying
\footnotesize

\hypertarget{episodes-in-the-stimuli}{%
\section{Episodes in the stimuli}\label{episodes-in-the-stimuli}}

\label{section:b}

\textbf{Narration 1}

1. Tom chases Jerry, wearing a wooden plank

2. Tom and Jerry enter a tube

3. Tom falls down and Jerry runs away

4. Tube gets thinner and Tom chases Jerry

5. Jerry gets out and escapes

6. Tom follows Jerry in shape of a stick

\textbf{Narration 2}

1. Tom walks past a female cat sitting on a couch

2. Tom turns to and approaches female cat

3. Female cat blushes

4. Tom offers a fish in aquarium to female cat

5. Female cat rejects offer

6. Tom offers a bird in cage to female cat

7. Episode 5 is repeated

8. Tom leaves by walking in a flirtatious manner

\textbf{Narration 3}

1. Jerry falls down and tries to walk up stairs

2. Tom pulls carpet back

3. Jerry struggles

4. A piano falls down stairs

5. Jerry manages to escape

6. Tom sees piano and tries to run away

7. Piano smashes into Tom and flattens him

8. Tom falls on ground

\textbf{Narration 4}

1. Tom hides behind a cardboard house

2. Jerry walks to house with a toy mouse

3. Toy leaves Jerry who stops at entrance

4. Toy enters house

5. Toy gets smashed into pieces

6. Jerry is surprised

7. Tom hurts his stomach after swallowing toy

8. Tom looks at his broken teeth on a mirror

9. Tom gets angry and throws down mirror

\textbf{Narration 5}

1. Tom fakes sleeping and suddenly gets up

2. Tom pulls a rope inside a couch

3. Jerry appears holding a bar of soap

4. Jerry washes Tom's mouth with bar and flees

5. Tom spits soap out

6. Jerry escapes to nest with rope

7. Tom catches rope and stars pulling it

8. A mouse trap appears and catches Tom's nose

9. Tom yells and removes trap

\textbf{Narration 6}

1. Tom points a sword at Jerry and catches him

2. Tom looks at Jerry

3. Red spots appear on Jerry's face

4. Tom gets scared and drops Jerry

5. Tom rushes into bathroom

6. Tom washes his hands

7. Tom opens a drawer

8. Tom swallows pills one after another

9. Feeling ill, Tom looks out of window

10. Tom and Jerry look at each other

\textbf{Narration 7}

1. Two male cats chase one another

2. Tom sees a female cat and kisses her

3. Tom hides away when black cat returns

4. Tom pokes his head behind couch again

5. Episode 3 is repeated

6. black cat sees female cat and hugs her

7. Tom appears and black cat kisses Tom

8. Two male cats move away from one another

\textbf{Narration 8}

1. Jerry hits Tom's knee with a hammer

2. Tom suddenly jumps up

3. Tom yells and holds his knee

4. Jerry appraches Tom holding a thermometer

5. Jerry sticks thermometer into Tom's mouth

6. Jerry lights a fire under thermometer

7. Red liquid starts to rise up

8. Thermometer explodes

\textbf{Narration 9}

1. Tom sits on Jerry who tries to get loose

2. Jerry looks around and sees a hat ribbon

3. Jerry finds a needle attached to hat

4. Jerry takes needle and sticks Tom with it

5. Tom turns red and jumps

6. Jerry placesn eedle between two cushions

7. Tom dodges needle

\textbf{Narration 10}

1. Wearing a stethoscope, Tom catches Jerry

2. Jerry yells into stethoscope

3. Tom's head pops up

4. Jerry runs and enter a mouse nest

5. Tom inserts a gun into nest

6. Gun bends over toward Tom

7. Tom fires gun and his head explodes

\clearpage

\onecolumn

\hypertarget{effect-of-narrating-to-the-camera-vs.-experimenter}{%
\section{Effect of narrating to the camera
vs.~experimenter}\label{effect-of-narrating-to-the-camera-vs.-experimenter}}

\label{section:c}

\begin{verbatim}
## Warning: Ignoring unknown aesthetics: height
\end{verbatim}

\begin{figure}
\centering
\includegraphics{manuscript_v2_files/figure-latex/model-camera-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Null) with the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late)}
\end{figure}

\clearpage

\hypertarget{additional-model-with-3-levels-for-nativeness}{%
\section{Additional model with 3 levels for
Nativeness}\label{additional-model-with-3-levels-for-nativeness}}

\label{section:d}

\begin{table}[hbt!]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-4}Model results for Nativeness with 3 levels}

\small{

\begin{tabular}{llllll}
\toprule
Coefficients & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{EstError} & \multicolumn{1}{c}{lower95} & \multicolumn{1}{c}{upper95} & \multicolumn{1}{c}{Rhat}\\
\midrule
1 & 1.23 & 0.20 & 0.83 & 1.62 & 1.00\\
2 & -4.25 & 3.56 & -13.27 & 0.09 & 1.01\\
3 & 0.49 & 0.54 & -0.55 & 1.56 & 1.01\\
4 & -0.41 & 0.49 & -1.37 & 0.57 & 1.00\\
5 & 3.76 & 0.25 & 3.29 & 4.25 & 1.00\\
6 & -5.24 & 0.21 & -5.66 & -4.82 & 1.00\\
7 & 1.01 & 0.63 & -0.16 & 2.30 & 1.00\\
8 & 0.02 & 0.64 & -1.24 & 1.24 & 1.00\\
9 & -0.33 & 0.55 & -1.43 & 0.73 & 1.00\\
10 & -0.92 & 0.52 & -1.96 & 0.11 & 1.00\\
11 & -13.40 & 10.69 & -40.57 & -0.41 & 1.01\\
12 & 14.22 & 10.69 & 1.29 & 41.29 & 1.01\\
13 & -21.98 & 14.25 & -58.10 & -4.70 & 1.01\\
14 & 11.27 & 7.13 & 2.67 & 29.30 & 1.01\\
15 & -54.37 & 42.74 & -162.14 & -2.68 & 1.01\\
16 & 55.46 & 42.75 & 3.99 & 164.16 & 1.01\\
17 & 26.68 & 21.38 & 0.88 & 80.69 & 1.01\\
18 & -27.59 & 21.38 & -81.73 & -1.81 & 1.01\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Rhat values above 1 indicate that the regression model chains did not converge properly. There were 4 chains in our regression model.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\clearpage

\renewcommand\refname{References}
\bibliography{./bibliography/references.bib}

\end{document}