### 2023-05-02 revised_manuscript.tex
model results added
Added the model results to the appendix.

\documentclass[review]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{Journal of Pragmatics} % Sets Journal name

\usepackage{lineno} % add

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\usepackage{graphicx}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{floatrow}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{float}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\floatsetup[figure]{capposition=top}
\floatsetup[table]{capposition=top}
\usepackage{times}
\usepackage{smartdiagram}
\usepackage{ragged2e}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{langsci-gb4e}
\usepackage{caption}
\usepackage{multicol}
\setlength{\columnsep}{-2cm}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\begin{frontmatter}

  \title{Reference Tracking Strategies of Deaf Adult Signers in Turkish
Sign Language}
      \cortext[cor1]{Corresponding author}

  \begin{abstract}
  We investigated the reference tracking strategies among deaf adults
  with signing deaf parents (DoD) and adult deaf signers with
  non-signing hearing parents or caregivers (DoH) who had late exposure
  to Turkish Sign Language (TİD). Consistent with the theories of
  saliency and referential accessibility, regardless of their acquisiton
  groups or parental hearing status, signers mainly used nominals and
  extension classifiers for introducing referents. To maintain a
  referent from the previous clause, the signers used zero anaphora
  (e.g., constructed action, agreement and plain verbs). For topic
  shifts or re-introduced contexts, nominals and pronouns were chiefly
  favored although we observed very little pronominal use. As for the
  effect of native acquisition from deaf caregivers, we only report
  limited over-redundancy for DoH signers who used zero anaphora less
  compared to DoD especially for introduced and maintained contexts. We
  conclude that DoH signers are still able to achieve native-like
  competency in terms of reference tracking in simple narratives but DoD
  signers utilize the spatial affordances of the visual-spatial modality
  better than DoH signers only to a certain extent.
  \end{abstract}
    \begin{keyword}
    Referential accessibility, signed narratives, pragmatic competence,
delayed language acquisition, Turkish Sign Language \sep 
    Referential accessibility, signed narratives, pragmatic competence,
delayed language acquisition, Turkish Sign Language
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Narratives are extensive discourses that form an important part of human
communication. In order to create a systematic and coherent organization
of experiences, narrators must choose appropriate linguistic forms to
refer to entities \citep{ariel1990, givon1983, gundel1993}. Certain
forms such as nominal expressions are commonly used for the first
mention of a referent in a discourse, and other forms such as pronouns
and zero anaphora may track previously mentioned entities which could
either be maintained across multiple clauses or re-introduced back into
the discourse. The selection of these referring expressions (henceforth
REs) is contingent on cognitive abilities, such that individuals must
rely on the activation of the working memory to keep in mind previously
referred expressions and long strings of information
\citep{morgan2005, bamberg1997}. Pragmatic abilities too play a pivotal
role in the tracking of referents. The form with the least amount of
redundant information must be selected based on the accessibility levels
of the referent (i.e., how active a referent is in the addressee's mind
given a point in discourse) \citep{ahn2019, ariel1990}. Because of these
cognitive and pragmatic necessities, it is not surprising to observe
that producing a coherent narrative is one of the latest milestones in
language acquisition among children \citep{hickmann1996}. As can be
expected, there is a plethora of research examining how narrative
competence is achieved in both monolingual and bilingual acquisition
contexts among children and adults. Although narrative competence and
referent selection have been studied extensively for typically
developing speakers \citep{aksu-koc2015, hickmann1996, williams1988},
deaf individuals who acquired a sign language in atypical learning
contexts (i.e., deaf signers with delayed exposure to a sign language)
have remained mostly understudied with an exception of a handful of
studies \citep{cormier2013, gur2018, becker2009}. Nevertheless, it is
paramount that we investigate and describe the effects of late exposure
to a sign language among deaf adults growing up in a signing
vs.~non-signing environment on narrative competence more in depth. This
will help us understand whether the lack of systematic and frequent
language exposure in the early years of life affect how deaf signers
produce a narrative in their adulthood. Given this, the present study
will focus on two groups of signers and examine their reference tracking
strategies in signed narratives: (i) deaf of deaf (DoD) adult signers
(i.e., signers who have deaf caregivers and natively acquired a sign
language from birth) and (ii) deaf of hearing (DoH) adult signers (i.e.,
signers who have hearing non-signing caregivers and state to have
started acquiring a sign language later in life).

Sign languages are the natural languages of the Deaf communities all
around the world. They are independent languages, which do not derive
from majority spoken language(s) but emerge naturally when deaf people
form a community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and
has been the officially recognized language of the Deaf community in
Turkey since 2005 \citep[see][ for a brief review of the historical
development of the language]{nuhbalaoglu2018}. Deaf or hearing children
acquiring a SL within a family where there is one or more deaf
caregivers display the same developmental milestones as hearing children
acquiring a spoken language in a hearing family
\citep{chamberlain_language_1999, meier1990}. However, reports indicate
that over 90\% of all deaf individuals are born into hearing families
(DoH) who do not know how to sign \citep{woll2013}. This means that the
auditory modality used in the household is not accessible to many deaf
children. That is why the age of acquisition of a first functional
language varies greatly among deaf children, with many of whom being
deprived of their first languages typically until after age 7 when they
start formal education in a deaf school, especially in the context of
Turkey \citep{sari_analysis_2005}. Despite this, many of the early
intervention programs for late signing children prioritize spoken
language development and discourage the use of a sign language
\citep{kemaloglu2012}. The \citet{ministryofnationaleducation2020}
identifies 72 deaf schools in Turkey for the K-12 level. Even when deaf
children with hearing parents start their formal education in one of
these schools, they do not receive proper instruction in TİD since the
teachers themselves are mainly hearing and only have a very rudimentary
knowledge of a sign language \citep{ilkbasaran2015, sari_analysis_2005}.
In some cases, it has been argued that late exposure to a sign language
among certain DoH signers has negative effects on some cognitive and
linguistic measures. To name a few, delayed exposure to a first
functional language is reported to negatively affect deaf children and
adults' morphosyntactic production and processing
\citep{kayaba2022, boudreault2006, newport1990, cheng2021, karadoller2017, cormier2012}
and narrative abilities \citep[e.g.,][]{becker2009}. In our overall
research, we are concerned with the effects of this late exposure in the
ultimate adult forms of TİD users.

\hypertarget{referring-expression-re-selection}{%
\subsection{Referring Expression (RE)
Selection}\label{referring-expression-re-selection}}

Natural languages can be systematically used to keep track of previous
referents in the discourse and mark the changing cognitive statuses of
the REs (i.e., from being active to less active in the addressee's mind)
to create a more coherent narrative. Among the theoretical frameworks
that aim to explain that different linguistic forms are selected in
accordance with the addressee's mental representations of the referents
in the discourse is the Givenness Hierarchy \citep{gundel1993}.
According to this model, given or familiar information is expressed in
more attenuated forms such as pronouns and zero anaphora (i.e., there is
no explicit mention of the referent) since the mental status of the
expressed referent is already at the center of attention, as indicated
by the present contextual information \citep{chafe1976}. The Givenness
Hierarchy is akin to other previously proposed models of reference
control within the topicality and saliency paradigm such as the
Accessibility Hierarchy \citep{ariel1990}, Topic Continuity
\citep{givon1983}, and Familiarity Scale \citep{prince1981}.
\citet{gundel1993} argued that the distribution of forms is not random
but abides by the Gricean Maxim of Quantity \citep{grice1975}. That is
to say, individuals must make the most informative and succinct choice
possible to retain economy of form. When multiple expressions are
available for the same referent as in an utterance like (\ref{ex:1}),
phonetically simplest form is selected with the least redundant
information based on an economy principle. In the utterance, REs for
``Eddie'' are given in bold. All three REs (nominal, pronominal, and
zero anaphora) refer to the same entity; however, the quantity of
marking decreases as the referent's accessibility increases. On this
ground, for highly accessible referents (i.e., those that are highly
active in the mind), linguistic forms with the least phonetic content
are selected for the least amount of processing effort.

\begin{exe}
    \ex \label{ex:1}
    So \textbf{Eddie} turned around. \hfill Williams (1988; p. 343, ex. 7) \\
\textbf{He} said, "youse got a problem?" \\
"Yeah, we want you," they say.
So-$\pmb{\emptyset}$ walked right up to them \\
\end{exe}

It is also possible to categorize REs into three according to their
discourse status \citep{gullberg2006}. A referent can be introduced or
mentioned for the first time, maintained, that is, continued across two
or more clauses, or re-introduced into the discourse following a topic
shift. As a result, interlocutors mark some REs differently than others
based on the principles of reference accessibility. For spoken
languages, this observation is well established
\citep{ariel1990, givon1983, gundel1993, debreslioska2013, ahrenholz2005, chafe1976, hickmann1996, azar2015}.
These research indicate that the introduction and re-introduction of a
referent are often accomplished by full noun phrases which can be
commonly modified by an adjective or demonstrative. Maintenance, on the
other hand, is typically marked with pronominals and zero anaphora since
the continued referent is still highly accessible to the addressee.

However, this particular observation is not peculiar to spoken
languages. There is an emerging body of literature that reports the
distribution of REs in signed narratives
\citep{czubek2017, ferrara2022, swabey2002, frederiksen2016, perniss2015, morgan2005, hodge2019, ahlgren1994, garcia_reference_2014, garcia2018impersonal}.
Building on the work by \citet{gundel1993}, \citet{swabey2002}
demonstrated that, like other languages such as Russian, Chinese, and
Japanese, American Sign Language (ASL) signers use bare nouns for
introduced and re-introduced discourse contexts whereas zero anaphora is
used for referents which are ``in-focus'' or very accessible to the
addressee. Overt REs like pronouns or nouns are disprefered for such
referents. \citet{swabey2002} concluded that ``native'' signers in her
study abided by the Gricean maxim of quantity and thus used more
economic forms of reference tracking. Similar findings have been
reported for German Sign Language (Deutsche Gebärdensprache - DGS)
\citep{perniss2015} and British Sign Language (BSL) \citep{morgan2005}.
It is important to note, however, that this research either collapsed
DoD and DoH signers into a single group and labeled them as
``native-like'' or explicitly compared native deaf signers with late
deaf signers (with delayed exposure to a sign language) in terms of
narrative competence. Most recently, \citet{ferrara2022} examined
corpora including video-recorded narratives from five different sign
languages including Irish Sign Language, Finnish Sign Language,
Norwegian Sign Language and Swedish Sign Language and investigated how
these languages established and maintained referents. Like other
observations, this research showed that, across each of these languages,
bare nouns were used to introduce new referents and zero anaphora or
more reduced forms were preferred for referent maintenance.

\hypertarget{linguistic-tools-for-tracking-referents-in-signed-narratives}{%
\subsection{Linguistic Tools for Tracking Referents in Signed
Narratives}\label{linguistic-tools-for-tracking-referents-in-signed-narratives}}

The linguistic inventory for discourse referents can include nominal
constructions (bare nouns, fingerspelled nouns, and modified nouns),
null or zero
anaphors\footnote{Zero anaphora does not refer to the lack of a referent. It only alludes to the non-presence of a more explicit anaphoric form for the referent such as a nominal, pronominal, or classifiers.}
(predicated by a plain verb, an agreement verb, or constructed action),
classifier predicates (whole entity classifiers, body part classifiers,
handling classifiers, and extension classifiers), and overt pronouns
(pronominal index or IX) \citep{frederiksen2016}. This interim section
describes such forms attested for sign languages under four main
headings (nominals, pronouns, classifiers, and zero anaphora).

\hypertarget{nominals}{%
\subsubsection{Nominals}\label{nominals}}

In TİD and many other sign languages, it is possible to use bare nouns
as in \ref{ex:2a} as well as fingerspelled nouns as in \ref{ex:2b}.
Fingerspelling refers to the use of the manual alphabet to spell all or
part of the letters of a word borrowed from a surrounding written
language \citep{kubus2008}. Nouns can also be modified by a prenominal
or postnominal extended index finger pointing sign, glossed as IX,
functioning as a determiner or a demonstrative as in \ref{ex:2c}
\citep{maclaughlin1997} or by a modifying classifier within the same NP
as in \ref{ex:2d}
\citep{frederiksen2016}.\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990). In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.}

\newpage
\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:2a} Full noun \\\glll
{} \includegraphics[width=35pt]{pictures/p1.png} {} \\ \textbf{Gloss:} \textsc{mouse\footnotemark} \textsc{sick} \\ 
\textbf{Status:} Intro. \\
\glt \textbf{Context:} Narration 6, Event 1. Here, the signer introduces the mouse, Jerry, at the beginning of the narration with a bare noun.
\glt \textbf{Translation:} There is a sick mouse.
\vspace{0.35cm}

\ex \label{ex:2b} Fingerspelled noun \\\glll
{} \includegraphics[width=40pt]{pictures/p2.1.png} \includegraphics[width=40pt]{pictures/p2.2.png} \includegraphics[width=40pt]{pictures/p2.3.png} {} \\ \textbf{Gloss:} \textsc{t-} \textsc{-o-} \textsc{-m} \textsc{come.back} \\ 
\textbf{Status:} Re-intro. \\
\glt \textbf{Context:} Narration 7, Event 2. The signer introduces Tom two sentences before. She introduces the black cat in the preceding clause and then here re-introduces Tom by using fingerspelling.
\glt \textbf{Translation:} Tom comes back.
\vspace{0.35cm}

\ex \label{ex:2c} Noun modified by index (IX) \\\glll
{} \includegraphics[width=45pt]{pictures/p3.1.png} \includegraphics[width=40pt]{pictures/p3.2.png} {} {} {} \\ \textbf{Gloss:} \textsc{ix} \textsc{cat} \textsc{pipe.extend} \textsc{go.through} \textsc{exit} \\ 
\textbf{Status:} Re-intro. \\
\glt \textbf{Context:} Narration 1, Event 6. The signer introduces the cat in the first clause of the first event and re-introduces it several times including this last event of the narrative.
\glt \textbf{Translation:} The cat goes through and exits the thin pipe.
\vspace{0.35cm}

\ex \label{ex:2d} Noun modified by classifier (CL) \\\glll
{} \includegraphics[width=35pt]{pictures/p4.1.png} \includegraphics[width=35pt]{pictures/p4.2.png} \\ \textbf{Gloss:} \textsc{soap} \textsc{soap.cl} \\ 
\textbf{Status:} Intro. \\
\glt \textbf{Context:} Narration 5, Event 3. Here the signer introduces the soap to the discourse. The signer produces the lexical sign, \textsc{soap}, plus a classifier, \textsc{soap.cl}.
\glt \textbf{Translation:} The soap, which is like this.
\end{xlist}
\end{exe}

\footnotetext{We use small capitals to indicate sign glosses.}

\hypertarget{referential-use-of-space-and-pronouns}{%
\subsubsection{Referential Use of Space and
Pronouns}\label{referential-use-of-space-and-pronouns}}

In addition to being a modifier of a noun, the IX sign can also be
pronominal by referring to established nominals through pointing to a
previously assigned abstract or physical locus in the signing space when
re-introducing a topic in discourse \citep{emmorey1996}. This function
has also been identified for TİD \citep{sevinc2006}. The example in
\ref{ex:3} illustrates this spatially-anchored pronominal use.

\ea 
\ea \label{ex:3} Pronominal \\\glll
{} \includegraphics[width=35pt]{pictures/p5.png} {} {} {} {} \\ \textbf{Gloss:} \textsc{ix} \textsc{look} \textsc{cat.sad(constructed action)} \textsc{sad} \textsc{sick} \\ 
\textbf{Status:} Re-intro. \\
\glt \textbf{Context:} Narration 9, Event 1: The signer introduces the mouse then signs an action of the cat. Here, he re-introduces the mouse with a pointing sign that functions as a pronoun.
\glt \textbf{Translation:} It (the mouse) looks sad and sick.
\z \z

These pronominal signs are rarely used for the first mention of an
entity \citep{perniss2015}. Other researchers consider pointing signs
without noun phrases as examples of demonstratives and not personal
pronouns \citep{koulidobrova2016}. It is also important to mention here
that in some spoken languages (e.g., German and Dutch), demonstratives
can also refer to persons \citep{bosch2007} and behave similarly to
personal pronouns. However, for the purposes of the present study, we
will refer to IX signs as pronominal IX and particularly examine their
non-deictic or personal pronoun realizations as a possible discourse
strategy to mainly mark topic shift (re-introduction) among DoD and DoH
signers of TİD.

\hypertarget{classifiers}{%
\subsubsection{Classifiers}\label{classifiers}}

In sign languages, a verbal classifier morpheme is morpho-phonologically
encoded by a handshape. This handshape is articulated co-temporally with
a movement that encodes the verbal meaning
\citep{supalla1990, benedicto2004}. Semantically, the classifiers denote
sets of objects based on criteria such as animacy, size, shape,
body-part, texture and sometimes function \citep{talmy2003}. The
classifier articulated simultaneously with a movement is called a
classifier predicate and the sentence in which such a predicate is used
is usually called a classifier construction in sign language research.
As in most sign languages hitherto studied, classifier constructions are
also commonly found in TİD \citep[see][ for a review]{dikyuva2015}.
Following the classification of \citep{engberg-pedersen1993}, in TİD, we
distinguish between four types of classifier handshapes. Whole entity
classifiers (WECL) represent the entirety of a referent (e.g.~an entire
flat horizontal entity, or an upright entity) (e.g., \ref{ex:4a})
whereas body part classifiers (BPCL) represent a specific part of an
animate entity (e.g., limbs, head, mouth, tongue, or teeth, etc.)
(e.g.,\ref{ex:4b}). Handling classifiers (HCL) occur with transitive
verbal roots and represent both the size, shape and surface of an object
and how that object is handled or manipulated by an agent (e.g.,
\ref{ex:4c}). Extension classifiers (ExtCL), on the other hand,
represent the size and shape of an entity by following a trajectory to
trace its shape (e.g., \ref{ex:4d}). We exemplify these handshapes with
the discourse status of the referent they track and a description of the
previous discourse when available.

\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:4a} Whole entity classifier  (WECL) \\\glll
{} \includegraphics[width=40pt]{pictures/p6.png} \\ \textbf{Gloss:} \textsc{two.mice.go.together} \\ 
\textbf{Status:} Maintain. \\
\glt \textbf{Context:} Narration 4, Event 2: The signer introduces Jerry and the robot mouse in the preceding clause. She maintains reference to them by using a whole entity classifier on each hand. 
\glt \textbf{Translation:} (Jerry and the robot mouse from the previous context) go together.
\vspace{0.35cm}

\ex \label{ex:4b} Body part classifier (BPCL) \\\glll
{} \includegraphics[width=40pt]{pictures/p7.png} \\ \textbf{Gloss:} \textsc{cat.move.thin.legs} \\ 
\textbf{Status:} Maintain. \\
\glt \textbf{Context:} Narration 1, Event 6: The signer re-introduces the cat three clauses before and maintains reference to him in the two clauses before and here with a body part classifier that refers to the legs of the cat.
\glt \textbf{Translation:} (The cat) walks (i.e. moving his thin legs forward).
\vspace{0.35cm}

\ex \label{ex:4c} Handling classifier (HCL) \\\glll
{} {} {} {} {} \includegraphics[width=40pt]{pictures/p8.png} \\ \textbf{Gloss:} {} \textsc{rope see-3a} \textsc{soon} \textsc{go} \textsc{cat.pull.rope} \\ 
\textbf{Status:} Maintain. \\
\glt \textbf{Context:} Narration 5, Event 2: The signer introduces Tom three clauses before. She maintains reference to him in the preceding two clauses and here she still maintains reference to him by using a handling classifier on both hands.
\glt \textbf{Translation:} When (he/Tom) sees the rope, (he) soon goes to pull it.
\vspace{0.35cm}

\ex \label{ex:4d} Extension classifier (ExtCL) \\\glll
{} \includegraphics[width=80pt]{pictures/p9.png} \\ \textbf{Gloss:} \textsc{cat.extend} \\ 
\textbf{Status:} Maintain. \\
\glt \textbf{Context:} Narration 1, Event 6: The signer re-introduces Tom two clauses before and she maintains reference to him with constructed action in the previous clause and here with an extension classifier.
\glt \textbf{Translation:} (The cat) extends thin.
\end{xlist}
\end{exe}

\hypertarget{zero-anaphora}{%
\subsubsection{Zero Anaphora}\label{zero-anaphora}}

Parallel to spoken languages, many sign languages have been reported to
allow null subjects and/or objects and make use of null arguments when
referring to highly accessible referents that are still active in the
mental representation of the addressee. In signed discourse, this can be
achieved by employing certain verbal constructions or constructed action
(CA). The latter construction exemplified in \ref{ex:5a} refers to cases
in which the signers engage with the space in front of themselves as if
they are characters in the story interacting with objects and other
characters. This is often achieved with the signer taking on the role of
a referent in the narrative and performing the actions
\citep{metzger1995}. CA is often marked by certain non-manual cues such
as the facial expression and the breaking of the eye contact. This is to
say, the narrator's head, face, torso, arms, and as an optional marker,
hands represent those of the associated referent(s) \citep{smith2014}.
As for the verbal constructions, to identify the predicate types in ASL,
\citet{padden1986} suggested a triadic categorization of verbs in sign
languages: plain verbs, agreement verbs, and spatial verbs. According to
this classification, plain verbs (as in \ref{ex:5b}) do not agree with
the subject or the object. This means that they are not inflected for
number or person. This contrasts with agreeing verbs (as in \ref{ex:5c})
the starting or the ending point of which can be modified to specific
loci for person and number agreement while spatial verbs (as in
\textsc{ix book put} `I placed the book there') agree with locative
arguments but like plain verbs, they are not inflected for number or
person. TİD too licenses null arguments. Following this classification,
\citet{sevinc2006} argued that although it is common in TİD to drop
arguments, there were certain limitations on the null marking of the
arguments for agreement and plain verbs.

In terms of referential accessibility, zero anaphors predicated by such
constructions in narratives are assumed to rank last in the hierarchy,
preceded by nominal classifiers and then pronouns and other nominal
forms. The reason that we distinguish zero anaphora from classifer
handshapes or predicates (which can also be considered implicit markers
of reference) is because classifiers typically denote certain
topographic or semantic properties (e.g., size and shape) of the
referent noun due to their iconic nature. They are not as explicit as
pronouns or nouns since they do not target a specific individual (e.g.,
\textsc{eddie}) in a set but rather they refer to a set of individuals
that are semantically similar to each other \citep{morgan2005}. For
example, the extended index finger can be used to refer to a human
entity. Finally, zero anaphors predicated by constructed action,
agreement and plain verbs are least explicit forms of referent tracking
since the implicit nature of such constructions do not easily restrict
the referent set based on cues such as size, shape or semantic category.
In other words, zero anaphora requires preceding context or information
the most compared to other forms.

\begin{exe}
\ex
\begin{xlist}
\ex \label{ex:5a} Constructed action \\\glll
{} \includegraphics[width=40pt]{pictures/p10.png} \\ \textbf{Gloss:} \textsc{mouse.stay.like.this} \\ 
\textbf{Status:} Maintain. \\
\glt \textbf{Context:} Narration 6, Event 9: The signer re-introduces Jerry in the previous clause and maintains reference to him by using this constructed action.
\glt \textbf{Translation:} (The mouse) stays like this.
\vspace{0.35cm}

\ex \label{ex:5b} Plain verb \\\glll
{} \includegraphics[width=80pt]{pictures/p11.png} \\ \textbf{Gloss:} \textsc{run.away} \\ 
\textbf{Status:} Re-introduce. \\
\glt \textbf{Context:} Narration 5, Event 4: In this sentence, the signer re-introduces Jerry with the plain verb \textsc{run.away}.
\glt \textbf{Translation:} (Jerry) runs away.
\vspace{0.35cm}

\ex \label{ex:5c} Agreement verb \\\glll
{} \includegraphics[width=80pt]{pictures/p12.png} \\ \textbf{Gloss:} \textsc{3a(mouse)-hit-3b(cat)} \\ 
\textbf{Status:} Maintain. \\
\glt \textbf{Context:} Narration 8, Event 2: The signer introduces Jerry in the preceding clause and maintains reference to him by using the agreement verb \textsc{hit}.
\glt \textbf{Translation:} (Jerry) hits (Tom).
\vspace{0.35cm}

\end{xlist}
\end{exe}

\hypertarget{learning-how-to-track-referents-in-sign-languages}{%
\subsection{Learning How to Track Referents in Sign
Languages}\label{learning-how-to-track-referents-in-sign-languages}}

In both communication modalities (visual and auditory), children start
to produce more complex and coherent stories with appropriate and
informative linguistic forms to indicate anaphoric or cataphoric
references after age 10 \citep{hickmann1996, rathmann2007}. Prior to
this age, both signing and speaking children fail to refer to entities
consistently since their pragmatic abilities are yet to develop. The
manipulation of linguistic forms for discourse functions to signal topic
continuance or shift is difficult to master until the beginning of
formal education. These findings are also comparable across different
sign languages including BSL \citep{morgan2005}, ASL \citep{loew1984},
and DGS \citep{becker2009}, pointing out to the fact that the mastery of
reference tracking in narratives is one of the most protracted
achievements in the acquisition of signed discourse.

Like in first language development, narrative skills have been
extensively studied among second language learners
\citep[see][]{bel2015, frederiksen2019}. Comparing 11 deaf adult signers
and 13 hearing advanced learners of Catalan Sign Language (Llengua de
Signes Catalana - LSC), \citet{bel2015} found that hearing signers
resort to more overt forms (nominals and pronouns) than native deaf
signers to avoid potential ambiguities among referents. Nevertheless,
their results did not reach statistical significance for the interaction
of discourse status and acquisition group. However, the control group
here consisted of both deaf signers who had one or more deaf parents and
also deaf signers who had hearing parents and their age of acquisition
varied. Out of 11 deaf L1 signers in Bel et al.'s \citeyearpar{bel2015}
control sample, 5 had deaf parents and acquired LSC from birth, and the
other 6 had hearing parents but their exposure to LSC started before age
5. A similar line of research came from \citet{frederiksen2019} who
found a statistically significant difference between both groups, that
is DoD signers and hearing L2 learners of ASL, only for classifier use
especially in re-introduced contexts. L2 learners in the study used more
nominals and fewer instances of zero anaphora in number, indicative of a
limited effect of overexplicitness among second language learners.
Although Frederiksen and Mayberry's control group included some signers
with hearing parents, they had all learned to sign from birth since they
received sign language input from older deaf siblings.

Still, only very few researchers explicitly compared DoD (i.e., deaf
signers with native exposure to a sign language) and DoH (i.e., deaf
signers with delayed exposure to a sign language) also display
redundancy in their reference tracking. \citet{becker2009} examined 6
DGS signing deaf children's (3 native and 3 late) narrative abilities.
She reported that late signing children generated more explicit forms
such as nominals when referring to the characters in the story compared
to native learners, concluding that first language deprivation results
in an inadequate representation of the addressee and less efficient use
of morphosyntax. Comparably, \citet{cormier2013} investigated the use of
CA in BSL among 15 deaf native, early, and late adult signers (5
participants in each group) and found that DoH signers were redundant in
topic continuity. In other words, the DoH signers produced more explicit
forms like nominals, and fewer instances of zero anaphora when
maintaining previous referents. Most relevantly to the present study,
\citet{gur2018} studied the development of narrative skills among 10
native and 10 late deaf children and deaf adult signers of TİD. She
found that native-signing children and adults had more instances of
explicit referent introduction, set the scene more frequently, and
employed CA in an appropriate manner more than corresponding
late-signing children and adults. Although this study presented
preliminary findings on age of acquisition effects on narrative
competence, it does not investigate how DoD and DoH signers employ
different linguistic tools based on the accessibility of the referent.
Therefore, to our knowledge, there is a lack of research directly
examining the effects of first language delay especially among deaf
adult signers with non-signing hearing parents on the selection of REs
in signed narratives.

Considering this, it is also not very clear whether the age of
acquisition effects on narrative abilities, observed among late signing
deaf children, hold when they become adults. Understanding the impact of
delayed first language learning on narrative competence is crucial for
three reasons. First, studying signers from this population will shed
light on the possible effect of lack of early language experience on
future cognitive and pragmatic abilities since the narrative competence
is dependent on such faculties. Second, examining reference tracking
strategies in particular will also inform about the linguistic inventory
that DoD and DoH signers use to refer to entities. Finally, any possible
difference between the two acquisition groups in terms of narrative
competence will have further practical implications for the importance
of early access to a first functional sign language to retain future
linguistic and cognitive abilities.

\hypertarget{present-study}{%
\subsection{Present Study}\label{present-study}}

In an attempt at addressing this gap in the literature, the present
study aims to investigate the reference tracking strategies of DoD and
DoH adult signers in TİD narratives with a story-telling paradigm. We
explore the forms available for referent tracking in TİD and examine the
nature of RE selection as an indicator of possible pragmatic differences
between the acquisition groups.

Overall, considering what has been suggested for the distribution of REs
in signed narratives, we expect more explicit forms like nominals for
referent introduction and and less explicit forms such as zero anaphora,
pronominal and/or classifier use for referent maintainance. For the
re-introduction contexts, where referents which are discontinued across
one or more clauses are brought back into the discourse, we expect a
less clear picture. Some re-introduced referents will have higher
accessibility and will thus be tracked with more implicit REs, while
others that have lower accessibility will be tracked with more explicit
REs. Also, based on previous research, we hypothesize that DoH signers
will be more over-redundant in their reference tracking compared to DoD
signers. This would translate to a preference by DoH signers for
nominals over zero anaphora especially in maintenance and
re-introduction contexts. In contrast, we would expect that DoD signers
use more zero anaphora to track referents.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We tested 29 deaf adult signers, 15 DoD signers with exposure to TİD
from birth from deaf caregivers (e.g., parents or siblings), 14 DoH
signers who had hearing caregivers and were exposed to TİD between ages
3 and 17. (\emph{M}\textsubscript{Age of acquisition} = 7.7 years,
\emph{SD} = 3.3 years). As a result, the criterion to be classified as a
DoH signer was to have hearing and non-signing caregivers. All the
signers satisfying this criterion were regarded as
such.\footnote{Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3. In order to capture in-group variation among DoH signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative. For the analysis, see \ref{section:d}.}
Detailed information about the linguistic and educational background of
the participants can be found in \ref{section:a}.

All participants indicated that their preferred language of
communication was TİD which they used for most contexts. 1 DoD signer
was excluded from the data analysis because they self-reported to have
signing deaf parents but also only used home sign (a more crude form of
communication) at home. As a result, we present data from only 14 DoD
signers.

All participating DoD signers (7 males, 7 females) were born into deaf
signing families and started acquiring TİD from at least one of their
caregivers after birth. The parents of 14 participating DoD signers were
deaf. One native signer had a deaf father and a hearing mother. In
addition to TİD, the home language for some DoD signers (n = 4) in early
childhood included the use of home
sign.\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers. Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages.}.
The age of the native participants ranged between 18 and 35 (\emph{M} =
26.4 years, \emph{SD} = 5.0 years) This also corresponded to the total
average years of TİD use for DoD signers. All DoD signers were
pre-lingually deaf, that is they were either born deaf or they had
become deaf before the age of 3, with the exception of one signer who
became deaf between the ages of 4 and 7. DoD signers reported to
complete at least high school education and their level of education
(LOE) was between 12 and 16 years (\emph{M} = 12.7 years, \emph{SD} =
1.5 years). In addition, all native participants attended at least two
deaf schools, one being primary and the other middle school. 10 native
participants also attended a deaf school for their high school
education. Based on a self-reported 5-point scale of proficiency (1:
poor, 5: proficient), native participants declared that they have high
receptive and productive skills in TİD (\emph{M} = 5 points, \emph{SD} =
0 points). As for their proficiency in (spoken) Turkish, their
self-reported rankings were on average 3.7 (\emph{SD} = 0.7), 3.7
(\emph{SD} = 0.7), and 2.6 (\emph{SD} = 1.2) for reading, writing, and
speaking abilities, respectively.

Comparably, DoH signers (8 males, 6 females) were all born into hearing
non-signing families and started learning TİD later in life after age 3,
usually with the start of formal education. The age of sign language
acquisition for DoH signers ranged between years 3-17 (\emph{M} = 8.7
years, \emph{SD} = 3.5 years). The caregivers of all participating DoH
signers were hearing, and the reported home languages used in childhood
mainly consisted of Turkish and home sign. The average age of DoH
signers was 34.6 years (\emph{SD} = 8 years) and ranged between 24 and
50. The derived mean number of years for TİD use following its
acquisition was comparable to DoD signers with 25.9 years (\emph{SD} =
7.9 years). Except for two signers who became deaf between the ages of 4
and 7, all the other DoH signers were pre-lingually deaf. All DoH
signers completed at least middle school education and the year of
schooling ranged between 8 and 16 years, and the average was 12.3 years
(\emph{SD} = 2.8 years). All of the DoH group attended at least one deaf
school. While 8 participants attended two deaf schools (either primary
and middle school, or middle and high school), 3 participants reported
to attend three deaf schools. Using the same scale, the signers
self-rated their abilities in TİD and Turkish. Both of their receptive
and productive TİD abilities rated on average 4.9 points (\emph{SD} =
0.3 points). DoH signers' mean rating for their reading abilities in
Turkish was 3.5 points (\emph{SD} = 0.7 points) while their writing and
speaking abilities rated on average 3.4 (\emph{SD} = 0.5 points) and 3.1
points (\emph{SD} = 0.7 points), respectively.

Both groups' sex, average years of schooling, TİD use following its
acquisition, and their self-reported receptive and productive TİD
ratings, and their reading, writing, and speaking ratings for Turkish
did not statistically differ; Bayesian \emph{p}'s \textgreater{} 0.1.
However, participants' ages statistically differed between the two
acquisition groups; \emph{t}(27) = 3.4, Bayesian \emph{p} \textless{}
0.01. Overall, the DoD group consisted of younger participants. This was
to ensure that the average years of TİD use among DoD signers
approximated those of DoH signers.

\hypertarget{stimuli}{%
\subsection{Stimuli}\label{stimuli}}

We used a story-telling paradigm to examine how DoD and DoH signers of
TİD track reference in signed discourse. Participants were presented
with 10 short video clips extracted from the cartoon \emph{Tom and
Jerry} and asked to retell them. Each clip had six to ten episodes
(\emph{M} = 8 episodes). The duration of the clips ranged from 17
seconds to 28 seconds (\emph{M} = 22.7 seconds). The narratives included
on average 2 recurring animate characters for which the present context
required the introduction, maintenance, and re-introduction of
referents. In addition, in each narrative, there were on average 2
inanimate objects with which the animate referents interacted. The
detailed description of the episodes in each narrative can be found in
\ref{section:b}.

\begin{table}[p]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:tab-2}Annotation Tiers with their Tags and Descriptions}

\footnotesize{

\begin{tabular}{p{3.2cm}p{3.8cm}p{3.8cm}}
\toprule
Tier & \multicolumn{1}{c}{Tag} & \multicolumn{1}{c}{Description}\\
\midrule
Clause Boundary &  & The boundary between clauses indicated by the presence of predicates and/or certain prosodic cues (e.g., head nod)\\
Discourse Status & Introduced & A first mention of an entity in a given discourse regardless of any syntactic role\\
 & Maintained & A subject referent continued across two or more immediate clauses\\
 & Reintroduced & A subject referent discontinued across at least one clause\\
RE Type & NOM: Bare Noun & A non-modified noun phrase (e.g., CAT, MOUSE)\\
 & NOM: Modified Noun & A noun phrase modified by an index sign, classifier etc.\\
 & NOM: Fingerspelled Noun & Borrowed nouns from Turkish spelled by the manual alphabet (e.g., K-E-D-İ ‘cat’)\\
 & PRO: Pronominal IX & An index sign used in isolation to refer to an entity in discourse\\
 & CL: Whole Entity Classifier & A handshape that refers to the entirety of an entity\\
 & CL: Body Part Classifier & A handshape that refers to a part of an entity\\
 & CL: Handling Classifier & A handshape that refers to the handling of an item\\
 & CL: Extension Classifier & A handshape that denotes the size and shape of an entity\\
 & ZERO: Constructed Action & A multifunctional construction where the signers take on the role of a referent and perform their actions\\
 & ZERO: Plain Verb & A non-agreeing predicate (e.g., LOVE)\\
 & ZERO: Agreement Verb & A predicate the beginning and ending of which agree with the person and number\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} NOM = nominal, PRO = pronominal, CL = classifier, ZERO = zero anaphora.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\hypertarget{task-procedure}{%
\subsection{Task Procedure}\label{task-procedure}}

Participants sat in front of a laptop computer which had the stimulus
stories. The instructions were all given by a deaf research assistant
who was present in all sessions and ensured that the task requirements
were clear. The informed consent form was translated into TİD and
explained to the participants by the same assistant before data
recording. Participants filled in the consent form and a background
questionnaire form before the onset of the trials. Following this step,
signers were seated across a hearing researcher who is an advanced late
learner of TİD. We instructed the participants to narrate as they would
to a deaf friend. We gave this instruction to make sure that signers
naturally sign, and simply do not try to simplify or adjust their
language to the level of the addressee. All trials were recorded by a
digital camera placed in front of the signers. The participants were
instructed to watch the narrative clip as many times as they needed, and
narrate the content of each clip either to the experimenter or the
recording digital
camera.\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining narrated to the camera. We did not make it obligatory for participants to  narrate to the digital camera or to the experimenter for few reasons. First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter. Therefore, most signers preferred to narrate to the camera. Furhermore, some signers did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available. We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers (see \ref{section:c}). The results imply that there is not much of a difference.}

\hypertarget{data-annotation}{%
\subsection{Data Annotation}\label{data-annotation}}

We coded the discourse status (Introduction, Maintenance,
Re-introduction) and RE type (Nominal, Pronominal, Classifier, Zero
Anaphora). We used ELAN Linguistic Annotation Software
\citep{crasborn2008} for data annotation. First Author who is a hearing
signer of TİD annotated all the data and translated the narratives into
Turkish together with a native deaf signer of TİD.

Following \citet{berman2013relating}, we first identified the clause
boundaries in each sentence indicated by the presence of predicates and
informative non-manual markers (e.g., head nod). On the second tier, we
coded the discourse status for each referent. Having adapted the
classification of \citet{gullberg2006}, we considered all referents
mentioned for the first time as introduced regardless of their syntactic
function in the sentence. Only subjects were coded for maintenance and
re-introduction. We coded subject referents as maintained only if they
had been referred to in the previous clause. To this end, the referents
which had been discontinued across one or more sentences were all
regarded re-introduced into the discourse. On a third tier, we had the
RE type for a single reference with the main categories nominal,
pronominal, zero anaphora, and classifier. Under nominal, we identified
bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we
distinguished among three types of handshapes: WECL, BPCL, and ExtCL.
Finally, we had CA\footnote{Note that CA can also be further subdivided
  into overt, reduced, and subtle CA, following
  \citep{cormier_rethinking_2015}. In overt CA, the signer is ``fully in
  character'' and does not use any other lexical or classifier
  constructions. In reduced CA, the signer can involve a small amount of
  narration or lexical material while tracking a referent using CA.
  However, subtle CA refers to cases where the referent is primarily
  tracked by a non-CA construction (e.g., nominal, pronominal, verbal or
  classifier constructions) but the signer subtly uses a few CA
  articulators at the same time (e.g., enacting the eye-gaze of the
  character). We only coded overt and reduced types as reference
  tracking by CA, and disregarded the subtle type. We'd also like to
  acknowledge that mixed CA, or simultaneous tracking of two or multiple
  references through CA and other tools can be possible
  \citep{dudis2004, liddell2003}. In such cases, as we explain here, we
  only coded only the subject of a clause for maintenance and
  re-introduction. If the referent was in subject position and tracked
  by using CA, we considered this an instance of CA, as well.},
agreement, plain tags for zero anaphora. We were unable to observe any
instance of HCLs occurring in isolation (i.e., not part of CA). We
should also highlight here that we coded referent status as nominal if
the clause included an explicit nominal referent. For example, if
constructed action, an index sign, or a classifier preceded or followed
its explicitly signed nominal referent, the clause was regarded as
having nominal reference (i.e., modified noun). The annotation tiers are
presented in Table \ref{tab:tab-2}.

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

In our analysis, we use Bayesian regression models with the help of the
brms package \citep{burkner2018} in R. In a Bayesian regression the
prior and the likelihood is used to calculate the posterior. Instead of
a single filter of significance, we get a posterior probability
distribution. We do not define our own priors and use the brms package
defaults. We are using Bayesian regression to avoid some common pitfalls
of other approaches \citep{wagenmakers2007, kruschke2011}.

For this particular type of experimental data, we fit the responses to a
multinomial distribution. Most of the time, experimental results have
yes/no-true/false responses which are fit to binomial distributions,
reading/response times which are fit to log-normal distributions, and
ratings which are fit to cumulative distributions. In these cases, the
researcher is not particularly interested in the `intercept' coefficient
which shows the mean value of the response. The researcher is usually
interested in the remaining coefficients which give information about
predictor level comparisons. Each coefficient is a comparison. That is
why for a predictor with two levels A and B, the plots or the results
only have ``A'' or ``B'' as labels.

In a multinomial distribution, the responses are in a categorical
relationship where the levels can be A, B, C, and more. These levels are
not ordered in time, they are not cumulative, and they are not additive.
If we were to make an analogy, they are like countries on Earth which
can have differences and similarities, but they do not have to, and they
are treated as distinct entities. In a multinomial distribution, one of
the responses becomes the level of comparison and other levels are
compared against it. In a way, there are multiple binomial
distributions. However, the intercepts in a multinomial distribution end
up having more of an interest for the researcher. This is because the
intercept is a comparison of two levels of the response to the exclusion
of the remaining response levels. There is another consideration that
should go into interpreting the results of a multinomial distribution.
Different from the intercepts, the remaining coefficients are predictor
level comparisons within the grand comparison of the response levels. To
indicate which predictor level comparison belongs to which response
comparison, we name our coefficients with prefixes followed by an
underscore.

In our model plots, we report median estimates as points, 50\% credible
intervals as a thick line and 95\% credible intervals as a thin line.
These numbers are relatively arbitrary, and they are intended to show
the posterior probability distribution of the coefficients. We use
interval plots to save space and have a more compact plot. If a
considerable amount (\textgreater90\%) of the posterior probability
distribution is towards a sign (-/+) we interpret it as a categorial
effect, if a good amount (\textgreater50\%) of the distribution is
towards a sign (-/+) we interpret it as a tendency, and we see the
remaining distributions as indifference. In addition to these, we look
at how the distribution is skewed in the extremes or how `wide' it is
from the median estimate to inform us mainly about uncertainty and
variation.

\hypertarget{results}{%
\section{Results}\label{results}}

We collected 4007 data points as a result of the
experiment.\footnote{For the purposes of this double-blind review, we share all our data and code in an anonymous online repository that can be accessed here: \url{https://anonymous.4open.science/r/reference_tracking_revision-A542/}}
Each data point is a referent tracked by the participant. Each of the
REs is classified in terms of its Discourse Status (Introduction,
Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer,
Zero
Anaphora),\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Zero Anaphora). However, we allude to the numerical distribution of the subtypes in this and following sections.}
and Participant Group (DoD,
DoH).\footnote{Similarly, we consider DoH signers a single group in our analyses and disregard the differing age of acquisition intervals (0-3, 4-7, 8-12, 13-17).}

Across our acquisition groups, we observed 1031 (0.26) instances of
referent introduction, 1589 (0.4) instances of referent maintenance, and
1387 (0.35) instances of referent re-introduction. For the introduction
of a referent, the majority of the REs were nominal constructions (Table
\ref{tab:my-table}). However, we observed a sporadic use of classifiers
and zero anaphora respectively. When the discourse status was
maintained, signers extensively used zero anaphora to refer to the
entities in the immediately previous clause. Classifier constructions
followed zero anaphora in frequency in maintained contexts. We observed
only a few instances of nominal use for referent maintenance. In
re-introduction contexts, signers used a greater proportion of nominal
markings which were followed by zero anaphora, classifier, and
pronominal use.

\begin{table}
\small
\caption{Distribution of all discourse status by RE types}
\label{tab:my-table}
\begin{tabular}{llll}
\hline
                     & Introduced    & Maintained     & Re-introduced \\ \hline
\textit{Nominal:}    & 92.7\% (956) & 5.9\% (93)    & 63.7\% (884) \\
\textit{Pronominal:} & 0.1\% (1)    & 0.3\% (4)     & 2.5\% (35)   \\
\textit{Classifier:} & 5.3\% (55)   & 18.3\% (291)  & 7.8\% (108)     \\
\textit{Null:}       & 1.8\% (19)   & 75.6\% (1201) & 26\% (360) \\ \hline
\end{tabular}%
\end{table}

Table \ref{tab:my-table2} shows that for all discourse statuses the most
numerous nominal type was bare nominal which was followed by nouns
modified by index or classifiers. In number, signers used modified nouns
more in introduction and re-introduction compared to maintenance
settings. Signers produced a total number of 40 pronouns (Table
\ref{tab:my-table3}). A great majority of the pronominal constructions
were used to re-introduce a referent into the discourse. We observed
only a very small amount of pronoun use for referent introduction and
maintenance.

\begin{table}
\small
\caption{Distribution of nominals by discourse status}
\label{tab:my-table2}
\begin{tabular}{lllll}
\hline
                        & Bare Noun  & Modified Noun & FS Noun \\ \hline
\textit{Introduced:}    & 48.7\% (813) & 53\% (134)     & 75\% (9) \\
\textit{Maintained:}    & 5\% (83)  & 4\% (10)      & 0\% (0) \\
\textit{Re-introduced:} & 46.3\% (772) & 43\% (109)     & 25\% (3) \\ \hline
\end{tabular}
\end{table}

\begin{table}
\small
\caption{Distribution of pronominals by discourse status}
\label{tab:my-table3}
\begin{tabular}{ll}
\hline
                        & Pronominal  \\ \hline
\textit{Introduced:}    & 2.5\% (1)   \\
\textit{Maintained:}    & 10\% (4)    \\
\textit{Re-introduced:} & 87.5\% (35) \\ \hline
\end{tabular}
\end{table}

As for the classifier constructions (Table \ref{tab:my-table4}), WECL
was the most common type, followed by BPCL and ExtCL. For the first
mention of a referent, signers mostly preferred to use ExtCL. However,
for referent maintenance and re-introduction WECL was used more compared
to BPCL and ExtCL.

Among the markers of zero anaphora, CA was most frequently employed, and
it was followed by plain verbs and agreement verbs (Table
\ref{tab:my-table5}). Overall, signers used very few markers of zero
anaphora for the introduction of a referent and the majority of the
existing ones were CA. Additionally, signers used CA more for referent
re-introduction than maintenance. We observed similar instances of
agreement verbs for maintained and re-introduced contexts. In contrast,
plain verbs were favored more in the maintenance setting than
re-introduction.

\begin{table}
\small
\caption{Distribution of classifiers by discourse status}
\label{tab:my-table4}
\begin{tabular}{llll}
\hline
                        & WECL          & BPCL        & ExtCL         \\ \hline
\textit{Introduced:}    & 2.2\% (6)    & 2.5\% (3)   & 70.8\% (46) \\
\textit{Maintained:}    & 67.5\% (181) & 77.7\% (94) & 24.6\% (16) \\
\textit{Re-introduced:} & 30.2\% (81)   & 19.8\% (24)   & 4.6\% (3)   \\ \hline
\end{tabular}
\end{table}

Across the discourse statuses, there is also a difference between
participant groups in the use of zero anaphora for reference tracking.
DoD signers of TİD used it more frequently (0.22) compared to the DoH
signers (0.17). Comparably, DoH signers used nominals more frequently
(0.26) than DoD signers (0.22). Figure 1 shows the numerical
distribution of the responses by RE type and Group.

\begin{table}
\small
\caption{Distribution of zero anaphora by discourse status}
\label{tab:my-table5}
\begin{tabular}{llll}
\hline
                        & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
\textit{Introduced:}    & 1.6\% (17)        & 0.3\% (1)    & 0.5\% (1)      \\
\textit{Maintained:}    & 71.5\% (749)      & 89.3\% (302) & 76.9\% (150)  \\
\textit{Re-introduced:} & 26.8\% (281)      & 10.4\% (35)   & 22.6\% (44)   \\ \hline
\end{tabular}
\end{table}

\begin{verbatim}
## Warning: The dot-dot notation (`..count..`) was deprecated in ggplot2 3.4.0.
## i Please use `after_stat(count)` instead.
## This warning is displayed once every 8 hours.
## Call `lifecycle::last_lifecycle_warnings()` to see where this warning was
## generated.
\end{verbatim}

\begin{figure}
\centering
\includegraphics{revised_manuscript_files/figure-latex/fig-count-plot-1.pdf}
\caption{Referent counts by Discourse Status (Introduction, Maintanence,
Re-introduction), RE Type (Nominal, Pronominal, Classifier, Zero
Anaphora), and Participant Group (DoD, DoH)}
\end{figure}

For more inference on the data and having pairwise comparisons, we fit a
regression model to RE Type using discourse status and acquisition group
as predictors. We used the brms package \citep{burkner2018} in R, with a
categorical distribution on the responses and sum contrasts for the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (DoD, DoH). As the overall number of produced
pronouns was very little (40 data points), we excluded pronominal use
from the following analyses. For the current analysis, we did not define
priors ourselves and used the preset uninformed priors. Figure
2\footnote{The regression model results are given in
  \ref{section:modelResult}. We ran another model where we used the
  self-reported acquisition time as a predictor with 3 levels 0-3:Early,
  4-7:Late1, and 7-above:Late2. However, our model did not converge
  properly and the credible intervals were very wide spread, yielding
  the results uninformative. We provide the model results as a plot in
  \ref{section:d}.} shows the posterior probability distributions where
the dot is the median estimate, the thick line is the 50\% credible
intervals, and the thin line is the 95\% credible intervals. There are 3
responses we are considering in this model. They are Classifier,
Nominal, and Zero Anaphora. Classifier response is the level that other
responses Nominal and Zero Anaphora are compared against.

\begin{figure}
\centering
\includegraphics{revised_manuscript_files/figure-latex/model-gram-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Zero,
Classifier) with the predictors Discourse Status (Introduction,
Maintenance, Re-introduction) and Participant Group (DoD, DoH)}
\end{figure}

In a multinomial distribution, the model takes one of the responses as a
reference and compares all other responses to it. For example, the first
coefficient, Nom\_Intercept (nominal intercept), in Figure 2 is the
comparison of a nominal versus a classifier. It means that participants
produced more nominals than classifiers in general regardless of
participant group and discourse status. The second coefficient shows
that zero anaphora was also used more than classifiers by the
participants in general, but not as much as nominals.

In the comparison of nominals and classifiers, acquiring TİD from deaf
signing caregivers did not have an effect. Nominals were used more when
the discourse status was introduction compared to re-introduction.
Again, the whole comparison is made against a classifier. In contrast to
the introduction status, nominals were used less in maintenance contexts
compared to the re-introduction contexts within the greater comparison
to classifier referents. There seems to be an interaction between the
introduction status of discourse and being a DoD signer of TİD. When the
acquirer is a DoD signer and the discourse function is introduction, the
chances of observing a nominal referent gets
higher.\footnote{Remember that this is an interaction term and the grand comparison is nominal vs classifier use.}
Also, there seems to be an interaction between maintaining a referent
and being a DoD signer of TİD. When it was a DoD signer of TİD and the
discourse context was maintenance, the chances of observing a nominal
referent gets
lower.\footnote{This coefficient too is an interaction term and needs to be considered withing the comparison of nominal versus classifier use.}

Compared to classifiers, being a DoD signer of TİD increased the odds of
using zero anaphora in general. Zero anaphora constructions were used
less in introduction contexts than in re-introduction. This contrasts
with the nominals (4th coefficient) as they were used more in referent
introduction versus re-introduction. In contrast to introduction
contexts, zero anaphora was used more in referent maintenance compared
to re-introduction, yet again within the greater comparison to the
classifiers. There seems to be an interaction between introducing a
referent and being a DoD signer of TİD. In that, the odds of using zero
anaphora increases if the participant is a DoD signer of TİD and the
referent is used in an introduction context. There is another
interaction with regards to referent maintenance, contrasting the first
interaction. In that, the odds of using zero anaphora decrease if the
participant is a DoD signer of TİD and the referent is maintained. This
does not mean that DoH signers used more zero anaphora in maintenance
contexts. It means that the use of zero anaphora did not linearly
increase as the sum of using a zero anaphora marker as a DoD signer plus
using zero anaphora for maintaining a referent versus re-introducing it.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In our study, we aimed to (i) describe how signers track referents in
signed discourse in TİD and (ii) examine whether late acquisition of TİD
results in more over-redundant referent selection. To this end, we
investigated the distribution of referring expressions (REs) that the
DoD and DoH signers select when employing referents in different
discourse statuses (introduced, maintained, and re-introduced).

In our first set of hypotheses regarding referent tracking strategies,
we had the general expectation that signers would use more explicit
forms (mainly nominals) for referent introduction and less implicit
forms such as zero anaphora, pronominals, and classifiers for
maintenance. In re-introduced contexts, we predicted that signers would
use both explicit and implicit markers depending on the accessibility of
the tracked referent.

For age of acquisition effects, based on what is proposed for child and
adult deaf signers with delayed language exposure in terms of their
morphosyntactic \citep{boudreault2006, cormier2012, kayaba2022} and
narrative development \citep{becker2009, cormier2013}, we expected that
DoH signers would be more over-redundant in their referent tracking,
perhaps akin to second language learners
\citep{frederiksen2019, bel2015}. That is, we had the expectation that
DoH signers would use more explicit markers like nouns and fewer
implicit markers like zero anaphora even for highly accessible referents
(when the discourse status is maintained or even re-introduced in same
cases) than DoD signers.

\hypertarget{tracking-referents-in-tid}{%
\subsection{Tracking referents in TİD}\label{tracking-referents-in-tid}}

The results of the present study only partly support our hypotheses. In
line with our expectations, regardless of the participant groups (DoD or
DoH) (i) signers mainly used nominals to introduce a referent into the
discourse, (ii) used zero anaphora and classifiers but not pronouns for
maintained contexts, and (iii) employed both nominals and less explicit
markers (classifiers, zero anaphora and pronouns) to re-introduce a
referent.

However, the proportion of the observed discourse statuses (26\% for
introduction, 40\% for maintenance, 34\% for re-introduction across
acquisition groups) in this study differed from that of some other
studies. For a comparison, while the percentage of introduced referents
were similar in both research (26\% versus 24\%), the participants in
the present study brought back referents into the discourse more
frequently compared to the participants in \citet{frederiksen2019} and
\citet{frederiksen2016} which reported that re-introduction contexts
accounted for only 7\% of the referents in ASL. On the other hand,
maintenance percentages were lower in the present study compared to
\citet{frederiksen2016} (40\% versus 69\%). Investigating the
distribution of REs across discourse statuses in ASL, a similar study by
\citet{czubek2017} also reported contrasting numbers (17\% for
introduction, 64\% for maintenance, and 19\% for re-introduction). We
suggest that the discrepancy might be due to the nature of the stimuli.
In \citet{frederiksen2016}, the authors used a 4-6-episode-long picture
story with 2 animate characters and 1 inanimate object. In contrast, the
stimuli in the present study consisted of 8-episode-long video clips on
average with a mean number of 2 animate characters and 3 inanimate
objects. Although the number of animate characters and inanimate
characters in Czubek's \citeyearpar{czubek2017} stimuli matched that of
the present study, their narrative structure (a 6-episode-long picture
story) was still shorter compared to our narrative stimuli. We believe
fewer possible referents and/or a simpler narrative structure in these
studies may have contributed to the lower number of re-introduced
contexts compared to the present study. Due to the story length and
composition in this study, the participants had to track more recurring
animate or inanimate referents for a longer period of time. As a result,
our participants might have engaged in referent re-introduction more
than what was reported in previous studies
\citep{czubek2017, frederiksen2016}.

As for the distribution of REs across different discourses statuses, we
observed that signers overwhelmingly preferred to use nominals over
other REs for the first mention of a referent and used zero anaphora
such as CA, agreement and plain verb for referents with high
accessibility (i.e., when the discourse status was maintained). This
finding parallels other accounts of referential accessibility suggested
for other sign languages (see \citet{frederiksen2016};
\citet{czubek2017}; and \citet{swabey2002}; for ASL, \citet{morgan2005};
for BSL \citet{bel2015}; \citet{barbera2009} for LSC;
\citet{ferrara2022} for other sign languages). Our results also indicate
that modified nouns (e.g., nouns that are preceded or followed by a
classifier or IX) appeared more in introduced contexts compared to
re-introduced contexts. This indicates that modified nouns are used more
to track referents with lower accessibility, probably due to their more
explicit structure.

Interestingly and to our surprise, the signers in this study used
pronominal IX very sporadically with only 40 observed instances that
were mainly used to re-introduce referents rather than maintain them.
This finding is comparable with what has been reported for ASL
\citep{swabey2002, frederiksen2016, czubek2017} but not with LSC
\citep{bel2015}, BSL \citep{morgan2005} and DGS \citep{perniss2015}, in
which the signers extensively made overt pronominal references (up to
30\%) to maintain entities. Comparable to our findings, it has been
reported that pronouns do not frequently occur in the local context and
instead follows a topic shift and thus rank higher than classifiers and
zero anaphora \citetext{\citealp[see][ for
LSC]{barbera2009}; \citealp[and][ for TİD]{nuhbalaoglu2018}}.
Considering this, the claim that the IX sign functions like
demonstratives that are also used to track referents in re-introduced
contexts has not been borne out \citep{koulidobrova2016}. Similar to
these research, our findings indicate that the pronominal IX in TİD was
used more concurrently with nominal constructions and pronominal IX was
used very infrequently in isolation. The observation that overt pronouns
are rare in ASL narratives comes from studies that include a simple
picture retelling task with one protagonist
\citep[e.g.,][]{frederiksen2016} as well as a more complex narrative
design as in \citep{czubek2017}. However, \citet{czubek2017} also
incorporated different genres and reported more instances of overt
pronouns in personal narratives (compared to picture or video
retelling). In fact, this finding suggests that the use of pronouns
might be genre-dependent. Moreover, another production study on referent
tracking in ASL reported object preference for pronominals
\citep{frederiksen2017}. However, we only coded subjects for maintained
and re-introduced referents. We argue that all of these factors
(genre-dependency as well as object preference) might contribute to the
explanation of why overt pronouns in our data are dispreferred in favor
of other markers for highly accessible referents such as zero anaphora
and classifiers.

A deeper look into the allocation of classifier handshapes revealed that
WECLs were the most common type of classifiers for the maintenance of a
referent followed by BPCLs and ExtCLs. We observed a similar distibution
for re-introduced contexts. In contrast, a great number of the overall
ExtCLs were used to introduce a new referent into the discourse. That is
to say, whereas WECLs were the preferred classifier types to signal
highly accessible referents, ExtCL constructions were mainly used for
referents which were not easily accessible. We suggest that the less
iconic and transparent nature of WECLs allowed signers to continue a
previous referent without having to use an explict form. As pointed out
by \citet{czubek2017}, signers prefer to rely on ExtCLs for the first
mention of an entity because these constructions are informative and
explicit by nature, and they provide visual and descriptive cues about
an entity and its location in space. We only observed 3 instances of
ExtCLs for re-introduced contexts, replicating Frederiksen and
Mayberry's \citeyearpar{frederiksen2019} findings for ASL.

As for the distribution of the zero anaphora, approximately 90\% of the
plain verbs were used to maintain a previous referent and this makes
these constructions least implicit REs used for TİD\footnote{As in other
  sign languages, in TİD, the lexical meaning of signs has an effect on
  the morphological grouping. For instance, usually (di)transitive
  agreement verbs denote some sort of physical or metaphorical transfer
  of an object or abstract notion \citep{meir1999}. As for plain verbs,
  both phonology of a sign and its semantics have an effect on
  morphology. For instance, \textsc{eat} in TİD is signed nearby the
  mouth and since it needs to remain in that phonological location, it
  is not suitable for agreement and thus it is regarded as a plain verb.
  Likewise the experiencer verb \textsc{love} is signed on the torso and
  as such it does not take part in agreement. These generalizations
  noted we should also be cautious in labelling a verb as categorially
  one type or the other. For instance, some intransitive plain verbs
  like \textsc{fall} or \textsc{jump} in TİD may display spatial person
  agreement with the single argument in the clause \citep{gokgoz2020}.
  See \citet{lourenço2018} also for a similar discussion in Brazilian
  Sign Language.}. Agreement verbs and CA were the following implicit
markers of reference, 0.77 and 0.72 of which were used for maintained
contexts, respectively. This contrasts with some of the research
\citep{czubek2017} in which agreement verbs ranked lowest on the scale,
indicating that these constructions were more implicit compared to plain
verbs. However, there seems to be a common ground for the position of CA
markers which have been claimed to rank higher than other markers of
zero anaphora as in the present study
\citep{frederiksen2016, czubek2017, barbera2009}. In Figure
\ref{fig-scale}, we report the distribution of REs in TİD as a function
of referent accessibility and compare our accessibility hierarchy with
that of \citep{frederiksen2016} for ASL. In the hierarchies, the
leftmost REs (e.g., modified and bare nouns for TİD) are used when the
accessibility is low. If the referent is already active in the
addressee's mind and its accessibility is high, then the rightmost
markers (e.g., agreement and plain verbs for TİD) are selected.

\begin{figure}
\caption{Comparing the present study and Frederiksen and Mayberry (2016) in terms of proposed accessibility hierarchies}
\label{fig-scale}
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16) -- (214.92,115.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15) -- (314.89,115.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34) -- (414.89,115.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize M NOM}};
% Text Node
\draw (191,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize BARE NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node
\draw (285,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize WECL \& BPCL}};
% Text Node
\draw (359,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize CA}};
% Text Node
\draw (405,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR}};
% Text Node
\draw (452,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in the present study for TİD
    \end{minipage}

\vspace{1 cm}

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node% Text Node
\draw (333,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR \& CA}};
% Text Node
\draw (445,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in Frederiksen and Mayberry (2016) for ASL
    \end{minipage}
\end{figure}

Both hierarchies illustrated in Figure \ref{fig-scale} treat nominals
and ExtCLs as markers of lowly accessibly referents. Likewise, in both
hierarchies, as markers of highly accessible referents, different types
of zero anaphora markers are further distinguished. Plain verbs rank the
lowest on the scales, preceded by agreement verbs and CA. On a similar
ground, both scales lack pronouns, which occurred very infrequently in
the data. Although \citet{frederiksen2016} did not include other
classifier types in the hierarchy, we propose that WECLs and BPCLs must
rank higher than zero anaphora but lower than ExtCLs that were mainly
used by interlocutors to introduce a referent into the discourse
context.

\hypertarget{effect-of-native-acquisition-on-referent-tracking}{%
\subsection{Effect of native acquisition on referent
tracking}\label{effect-of-native-acquisition-on-referent-tracking}}

The results in this study only somewhat support our second set of
hypotheses that we set for the effects of having signing deaf parents on
participants' RE selection. That is, while being a DoD signer of TİD
increased the odds of using zero anaphora in reference tracking across
different discourse statuses, we failed to find a difference for overall
nominal production. Taking a closer look at the interaction terms, we
observed that DoD signers, compared to DoH signers, were less likely to
use nominal constructions when the discourse status was maintained.
Native acquisition from deaf caregivers further increased the odds of
zero anaphora when the discourse status was introduced. However, we
found the opposite alignment for the interaction of maintenance and
native acquisition. That is, the coefficient for this interaction term
slightly decreased the combined main effects of having a maintained
context and being a DoD signer. This means that these main effects were
greater than the interaction or added effects, pointing out to the
implication that the type of discourse status (introduced, maintained,
re-introduced) accounts for the distribution of REs in TİD narratives
better than native acquisition from deaf caregivers (DoD, DoH) which, in
comparison, had a less salient effect.

While DoH signers were only partially over-redundant in maintained
contexts, they were more so in introduced contexts. In line with our
study, a tendency for limited over-redundancy has been suggested for
late signing children in DGS \citep{becker2009} and late signing adults
in BSL specifically for the use of CA \citep{cormier2013}. As for TİD,
to our knowledge, there is only one study \citep{gur2018} that directly
aimed at examining age of acquisition effects on adult signers'
narrative skills. \citet{gur2018} observed that late signing deaf adults
and children set the scene less frequently and used CA less accurately
compared to their native counterparts. However, this particular
investigation did not present any data on the distribution of REs
between two acquisition groups. In the present study, the finding that
DoD signers preferred to use null arguments even for the first mention
of a referent (19 in total) is surprising given some of the previous
literature in the topicality and saliency framework
\citep{cormier2013, swabey2002, ferrara2022}. These research reported
very little or no use of zero anaphora by DoD signers for introduced
contexts.

Given the paucity of research directed at the DoH or late signing group,
we now turn to the referent tracking strategies observed for second
language learners of a sign language. Our study replicated some of the
findings of the line of research that investigated second language
learners of a sign language \citep{bel2015, frederiksen2019}. The
authors reported that more instances of zero anaphora in the context of
referent introduction was attested for DoD signers than second language
learners. Although using zero anaphora to introduce a referent into the
discourse would normally result in referential ambiguity, it is
important to note that our stimuli consisted of 10 interconnected
narratives with 2 or more recurring characters. It may be possible that
DoD signers, having already introduced the referent with an explicit
nominal marker in a different discourse context, found it felicitous to
refer to the same entity with more implicit markers in the following
narratives even for introduced contexts. Then, our results imply a
difference in pragmatic sensitivity to referential accessibility between
DoD and DoH signers, which is more robustly reflected in introduced
contexts. To test this implication, one could use the same number and
length of stimuli from different cartoons with different characters and
see if the tendency to introduce a referent without an overt nominal for
the first time is dependent on the task or not.

In other contexts, we found only very small differences between the two
participant groups. This means that adult signers with late exposure to
sign language in our study were still able to achieve native-like
competency in terms of reference tracking in simple narratives.
Nevertheless, DoD and DoH signers still exhibit some differences
especially in terms of pragmatic sensitivity to economy of form despite
using TİD as their preferred communication language for similar numbers
of years. Since we only observed a small difference between the two
acquisition groups for the use of certain REs, the aforementioned
findings allude to the implication that the effect on native acquisition
on pragmatic factors should also be considered in different contexts.
For one, both groups of signers can be tested with different genres of
narratives and with a more complex story design. In order to capture
possible variation in the accessibility levels of re-introduced
referents, a more fine-tuned protocol \citep[as in][]{toole1996} can be
used for measuring referent distance and the number of other possible
competitors. Furthermore, our focus in this study was to look at
referent tracking strategies between DoD and DoH signers. However, for
future research, it might also be useful to look at pre and post-lingual
acquisiton of TİD. We only had 3 post-lingual signers of
TİD.\footnote{When we looked at the random effects by participants, they did not differ from their acquisition group (DoD or DoH).}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The present study aimed to investigate the reference tracking strategies
in TİD narratives and whether native and late deaf TİD signers differ
from each other in these strategies. In line with our expectations, we
found that the distribution of REs (nominal, pronominal, classifier, and
zero anaphora) followed from previous accounts of saliency and
referential accessibility. For introduced contexts, signers produced
more nominals and ExtCLs, and to maintain a previous referent, they
mainly used null arguments (in clauses with CA, agreement and plain
verbs) and certain classifiers (e.g., BPCL and WECL) to track highly
accessible referents. We observed only very few instances of pronouns
(mostly produced by DoH signers), unlike our expectations. The ones we
observed were mostly reserved for re-introduced contexts along with
nominals. Based on our findings across the two groups, we proposed a
hierarchy of referent accessibility for TİD. As for the effects of
native acquisition, we observed limited over-redundancy for signers with
delayed language acquisition (i.e., DoH signers) who employed fewer
implicit forms of reference tracking compared to DoD signers. DoD
signers were less likely to use a nominal for maintained contexts and
more likely to use zero anaphora for introduced contexts than DoH
signers.

We conclude that DoH signers have overall native-like competency in
simple narratives, however, following \citet{swabey2002}'s intuition,
our findings also suggest that DoD signers adhere to pragmatic norms and
leverage accessibility better than DoH signers to a certain extent
(given our finding of limited overexplicitness for DoH signers) since
they use less form when the referent is more active in the mind. To
avoid this possible discrepancy between the groups, our results
accentuates the necessity of engaging in signed narrative production
early in life as well as the importance of early intervention for DoH or
late signing deaf children. This is especially a crucial message for
policymakers, educators, and health professionals who work with deaf and
hard-of-hearing children. Providing early and effective access to sign
language from very early on can help bridge the gap and ensure equal
opportunities for language development and success in life for DoH
signers.

\clearpage

\appendix

\hypertarget{participant-background-information}{%
\section{Participant background
information}\label{participant-background-information}}

\label{section:a}

\begin{table}[hbt!]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-2}Participant demographic information based on self-report}

\small{

\begin{tabular}{llllllll}
\toprule
No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{Deaf schools} & \multicolumn{1}{c}{Home language}\\
\midrule
2 & DoD & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign\\
3 & DoD & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign\\
*13 & DoD & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign\\
14 & DoD & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign\\
18 & DoD & M & 0-3 & 28 & 12 & PS, MS & TİD\\
20 & DoD & M & 0-3 & 32 & 12 & PS, MS & TİD\\
22 & DoD & M & 0-3 & 35 & 12 & PS, MS, HS & TİD\\
24 & DoD & F & 0-3 & 23 & 12 & PS, MS, HS & TİD\\
26 & DoD & F & 0-3 & 30 & 12 & PS, MS, HS & TİD\\
27 & DoD & F & 0-3 & 25 & 12 & PS, MS & TİD\\
29 & DoD & F & 0-3 & 24 & 14 & PS, MS, HS & TİD\\
30 & DoD & M & 0-3 & 18 & 12 & PS, MS & TİD\\
33 & DoD & F & 0-3 & 21 & 16 & PS, MS, HS & TİD\\
34 & DoD & M & 0-3 & 19 & 16 & PS, MS, HS & TİD\\
38 & DoD & F & 0-3 & 25 & 12 & PS, MS & TİD\\
*10 & DoH & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR\\
23 & DoH & M & 4-7 & 36 & 12 & MS, HS & TR, home sign\\
4 & DoH & M & 4-7 & 33 & 12 & PS, HS & TR\\
16 & DoH & F & 4-7 & 33 & 12 & PS & TR, home sign\\
21 & DoH & F & 4-7 & 31 & 14 & PS, MS & Home sign\\
25 & DoH & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign\\
35 & DoH & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR\\
36 & DoH & M & 8-12 & 43 & 16 & PS, MS & TR\\
12 & DoH & M & 8-12 & 28 & 16 & MS & TR\\
15 & DoH & F & 8-12 & 24 & 12 & PS, MS, HS & TR\\
19 & DoH & F & 8-12 & 24 & 8 & PS, MS & Home sign\\
28 & DoH & M & 8-12 & 33 & 12 & PS, MS & Home sign\\
31 & DoH & M & 13-17 & 50 & 8 & MS & TR, home sign\\
32 & DoH & M & 13-17 & 31 & 16 & PS, MS, HS & TR\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish. *We excluded Participant 13 from our data analysis because the signer indicated to have deaf parents, thus grouped as DoD, but also reported only home sign as the home language and not TİD. Since our classification of DoD included the criterion of having deaf and TİD-signing parents, we decided to remove this participant's data. Also, Participant 10 reported to use TİD at home despite having hearing caregivers which may go against our classification criterion for being DoH that is to have non-signing hearing caregivers. However, Participant 10 informed us that the parents used very rudimentary sign language and that they often did not understand what he was signing.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\clearpage

\setlength{\columnsep}{10pt}
\twocolumn
\linespread{1}
\justifying
\footnotesize

\hypertarget{episodes-in-the-stimuli}{%
\section{Episodes in the stimuli}\label{episodes-in-the-stimuli}}

\label{section:b}

\textbf{Narration 1}

1. Tom chases Jerry, wearing a wooden plank

2. Tom and Jerry enter a tube

3. Tom falls down and Jerry runs away

4. Tube gets thinner and Tom chases Jerry

5. Jerry gets out and escapes

6. Tom follows Jerry in shape of a stick

\textbf{Narration 2}

1. Tom walks past a female cat sitting on a couch

2. Tom turns to and approaches female cat

3. Female cat blushes

4. Tom offers a fish in aquarium to female cat

5. Female cat rejects offer

6. Tom offers a bird in cage to female cat

7. Episode 5 is repeated

8. Tom leaves by walking in a flirtatious manner

\textbf{Narration 3}

1. Jerry falls down and tries to walk up stairs

2. Tom pulls carpet back

3. Jerry struggles

4. A piano falls down stairs

5. Jerry manages to escape

6. Tom sees piano and tries to run away

7. Piano smashes into Tom and flattens him

8. Tom falls on ground

\textbf{Narration 4}

1. Tom hides behind a cardboard house

2. Jerry walks to house with a toy mouse

3. Toy leaves Jerry who stops at entrance

4. Toy enters house

5. Toy gets smashed into pieces

6. Jerry is surprised

7. Tom hurts his stomach after swallowing toy

8. Tom looks at his broken teeth on a mirror

9. Tom gets angry and throws down mirror

\textbf{Narration 5}

1. Tom fakes sleeping and suddenly gets up

2. Tom pulls a rope inside a couch

3. Jerry appears holding a bar of soap

4. Jerry washes Tom's mouth with bar and flees

5. Tom spits soap out

6. Jerry escapes to nest with rope

7. Tom catches rope and stars pulling it

8. A mouse trap appears and catches Tom's nose

9. Tom yells and removes trap

\textbf{Narration 6}

1. Tom points a sword at Jerry and catches him

2. Tom looks at Jerry

3. Red spots appear on Jerry's face

4. Tom gets scared and drops Jerry

5. Tom rushes into bathroom

6. Tom washes his hands

7. Tom opens a drawer

8. Tom swallows pills one after another

9. Feeling ill, Tom looks out of window

10. Tom and Jerry look at each other

\textbf{Narration 7}

1. Two male cats chase one another

2. Tom sees a female cat and kisses her

3. Tom hides away when black cat returns

4. Tom pokes his head behind couch again

5. Episode 3 is repeated

6. black cat sees female cat and hugs her

7. Tom appears and black cat kisses Tom

8. Two male cats move away from one another

\textbf{Narration 8}

1. Jerry hits Tom's knee with a hammer

2. Tom suddenly jumps up

3. Tom yells and holds his knee

4. Jerry appraches Tom holding a thermometer

5. Jerry sticks thermometer into Tom's mouth

6. Jerry lights a fire under thermometer

7. Red liquid starts to rise up

8. Thermometer explodes

\textbf{Narration 9}

1. Tom sits on Jerry who tries to get loose

2. Jerry looks around and sees a hat ribbon

3. Jerry finds a needle attached to hat

4. Jerry takes needle and sticks Tom with it

5. Tom turns red and jumps

6. Jerry placesn eedle between two cushions

7. Tom dodges needle

\textbf{Narration 10}

1. Wearing a stethoscope, Tom catches Jerry

2. Jerry yells into stethoscope

3. Tom's head pops up

4. Jerry runs and enter a mouse nest

5. Tom inserts a gun into nest

6. Gun bends over toward Tom

7. Tom fires gun and his head explodes

\clearpage

\onecolumn

\hypertarget{effect-of-narrating-to-the-camera-vs.-experimenter}{%
\section{Effect of narrating to the camera
vs.~experimenter}\label{effect-of-narrating-to-the-camera-vs.-experimenter}}

\label{section:c}

\begin{figure}
\centering
\includegraphics{revised_manuscript_files/figure-latex/model-camera-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Null) with the
predictors Discourse Status (Introduction, Maintenance,
Re-introduction), Participant Group (Native, Late) and Narration Style
(all responses vs.~only camera)}
\end{figure}

\emph{Note:} ``All'' indicates model results including all participants
(both camera and experimenter) whereas ``camera'' indicates the results
from only those who narrated to the camera. That is, the 4 signers who
narrated to the experimenter were from analysis in the second dataset.

\clearpage

\hypertarget{regression-model-results-for-re-type}{%
\section{Regression model results for RE
Type}\label{regression-model-results-for-re-type}}

\label{section:modelResult}

\begin{table}[hbt!]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-4}Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)}

\small{

\begin{tabular}{llllllllll}
\toprule
Coefficients & \multicolumn{1}{c}{outerWidth} & \multicolumn{1}{c}{innerWidth} & \multicolumn{1}{c}{pointEstimate} & \multicolumn{1}{c}{ll} & \multicolumn{1}{c}{l} & \multicolumn{1}{c}{m} & \multicolumn{1}{c}{h} & \multicolumn{1}{c}{hh} & \multicolumn{1}{c}{model}\\
\midrule
Nom\_Intercept & 0.90 & 0.50 & median & 0.86 & 1.06 & 1.18 & 1.31 & 1.50 & All\\
Zero\_Intercept & 0.90 & 0.50 & median & 0.11 & 0.26 & 0.37 & 0.46 & 0.59 & All\\
Nom\_DoD & 0.90 & 0.50 & median & -0.67 & -0.31 & -0.06 & 0.18 & 0.55 & All\\
Nom\_Intro & 0.90 & 0.50 & median & 3.34 & 3.56 & 3.72 & 3.88 & 4.13 & All\\
Nom\_Maintain & 0.90 & 0.50 & median & -5.70 & -5.49 & -5.34 & -5.20 & -4.98 & All\\
Nom\_DoD*Intro & 0.90 & 0.50 & median & -0.09 & 0.34 & 0.65 & 0.98 & 1.44 & All\\
Nom\_DoD*Maintain & 0.90 & 0.50 & median & -1.86 & -1.45 & -1.16 & -0.88 & -0.46 & All\\
Zero\_DoD & 0.90 & 0.50 & median & 0.45 & 0.72 & 0.89 & 1.10 & 1.40 & All\\
Zero\_Intro & 0.90 & 0.50 & median & -4.77 & -4.22 & -3.87 & -3.55 & -3.13 & All\\
Zero\_Maintain & 0.90 & 0.50 & median & 1.75 & 1.99 & 2.17 & 2.37 & 2.66 & All\\
Zero\_DoD*Intro & 0.90 & 0.50 & median & 0.10 & 0.93 & 1.58 & 2.26 & 3.32 & All\\
Zero\_DoD*Maintain & 0.90 & 0.50 & median & -1.95 & -1.38 & -0.99 & -0.64 & -0.17 & All\\
\bottomrule
\end{tabular}

}

\end{threeparttable}
\end{center}

\end{table}

\clearpage

\hypertarget{additional-model-with-3-levels-for-nativeness}{%
\section{Additional model with 3 levels for
Nativeness}\label{additional-model-with-3-levels-for-nativeness}}

\label{section:d}

\begin{table}[hbt!]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-5}Model results for Nativeness with 3 levels}

\small{

\begin{tabular}{llllll}
\toprule
Coefficients & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{EstError} & \multicolumn{1}{c}{lower95} & \multicolumn{1}{c}{upper95} & \multicolumn{1}{c}{Rhat}\\
\midrule
Nom\_Intercept & 1.24 & 0.21 & 0.83 & 1.68 & 1.00\\
Null\_Intercept & -4.77 & 4.19 & -16.09 & 0.09 & 1.01\\
Nom\_Late1vsNative & 0.43 & 0.57 & -0.71 & 1.56 & 1.00\\
Nom\_Late2vsLate1 & -0.34 & 0.51 & -1.32 & 0.68 & 1.01\\
Nom\_Intro & 3.76 & 0.26 & 3.27 & 4.27 & 1.00\\
Nom\_Maintain & -5.24 & 0.22 & -5.67 & -4.80 & 1.00\\
Nom\_L1vsN*Intro & 1.05 & 0.67 & -0.23 & 2.41 & 1.00\\
Nom\_L2vsL1*Intro & -0.07 & 0.67 & -1.41 & 1.22 & 1.00\\
Nom\_L1vsN*Maintain & -0.35 & 0.57 & -1.49 & 0.75 & 1.00\\
Nom\_L2vsL1*Maintain & -0.87 & 0.54 & -1.89 & 0.20 & 1.00\\
Null\_Late1vsNative & -15.00 & 12.57 & -48.88 & -0.48 & 1.01\\
Null\_Late2vsLate1 & 15.88 & 12.57 & 1.30 & 49.79 & 1.01\\
Null\_Intro & -24.10 & 16.74 & -69.41 & -4.68 & 1.01\\
Null\_Maintain & 12.30 & 8.37 & 2.60 & 35.02 & 1.01\\
Null\_L1vsN*Intro & -60.77 & 50.28 & -195.99 & -2.68 & 1.01\\
Null\_L2vsL1*Intro & 61.91 & 50.26 & 3.62 & 197.45 & 1.01\\
Null\_L1vsN*Maintain & 29.89 & 25.14 & 0.81 & 97.46 & 1.01\\
Null\_L2vsL1*Maintain & -30.89 & 25.13 & -98.46 & -1.73 & 1.01\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Rhat values above 1 indicate that the regression model chains did not converge properly. There were 4 chains in our regression model.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\clearpage

\renewcommand\refname{References}
\bibliography{./bibliography/references.bib}

\end{document}